---
title: "Clemson Gymnastics Analysis Code"
author: "Alexander Harriman"
date: "2024-04-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Packages

```{r, echo = FALSE, results = 'hide', warning = FALSE}
library(readxl)
library(dplyr)
library(gt)
library(ggplot2)
library(tidyverse)
library(readr)
library(data.table)
library(writexl)
```


# Question 1

Insights are based on the GYMNASTICS_DATA dataset.

```{r, warning = FALSE}
# Load all sheets into a list of dataframes

allSheets <- function(file, tibble = FALSE){
  sheetNames <- excel_sheets(file)
  x <- lapply(sheetNames, 
              function(y){
    read_excel(file, sheet = y)
  })
  names(x) <- sheetNames
  x
}


allGymnastics <- allSheets('GYMNASTICS_DATA.XLSX')
```


## Number of Scanned In Customers

To check how many customers scanned into each meet, the ATTENDANCE spreadsheet will be used.


```{r}
customers <- allGymnastics[[3]]

#Convert distance travelled to a numeric variable

customers <- customers |>
  mutate(DISTANCE_TRAVELLED_IN_MILES = as.numeric(DISTANCE_TRAVELLED_IN_MILES))
```


```{r}
# Filter the customer data down to only gymnastics

gymCustomers <- customers |>
  filter(SPORT_NAME == "Women's Gymnastics")
```

During the 2024 Women's Gymnastics season, 43,498 tickets were sold according to the database.


```{r}
#Find the number of scanned in tickets

allScan <- gymCustomers |>
  filter(is.na(SCAN_TIME) == FALSE)

dim(allScan)[1]
```

During the 2024 Women's Gymnastics season, 31,168 tickets were scanned into the event.


### Tickets Sold and Scanned, 2024 Season

```{r}
seasonTable <- data.frame(
  meet = 'Full Season',
  date = '2024', 
  totalTickets = dim(gymCustomers)[1],
  scanTickets = dim(allScan)[1]
) |>
  mutate(scanRate = round((100.0*scanTickets / totalTickets), 1))
```

#### Per Meet

For additional context on how well Clemson did on getting sold tickets into the stadium as the season progressed. This information will also inform factors on meet attendance (Part 3 of Question 1).


```{r}
totalTicketsMeet <- gymCustomers |>
  group_by(ATHLETICS_EVENT_ID) |>
  summarise(date = as.character(ATHLETICS_EVENT_DATE)[1],
            totalTickets = length(PRICE),
            scanTickets = sum(ifelse(is.na(SCAN_TIME) == FALSE, 1, 0))) |>
  mutate(scanRate = round((100.0*scanTickets / totalTickets),1),
         meet = case_when(
           ATHLETICS_EVENT_ID == 319 ~ 'UNC',
           ATHLETICS_EVENT_ID == 320 ~ 'Air Force',
           ATHLETICS_EVENT_ID == 323 ~ 'W&M',
           ATHLETICS_EVENT_ID == 324 ~ 'Pitt',
           ATHLETICS_EVENT_ID == 325 ~ 'NC State'
         )) |>
  select(meet, date, totalTickets, scanTickets, scanRate)
```



```{r}
fullData <- rbind(totalTicketsMeet, seasonTable) |>
  arrange(factor(meet, levels = c('W&M', 'Pitt', 'NC State', 'UNC', 'Air Force', 'Full Season'))) |>
  select(meet, totalTickets, scanTickets, scanRate)


scannedTicketTable <- fullData |>
  gt() |>
  tab_header(title = md('**Gymnastics Attendance Figures, 2024 Season**')) |>
  cols_align(align = 'center', columns = c('meet', 'totalTickets', 'scanTickets', 'scanRate')) |>
  cols_label(meet = 'Meet', totalTickets = 'Tickets Sold', scanTickets = 'Tickets Scanned', scanRate = 'Scanned%') |>
  opt_stylize(style = 5, color = 'gray') |>
  cols_width(everything() ~ px(140)) |>
  tab_style(style = list(
    cell_text(weight = 'bold')
  ),
  locations = cells_body(
    rows = meet == 'Full Season'
  )) |>
  tab_style(
    style = cell_borders(
      sides = c('top'),
      color = 'black',
      weight = px(1.5),
      style = 'solid'
    ),
    locations = cells_body(rows = meet == 'Full Season') 
  )

scannedTicketTable
```

#### Overall Clemson Scanned Ticket Rate

Used as a comparison tool for how effective Clemson Gymnastics was at getting scanned tickets.


```{r}
allSports <- customers |>
  group_by(SPORT_NAME) |>
  summarise(n = n(),
            scans = sum(ifelse(is.na(SCAN_TIME) == FALSE, 1, 0))) |>
  mutate(scanRate = round((100.0*scans / n), 1)) |>
  arrange(desc(scanRate))

totalAverage <- data.frame(
  SPORT_NAME = 'All Sports',
  n = dim(customers)[1],
  scans = sum(ifelse(is.na(customers$SCAN_TIME) == FALSE, 1, 0)),
  scanRate = round((100.0*sum(ifelse(is.na(customers$SCAN_TIME) == FALSE, 1, 0))) / dim(customers)[1], 1)
)
```



```{r}
fullSports <- rbind(allSports, totalAverage)

scannedTicketTable2 <- fullSports |>
  gt() |>
  tab_header(title = md('**Clemson Scanned Ticket Rates, Spring 2024**')) |>
  cols_align(align = 'center', columns = c('SPORT_NAME', 'n', 'scans', 'scanRate')) |>
  cols_label(SPORT_NAME = 'Sport', n = '# Tickets', scans = '# Scanned', scanRate = 'Scanned%') |>
  opt_stylize(style = 5, color = 'gray') |>
  cols_width(SPORT_NAME ~ px(200), everything() ~ px(125)) |>
  tab_footnote('Accurate as of 16 April.') |>
  tab_style(style = list(
    cell_text(weight = 'bold')
  ),
  locations = cells_body(
    rows = SPORT_NAME %in% c("Women's Gymnastics", 'All Sports')
  )) |>
  tab_style(
    style = cell_borders(
      sides = c('top'),
      color = 'black',
      weight = px(1.5),
      style = 'solid'
    ),
    locations = cells_body(rows = SPORT_NAME == 'All Sports') 
  )

scannedTicketTable2
```

## Scanned in On Time Rates by Sport

Looking at how often people who scanned into women's gymnastics events scanned in on time, relative to the rest of the spring sports.


```{r}
allSports <- customers |>
  filter(is.na(SCAN_TIME) == FALSE) |>
  group_by(SPORT_NAME) |>
  summarise(scans = n(),
            onTime = sum(ifelse(SCAN_MINS_BEFORE_START > 0, 1, 0))) |>
  mutate(onTimeRate = round((100.0*onTime / scans), 1)) |>
  arrange(desc(onTimeRate))

totalAverage <- data.frame(
  SPORT_NAME = 'All Sports',
  scans = sum(ifelse(is.na(customers$SCAN_TIME) == FALSE, 1, 0)),
  onTime = sum(ifelse(is.na(customers$SCAN_TIME) == FALSE & customers$SCAN_MINS_BEFORE_START > 0, 1, 0)),
  onTimeRate = round(100.0*
    sum(ifelse(is.na(customers$SCAN_TIME) == FALSE & customers$SCAN_MINS_BEFORE_START > 0, 1, 0))
    / sum(ifelse(is.na(customers$SCAN_TIME) == FALSE, 1, 0)), 1)
)

onTimeInfo <- rbind(allSports, totalAverage)
```


```{r}
onTimeTable <- onTimeInfo |>
  gt() |>
  tab_header(title = md('**Clemson Tickets Scanned on Time, Spring 2024**')) |>
  cols_align(align = 'center', columns = c('SPORT_NAME', 'scans', 'onTime', 'onTimeRate')) |>
  cols_label(SPORT_NAME = 'Sport', scans = '# Scanned', onTime = '# On Time', onTimeRate = 'On Time%') |>
  opt_stylize(style = 5, color = 'blue') |>
  cols_width(SPORT_NAME ~ px(200), everything() ~ px(125)) |>
  tab_footnote('Accurate as of 16 April.') |>
  tab_style(style = list(
    cell_text(weight = 'bold')
  ),
  locations = cells_body(
    rows = SPORT_NAME %in% c("Women's Gymnastics", 'All Sports')
  )) |>
  tab_style(
    style = cell_borders(
      sides = c('top'),
      color = 'black',
      weight = px(1.5),
      style = 'solid'
    ),
    locations = cells_body(rows = SPORT_NAME == 'All Sports') 
  )

onTimeTable
```

### Per Meet


```{r}
onTimeMeet <- gymCustomers |>
  filter(is.na(SCAN_TIME) == FALSE) |>
  group_by(ATHLETICS_EVENT_ID) |>
  summarise(scanTickets = n(),
            onTimeTickets = sum(ifelse(SCAN_MINS_BEFORE_START > 0, 1, 0))) |>
  mutate(onTimeRate = round((100.0*onTimeTickets / scanTickets),1),
         meet = case_when(
           ATHLETICS_EVENT_ID == 319 ~ 'UNC',
           ATHLETICS_EVENT_ID == 320 ~ 'Air Force',
           ATHLETICS_EVENT_ID == 323 ~ 'W&M',
           ATHLETICS_EVENT_ID == 324 ~ 'Pitt',
           ATHLETICS_EVENT_ID == 325 ~ 'NC State'
         )) |>
  select(meet, scanTickets, onTimeTickets, onTimeRate)
```



```{r}
scannedGym <- gymCustomers |>
  filter(is.na(SCAN_TIME) == FALSE)

seasonOnTimeTable <- data.frame(
  meet = 'Full Season',
  scanTickets = dim(scannedGym)[1],
  onTimeTickets = sum(ifelse(scannedGym$SCAN_MINS_BEFORE_START > 0, 1, 0))
) |>
  mutate(onTimeRate = round((100.0*onTimeTickets / scanTickets), 1))
```


```{r}
fullOnTime <- rbind(onTimeMeet, seasonOnTimeTable) |>
  arrange(factor(meet, levels = c('W&M', 'Pitt', 'NC State', 'UNC', 'Air Force', 'Full Season'))) |>
  select(meet, scanTickets, onTimeTickets, onTimeRate)


onTimeTable2 <- fullOnTime |>
  gt() |>
  tab_header(title = md('**Gymnastics Tickets Scanned on Time by Meet, 2024**')) |>
  cols_align(align = 'center', columns = c('meet', 'scanTickets', 'onTimeTickets', 'onTimeRate')) |>
  cols_label(meet = 'Meet', scanTickets = '# Scanned', onTimeTickets = '# On Time', onTimeRate = 'On Time%') |>
  opt_stylize(style = 5, color = 'blue') |>
  cols_width(everything() ~ px(125)) |>
  tab_style(style = list(
    cell_text(weight = 'bold')
  ),
  locations = cells_body(
    rows = meet == c("Full Season")
  )) |>
  tab_style(
    style = cell_borders(
      sides = c('top'),
      color = 'black',
      weight = px(1.5),
      style = 'solid'
    ),
    locations = cells_body(rows = meet == 'Full Season') 
  )

onTimeTable2

```



## Reasons for Ticket Distribution


### Scanned Tickets vs. Rankings

```{r}
#End of Regular Season Rankings of All Opponents


fullRankingsData <- fullData |>
  mutate(ranking = case_when(
    meet == 'W&M' ~ 63,
    meet == 'Pitt' ~ 41,
    meet == 'NC State' ~ 16,
    meet == 'UNC' ~ 42,
    meet == 'Air Force' ~ 56
  )) |>
  filter(meet != 'Full Season') |>
  select(meet, ranking, scanRate) |>
  arrange(desc(scanRate))
```

```{r}
rankingsTable <- fullRankingsData |>
  gt() |>
  tab_header(title = md('**Scan Rate by Road to Nationals Ranking, 2024**')) |>
  cols_align(align = 'center', columns = c('meet', 'ranking', 'scanRate')) |>
  cols_label(meet = 'Meet', ranking = 'Ranking', scanRate = 'Scanned%') |>
  opt_stylize(style = 5, color = 'gray') |>
  cols_width(everything() ~ px(125)) |>
  tab_footnote(footnote = 'Rankings Taken at End of Regular Season (3-25).', locations = cells_column_labels(columns = ranking))

rankingsTable
```



### Scanned Tickets vs. Proximity to Stadium

```{r}
bucketedProx <- gymCustomers |>
  mutate(LocBucket = case_when(
    DISTANCE_TRAVELLED_IN_MILES < 5 ~ 'Locals',
    DISTANCE_TRAVELLED_IN_MILES >= 5 & DISTANCE_TRAVELLED_IN_MILES < 40 ~ 'Nearby Residents',
    DISTANCE_TRAVELLED_IN_MILES > 40 ~ 'Vacationers',
    is.na(DISTANCE_TRAVELLED_IN_MILES) ~ 'Unknown'
  ),
    meet = case_when(
      ATHLETICS_EVENT_ID == 319 ~ 'UNC',
      ATHLETICS_EVENT_ID == 320 ~ 'Air Force',
      ATHLETICS_EVENT_ID == 323 ~ 'W&M',
      ATHLETICS_EVENT_ID == 324 ~ 'Pitt',
      ATHLETICS_EVENT_ID == 325 ~ 'NC State'
      )
  ) |>
  filter(is.na(SCAN_TIME) == FALSE)

proxData <- bucketedProx |>
  group_by(meet) |>
  summarise(scanTickets = sum(ifelse(is.na(SCAN_TIME) == FALSE, 1, 0)),
            #scanTickets = n(),
            campus = sum(ifelse(LocBucket == 'Locals', 1, 0)),
            locals = sum(ifelse(LocBucket == 'Nearby Residents', 1, 0)),
            vacations = sum(ifelse(LocBucket == 'Vacationers', 1, 0)),
            unknown = sum(ifelse(LocBucket == 'Unknown', 1, 0)))

proxRates <- proxData |>
  mutate(campus = round(100.0*campus / scanTickets,1),
         locals = round(100.0*locals / scanTickets,1),
         vacations = round(100.0*vacations / scanTickets,1),
         unknown = round(100.0*unknown / scanTickets,1))
```





```{r}
colnames(proxRates) <- c('meet', 'scanTickets', 'Campus', 'Locals', 'Vacations', 'Unknown')

pivotLonger <- proxRates |>
  select(meet, Campus, Locals, Vacations, Unknown) |>
  pivot_longer(-meet)

pivotLonger$meet <- factor(pivotLonger$meet, levels = c('W&M', 'Pitt', 'NC State', 'UNC', 'Air Force'))
pivotLonger$name <- factor(pivotLonger$name, levels = c('Campus', 'Locals', 'Vacations', 'Unknown'))


ggplot(pivotLonger, aes(x = meet, y = value, fill = name)) +
  geom_col(position = 'dodge') +
  xlab('Meet') +
  ylab('% of Scanned Tickets') +
  ggtitle('Scanned Gymnastics Tickets by Distance from Littlejohn') +
  theme(legend.title = element_blank()) +
  scale_fill_manual('legend', values = c('Campus' = '#522D80', 'Locals' = '#F56600', 'Vacations' = '#00205B', 'Unknown' = '#C8C9C7'))
```





### Scanned Tickets vs. Ticket Acquisition Method



```{r}
bucketedAcq <- gymCustomers |>
  mutate(meet = case_when(
      ATHLETICS_EVENT_ID == 319 ~ 'UNC',
      ATHLETICS_EVENT_ID == 320 ~ 'Air Force',
      ATHLETICS_EVENT_ID == 323 ~ 'W&M',
      ATHLETICS_EVENT_ID == 324 ~ 'Pitt',
      ATHLETICS_EVENT_ID == 325 ~ 'NC State'
      )
  ) |>
  filter(is.na(SCAN_TIME) == FALSE)

acqData <- bucketedAcq |>
  group_by(meet) |>
  summarise(scanTickets = n(),
            primary = sum(ifelse(TICKET_ACQUISITION_TYPE == 'Primary', 1, 0)),
            secondary = sum(ifelse(TICKET_ACQUISITION_TYPE == 'Secondary', 1, 0)),
            forward = sum(ifelse(TICKET_ACQUISITION_TYPE == 'Forward', 1, 0)),
            comp = sum(ifelse(TICKET_ACQUISITION_TYPE == 'Comp', 1, 0)))

acqRates <- acqData |>
  mutate(primary = round(100.0*primary / scanTickets,1),
         secondary = round(100.0*secondary / scanTickets,1),
         forward = round(100.0*forward / scanTickets,1),
         comp = round(100.0*comp / scanTickets,1)) |>
  select(meet, primary, secondary, forward, comp)

```




```{r}
colnames(acqRates) <- c('Meet', 'Primary', 'Secondary', 'Forward', 'Comp')

pivotLonger2 <- acqRates |>
  pivot_longer(-Meet)

pivotLonger2$Meet <- factor(pivotLonger2$Meet, levels = c('W&M', 'Pitt', 'NC State', 'UNC', 'Air Force'))
pivotLonger2$name <- factor(pivotLonger2$name, levels = c('Primary', 'Secondary', 'Forward', 'Comp'))

ggplot(pivotLonger2, aes(x = Meet, y = value, fill = name)) +
  geom_col(position = 'dodge') +
  xlab('Meet') +
  ylab('% of Scanned Tickets') +
  ggtitle('Scanned Gymnastics Tickets by Acquisition Method') +
  theme(legend.title = element_blank()) +
  scale_fill_manual('legend', values = c('Primary' = '#522D80', 'Secondary' = '#F56600', 'Forward' = '#C8C9C7', 'Comp' = '#546223'))
```


### How Often Each Ticket Type is Scanned/On Time: Gymnastics


```{r}
typeTable <- gymCustomers |>
  group_by(TICKET_ACQUISITION_TYPE) |>
  summarise(tickets = n(),
            scanTickets = sum(ifelse(is.na(SCAN_TIME) == FALSE, 1, 0)),
            onTimeTickets = sum(ifelse(is.na(SCAN_TIME) == FALSE & SCAN_MINS_BEFORE_START > 0, 1, 0)))

typeTable <- typeTable |>
  mutate(scanRate = round((100.0*scanTickets) / tickets, 1),
         onTimeRate = round((100.0*onTimeTickets) / scanTickets, 1)) |>
  select(TICKET_ACQUISITION_TYPE, tickets, scanRate, onTimeRate)

typeTable
```


```{r}
cleanType <- typeTable |>
  gt() |>
  tab_header(title = md('**Gymnastics Attendee Metrics by Ticket Type**')) |>
  cols_align(align = 'center', columns = c('TICKET_ACQUISITION_TYPE', 'tickets', 'scanRate', 'onTimeRate')) |>
  cols_label(TICKET_ACQUISITION_TYPE = 'Ticket Type', tickets = '# Sold', scanRate = 'Scanned%', onTimeRate = 'On Time%') |>
  opt_stylize(style = 5, color = 'green') |>
  cols_width(everything() ~ px(125)) |>
  tab_footnote(footnote = '% of Scanned Tickets Entered on Time.', locations = cells_column_labels(columns = onTimeRate))

cleanType
```




# Question 2


In this section, the on-floor performance of Clemson Gymnastics will be analyzed.


## Combining Meet Data


```{r, warning = FALSE, echo = TRUE, results = 'hide', message = FALSE}
meetFiles <- list.files(path = '.', pattern = '.csv')

mergedData <- list()

for(i in 1:length(meetFiles)){
  inputFile <- read_delim(meetFiles[i], locale = locale(encoding = 'windows-1252'))
  colnames(inputFile) <- lapply(colnames(inputFile), gsub, pattern = 'Ê', replacement = '')
  inputFile[] <- lapply(inputFile, gsub, pattern = 'Ê', replacement = '')
  mergedData[[i]] <- inputFile
}
```



```{r, warning = FALSE}
# Create a single format for column names

# The long files have two Event columns: one for the name of the match and one for the real event.
## Edit these column names for formatting:

for(i in 1:length(mergedData)){
  if('Event...5' %in% colnames(mergedData[[i]])){
    mergedData[[i]] <- mergedData[[i]] |>
      rename_at('Event...5', ~'Event') |>
      rename_at('Event...14', ~'Match')
  }
  else{
    next
  }
}

# The newly-made 'Event' column does not have the same syntax as the smaller data frames with EVENT.
## Regulate syntax:

for(i in 1:length(mergedData)){
  if(dim(mergedData[[i]])[2] > 10){
    mergedData[[i]] <- mergedData[[i]] |>
      mutate(Event = case_when(
        Event == 'VT' ~ 'VAULT',
        Event == 'UB' ~ 'BARS',
        Event == 'BB' ~ 'BEAM',
        Event == 'FX' ~ 'FLOOR'
        )
      )
  }
  else{
    next
  }
}

### Commonly found inaccuracies:


# EVENT instead of Event
# Gymnast instead of Name
# Meets with only one SV have SV; meets with 2+ have SV1 instead

# Edit column names based on above format:

for(i in 1:length(mergedData)){
  if('EVENT' %in% colnames(mergedData[[i]])){
    mergedData[[i]] <- mergedData[[i]] |>
      rename_at('EVENT', ~'Event')  
  } 
}

for(i in 1:length(mergedData)){
  if('Gymnast' %in% colnames(mergedData[[i]])){
    mergedData[[i]] <- mergedData[[i]] |>
      rename_at('Gymnast', ~'Name')
  }
}

for(i in 1:length(mergedData)){
  if('SV' %in% colnames(mergedData[[i]])){
    mergedData[[i]] <- mergedData[[i]] |>
      rename_at('SV', ~'SV1')
  }
}

# The data file from UNC has all capitalized columns (besides the edited Event)
## The same file used CU instead of Clemson.

### Fix this file:

## Find where in list it is
uncLocator <- which(sapply(meetFiles, function(meetFiles) "Away_meet_UNC.csv" %in% meetFiles))

# Fix columns (should be Place, Name, Team, Score, Event)
for(i in 1:length(mergedData)){
  if(i == uncLocator){
    colnames(mergedData[[i]]) <- c('Place', 'Name', 'Team', 'Score', 'Event')
  }
}

# Add in date and match information where not available

for(i in 1:length(meetFiles)){
  name <- meetFiles[i]
  for(j in 1:dim(mergedData[[i]])[1]){
    mergedData[[i]]$Match[j] <- case_when(
      name == 'ACC_champ.csv' ~ 'ACC Championship',
      name == 'Away_meet_NCSU.csv' ~ 'at NC State',
      name == 'Away_meet_PITT.csv' ~ 'at Pitt',
      name == 'Away_meet_QUAD.csv' ~ 'at TWU',
      name == 'Away_meet_UCLA.csv' ~ 'at UCLA',
      name == 'Away_meet_UNC.csv' ~ 'at UNC',
      name == 'Home_meet_AF.csv' ~ 'vs Air Force',
      name == 'Home_meet_NCSU.csv' ~ 'vs NC State',
      name == 'Home_meet_PITT.csv' ~ 'vs Pitt',
      name == 'Home_meet_UNC.csv' ~ 'vs UNC',
      name == 'NCAA_champ.csv' ~ 'NCAA Championship'
    )
    mergedData[[i]]$Date[j] <- case_when(
      name == 'ACC_champ.csv' ~ '2024-03-23',
      name == 'Away_meet_NCSU.csv' ~ '2024-02-17',
      name == 'Away_meet_PITT.csv' ~ '2024-02-09',
      name == 'Away_meet_QUAD.csv' ~ '2024-03-03',
      name == 'Away_meet_UCLA.csv' ~ '2024-03-16',
      name == 'Away_meet_UNC.csv' ~ '2024-01-26',
      name == 'Home_meet_AF.csv' ~ '2024-03-08',
      name == 'Home_meet_NCSU.csv' ~ '2024-02-02',
      name == 'Home_meet_PITT.csv' ~ '2024-01-21',
      name == 'Home_meet_UNC.csv' ~ '2024-02-23',
      name == 'NCAA_champ.csv' ~ '2024-04-04'     
    )
  }
}


# Add in placing in each event

for(i in 1:length(meetFiles)){
  mergedData[[i]] <- mergedData[[i]] |>
    group_by(Event) |>
    mutate(Place = rank(-1*as.numeric(Score), ties.method = 'min'))
}


# Add in team placing in each event

for(i in 1:length(meetFiles)){
  mergedData[[i]] <- mergedData[[i]] |>
    group_by(Event, Team) |>
    mutate(PlaceTeam = rank(-1*as.numeric(Score), ties.method = 'first'))
}

# Final row binding into one data frame


finalProduct <- bind_rows(mergedData)
```



```{r}
# Some team names are represented under two different values:

# NC State and North Carolina State
# CU and Clemson

## Fix team names:

for(i in 1:dim(finalProduct)[1]){
  if(finalProduct$Team[i] == 'CU'){
    finalProduct$Team[i] <- 'Clemson'
  }
  else if(finalProduct$Team[i] == 'North Carolina State'){
    finalProduct$Team[i] <- 'NC State'
  }
}
```


```{r}
# Score corrections for individual extra competitors:

## vs. Air Force: Maggie Holman's 9.9 on Floor is an extra score and will not count; adjust rank

finalProduct <- finalProduct |>
  mutate(PlaceTeam = case_when(
    Match == 'vs Air Force' & Event == 'FLOOR' & Name == 'Maggie Holman' ~ 6, #To account for her score not counting
    Match == 'vs Air Force' & Event == 'FLOOR' & Name == 'Rebecca Wells' ~ 5, #Real fifth score
    .default = PlaceTeam)
  )

## at Pitt: Kennedy Duke's 9.725 on Beam is an extra score and will not count; adjust rank

finalProduct <- finalProduct |>
  mutate(PlaceTeam = case_when(
    Match == 'at Pitt' & Event == 'BEAM' & Name == 'Kennedy Duke' ~ 6, #To account for her score not counting
    Match == 'at Pitt' & Event == 'BEAM' & Name == 'Julie Madso' ~ 5, #Real fifth score
    .default = PlaceTeam)
  )

## vs UNC: Several extra gymnasts scored above scoring gymnasts for UNC; adjust ranks

finalProduct <- finalProduct |>
  mutate(PlaceTeam = case_when(
    Match == 'vs UNC' & Event == 'VAULT' & Name == 'Elly King' ~ 6,
    Match == 'vs UNC' & Event == 'VAULT' & Name == 'Paige Prejean' ~ 5, 
    Match == 'vs UNC' & Event == 'BARS' & Name == 'Amy Wozniak' ~ 6,
    Match == 'vs UNC' & Event == 'BARS' & Name == 'Jamie Shearer' ~ 5, 
    Match == 'vs UNC' & Event == 'BEAM' & Name == 'Jamie Shearer' ~ 6, 
    Match == 'vs UNC' & Event == 'BEAM' & Name == 'Elly King' ~ 4, #Tie for 4th/5th place with another correctly tagged gymnast
    .default = PlaceTeam)
  )


## In the quad meet: Oklahoma's Faith Torrez is incorrectly scored as a 9.9 instead of her actual 9.95 on vault. Corrections required:

# Fix Faith's score
# Make her ranked 1st overall and 1st on Oklahoma
# Hannah Scheible and Jordan Bowers should be ranked 2nd overall and 2nd on Oklahoma
# Madison Minner should be ranked 4th overall

finalProduct <- finalProduct |>
  mutate(
    Score = case_when(
      Match == 'at TWU' & Event == 'VAULT' & Name == 'Faith Torrez' ~ '9.95',
      .default = Score
    ),
    Place = case_when(
      Match == 'at TWU' & Event == 'VAULT' & Name == 'Faith Torrez' ~ 1,
      Match == 'at TWU' & Event == 'VAULT' & Name == 'Hannah Scheible' ~ 2,
      Match == 'at TWU' & Event == 'VAULT' & Name == 'Jordan Bowers' ~ 2,
      Match == 'at TWU' & Event == 'VAULT' & Name == 'Madison Minner' ~ 4,
      .default = Place
    ),
    PlaceTeam = case_when(
      Match == 'at TWU' & Event == 'VAULT' & Name == 'Faith Torrez' ~ 1,
      Match == 'at TWU' & Event == 'VAULT' & Name == 'Hannah Scheible' ~ 2,
      Match == 'at TWU' & Event == 'VAULT' & Name == 'Jordan Bowers' ~ 2,
      .default = PlaceTeam
    )
  )

```




```{r}
# Send to an Excel file to place into Tableau
## Excel used instead of CSV to not cause any errors in above code
write_xlsx(finalProduct, 'Clemson_2024_Gymnastics.xlsx')
```


