---
title: "Edge% and Heart% Calculator"
author: "Alexander Harriman"
date: '2022-07-18'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This code takes a Trackman file with the Ball/Edge/Heart column defined and calculates the Edge% and Heart% for every pitcher within the file.

Those columns are exported into their own csv file at the end.

# Load Data
```{r}
pitchers2022 <- read.csv("2022 NCAA Trackman with Command.csv")

pitchers2022 <- pitchers2022[is.na(pitchers2022$ZoneType) == FALSE,]
pitchers2022 <- pitchers2022[is.na(pitchers2022$Pitcher) == FALSE,]
```

# Filter Data for Pitch Requirements

## Fastball Filtering

```{r}
# Filter by Fastball

ncaa2022 <- pitchers2022[pitchers2022$TaggedPitchType == "Fastball" | pitchers2022$TaggedPitchType == "Sinker",]

# Filter by Minimum Speed
ncaa2022 <- ncaa2022[ncaa2022$RelSpeed > 85.999,]

# Separating lefties and righties

ncaaRight <- ncaa2022[ncaa2022$PitcherThrows == "Right",]
ncaaLeft <- ncaa2022[ncaa2022$PitcherThrows == "Left",]

ncaaRight$Tilt.Coordinate
# Using tilt to separate four-seamers and two-seamers

rightTilt <- function(){
  for(i in 1:dim(ncaaRight)[1]){
    if(is.na(ncaaRight$Tilt.Coordinate[i]) == TRUE){
      ncaaRight$`Pitch Assignment`[i] <<- "None"
    }
    else if(ncaaRight$Tilt.Coordinate[i] > 0.07 && ncaaRight$Tilt.Coordinate[i] < .135){
      ncaaRight$`Pitch Assignment`[i] <<- "Right Two-Seam"
    }
    else if(ncaaRight$Tilt.Coordinate[i] > 0.04  && ncaaRight$Tilt.Coordinate[i] < .065){
      ncaaRight$`Pitch Assignment`[i] <<- "Right Four-Seam"
    }
    else if(ncaaRight$Tilt.Coordinate[i] > 0.499 && ncaaRight$Tilt.Coordinate[i] < .532){
      ncaaRight$`Pitch Assignment`[i] <<- "Right Four-Seam"
    }
  }
  return(ncaaRight$`Pitch Assignment`)
}

rightTilt()

ncaaRightFour <- ncaaRight[ncaaRight$`Pitch Assignment` == "Right Four-Seam",]
ncaaRightTwo <- ncaaRight[ncaaRight$`Pitch Assignment` == "Right Two-Seam",]


#Filter Out Bad Pitchers
ncaaRightTwo <- ncaaRightTwo[ncaaRightTwo$Pitcher != "Maxwell, Zachary",]

ncaaRightTwo <- ncaaRightTwo[ncaaRightTwo$Pitcher != "Stohr, Ryan",]
ncaaRightTwo <- ncaaRightTwo[is.na(ncaaRightTwo$Pitcher) == FALSE,]

ncaaRightFour <- ncaaRightFour[is.na(ncaaRightFour$Pitcher) == FALSE,]

#Zachary Maxwell's two-seamers break the code due to their unorthodox measurables. These are almost certainly incorrectly marked in Trackman.

#Ryan Stohr's fastballs have missing values in them, particularly for IVB.

#Both types of fastballs contain NA values, which break the code.

dim(ncaaRightFour)
dim(ncaaRightTwo)

leftTilt <- function(){
  for(i in 1:dim(ncaaLeft)[1]){
    if(is.na(ncaaLeft$Tilt.Coordinate[i]) == TRUE){
      ncaaLeft$`Pitch Assignment`[i] <<- "None"
    }
    else if(ncaaLeft$Tilt.Coordinate[i] > .3749 && ncaaLeft$Tilt.Coordinate[i] < .428){
      ncaaLeft$`Pitch Assignment`[i] <<- "Left Two-Seam"
    }
    else if(ncaaLeft$Tilt.Coordinate[i] >  .437 && ncaaLeft$Tilt.Coordinate[i] < 0.501){
      ncaaLeft$`Pitch Assignment`[i] <<- "Left Four-Seam"
    }
  }
  return(ncaaRight$`Pitch Assignment`)
}

leftTilt()


ncaaLeftFour <- ncaaLeft[ncaaLeft$`Pitch Assignment` == "Left Four-Seam",]
ncaaLeftTwo <- ncaaLeft[ncaaLeft$`Pitch Assignment` == "Left Two-Seam",]
dim(ncaaLeftFour)
dim(ncaaLeftTwo)
```

```{r}
# Set Pitch Minimum at 50 Pitches:

##This code takes 6 runs
###Make sure not to continue until the dimensions line up after running

#### Then continue

pitcherClearLF <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    x <- ncaaLeftFour
    base <- unique(x$Pitcher)[i]
    if(length(x[x$Pitcher == base,]$PitchNo) < 50){
      ncaaLeftFour <<- subset(x, x$Pitcher != base)
    }
  }
  return(ncaaLeftFour)
}

pitcherClearLT <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    x <- ncaaLeftTwo
    base <- unique(x$Pitcher)[i]
    if(length(x[x$Pitcher == base,]$PitchNo) < 50){
      ncaaLeftTwo <<- subset(x, x$Pitcher != base)
    }
  }
  return(ncaaLeftTwo)
}

pitcherClearRF <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    x <- ncaaRightFour
    base <- unique(x$Pitcher)[i]
    if(length(x[x$Pitcher == base,]$PitchNo) < 50){
      ncaaRightFour <<- subset(x, x$Pitcher != base)
    }
  }
  return(ncaaRightFour)
}

pitcherClearRT <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    x <- ncaaRightTwo
    base <- unique(x$Pitcher)[i]
    if(length(x[x$Pitcher == base,]$PitchNo) < 50){
      ncaaRightTwo <<- subset(x, x$Pitcher != base)
    }
  }
  return(ncaaRightTwo)
}

dim(ncaaLeftFour)
pitcherClearLF(ncaaLeftFour)
dim(ncaaLeftFour)

dim(ncaaRightTwo)
pitcherClearRT(ncaaRightTwo)
dim(ncaaRightTwo)

dim(ncaaRightFour)
pitcherClearRF(ncaaRightFour)
dim(ncaaRightFour)

dim(ncaaLeftTwo)
pitcherClearLT(ncaaLeftTwo)
dim(ncaaLeftTwo)
```
## Slider Filtering

```{r}
# Separating lefties and righties

ncaa2022 <- pitchers2022[pitchers2022$TaggedPitchType == "Slider",]

ncaaRight <- ncaa2022[ncaa2022$PitcherThrows == "Right",]
ncaaLeft <- ncaa2022[ncaa2022$PitcherThrows == "Left",]

ncaaRightSlider = subset(ncaaRight, ncaaRight$Tilt.Coordinate >= .333 & 
                        ncaaRight$Tilt.Coordinate <= .4271)
ncaaLeftSlider = subset(ncaaLeft, ncaaLeft$Tilt.Coordinate >= .0729 &
                       ncaaLeft$Tilt.Coordinate <= .1667)

dim(ncaaRightSlider)
```

```{r}
# Set Pitch Minimum at 20 Pitches:

##This code takes 6 runs
###Make sure not to continue until the dimensions line up after running

#### Then continue

pitcherClearLSlider <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    x <- ncaaLeftSlider
    base <- unique(x$Pitcher)[i]
    if(length(x[x$Pitcher == base,]$PitchNo) < 50){
      ncaaLeftSlider <<- subset(x, x$Pitcher != base)
    }
  }
  return(ncaaLeftSlider)
}

pitcherClearRSlider <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    x <- ncaaRightSlider
    base <- unique(x$Pitcher)[i]
    if(length(x[x$Pitcher == base,]$PitchNo) < 50){
      ncaaRightSlider <<- subset(x, x$Pitcher != base)
    }
  }
  return(ncaaRightSlider)
}

dim(ncaaLeftSlider)
pitcherClearLSlider(ncaaLeftSlider)
dim(ncaaLeftSlider)

dim(ncaaRightSlider)
pitcherClearRSlider(ncaaRightSlider)
dim(ncaaRightSlider)
```

## Changeup Filtering
```{r}
ncaaChangeup2022 <- pitchers2022[pitchers2022$TaggedPitchType == "ChangeUp",]
```

```{r}
# Separating lefties and righties

ncaaRightChangeup <- ncaaChangeup2022[ncaaChangeup2022$PitcherThrows == "Right",]
ncaaLeftChangeup <- ncaaChangeup2022[ncaaChangeup2022$PitcherThrows == "Left",]
```

```{r}
# Set Pitch Minimum at 20 Pitches:

##This code takes 6 runs
###Make sure not to continue until the dimensions line up after running

#### Then continue

pitcherClearLChangeup <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    x <- ncaaLeftChangeup
    base <- unique(x$Pitcher)[i]
    if(length(x[x$Pitcher == base,]$PitchNo) < 50){
      ncaaLeftChangeup <<- subset(x, x$Pitcher != base)
    }
  }
  return(ncaaLeftChangeup)
}

pitcherClearRChangeup <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    x <- ncaaRightChangeup
    base <- unique(x$Pitcher)[i]
    if(length(x[x$Pitcher == base,]$PitchNo) < 50){
      ncaaRightChangeup <<- subset(x, x$Pitcher != base)
    }
  }
  return(ncaaRightChangeup)
}

dim(ncaaLeftChangeup)
pitcherClearLChangeup(ncaaLeftChangeup)
dim(ncaaLeftChangeup)

dim(ncaaRightChangeup)
pitcherClearRChangeup(ncaaRightChangeup)
dim(ncaaRightChangeup)
```

# Curveball Filtering

```{r}
# Separating lefties and righties

ncaaRight <- pitchers2022[pitchers2022$PitcherThrows == "Right",]
ncaaLeft <- pitchers2022[pitchers2022$PitcherThrows == "Left",]

ncaaRightCurve = subset(ncaaRight, ncaaRight$TaggedPitchType == "Curveball")
ncaaLeftCurve = subset(ncaaLeft, ncaaLeft$TaggedPitchType == "Curveball")
```

```{r}
# Set Pitch Minimum at 20 Pitches by filtering out the rest:

##This code takes 6 runs
###Make sure not to continue until the dimensions line up after running

#### Then continue

pitcherClearLCurve <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    x <- ncaaLeftCurve
    base <- unique(x$Pitcher)[i]
    if(length(x[x$Pitcher == base,]$PitchNo) < 50){
      ncaaLeftCurve <<- subset(x, x$Pitcher != base)
    }
  }
  return(ncaaLeftCurve)
}

pitcherClearRCurve <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    x <- ncaaRightCurve
    base <- unique(x$Pitcher)[i]
    if(length(x[x$Pitcher == base,]$PitchNo) < 50){
      ncaaRightCurve <<- subset(x, x$Pitcher != base)
    }
  }
  return(ncaaRightCurve)
}

dim(ncaaLeftCurve)
pitcherClearLCurve(ncaaLeftCurve)
dim(ncaaLeftCurve)

dim(ncaaRightCurve)
pitcherClearRCurve(ncaaRightCurve)
dim(ncaaRightCurve)
```


# Find Edge, Heart, and Ball Counts
```{r}
edge <- vector()
heart <- vector()
ball <- vector()
edges <- 0
hearts <- 0
balls <- 0


heartCounter <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    base <- unique(x$Pitcher)[i]
    pitcher <- x[x$Pitcher == base,]
    for(a in 1:length(pitcher$PitchNo)){
      if(pitcher$ZoneType[a] == "Heart"){
        hearts <- hearts + 1
      }
    }
    heart[i] <- hearts
    hearts <- 0
  }
  return(heart)
}

edgeCounter <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    base <- unique(x$Pitcher)[i]
    pitcher <- x[x$Pitcher == base,]
    for(a in 1:length(pitcher$PitchNo)){
      if(pitcher$ZoneType[a] == "Edge"){
        edges <- edges + 1
      }
    }
    edge[i] <- edges
    edges <- 0
  }
  return(edge)
}

ballCounter <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    base <- unique(x$Pitcher)[i]
    pitcher <- x[x$Pitcher == base,]
    for(a in 1:length(pitcher$PitchNo)){
      if(pitcher$ZoneType[a] == "Ball"){
        balls <- balls + 1
      }
    }
    ball[i] <- balls
    balls <- 0
  }
  return(ball)
}

ELF <- edgeCounter(ncaaLeftFour)
ELT <- edgeCounter(ncaaLeftTwo)
ERF <- edgeCounter(ncaaRightFour)
ERT <- edgeCounter(ncaaRightTwo)
ELS <- edgeCounter(ncaaLeftSlider)
ERS <- edgeCounter(ncaaRightSlider)
ELU <- edgeCounter(ncaaLeftChangeup)
ERU <- edgeCounter(ncaaRightChangeup)
ELC <- edgeCounter(ncaaLeftCurve)
ERC <- edgeCounter(ncaaRightCurve)

HLF <- heartCounter(ncaaLeftFour)
HLT <- heartCounter(ncaaLeftTwo)
HRF <- heartCounter(ncaaRightFour)
HRT <- heartCounter(ncaaRightTwo)
HLS <- heartCounter(ncaaLeftSlider)
HRS <- heartCounter(ncaaRightSlider)
HLU <- heartCounter(ncaaLeftChangeup)
HRU <- heartCounter(ncaaRightChangeup)
HLC <- heartCounter(ncaaLeftCurve)
HRC <- heartCounter(ncaaRightCurve)

BLF <- ballCounter(ncaaLeftFour)
BLT <- ballCounter(ncaaLeftTwo)
BRF <- ballCounter(ncaaRightFour)
BRT <- ballCounter(ncaaRightTwo)
BLS <- ballCounter(ncaaLeftSlider)
BRS <- ballCounter(ncaaRightSlider)
BLU <- ballCounter(ncaaLeftChangeup)
BRU <- ballCounter(ncaaRightChangeup)
BLC <- ballCounter(ncaaLeftCurve)
BRC <- ballCounter(ncaaRightCurve)
```

# Find Pitch Count
```{r}
pitch <- vector()
pitchCounter <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    base <- unique(x$Pitcher)[i]
    pitcher <- x[x$Pitcher == base,]
    pitch[i] <- dim(pitcher)[1]
  }
  return(pitch)
}

PLF <- pitchCounter(ncaaLeftFour)
PLT <- pitchCounter(ncaaLeftTwo)
PRF <- pitchCounter(ncaaRightFour)
PRT <- pitchCounter(ncaaRightTwo)
PLS <- pitchCounter(ncaaLeftSlider)
PRS <- pitchCounter(ncaaRightSlider)
PLU <- pitchCounter(ncaaLeftChangeup)
PRU <- pitchCounter(ncaaRightChangeup)
PLC <- pitchCounter(ncaaLeftCurve)
PRC <- pitchCounter(ncaaRightCurve)
```

# Find Edge% and Heart%

```{r}
percentage <- vector()

percentCounter <- function(x,y){
  for(i in 1:length(x)){
    percentage[i] <- 100*x[i] / y[i]
  }
  return(percentage)
}

leftFourEdge <- percentCounter(ELF, PLF)
leftTwoEdge <- percentCounter(ELT, PLT)
rightFourEdge <- percentCounter(ERF, PRF)
rightTwoEdge <- percentCounter(ERT, PRT)
leftSliderEdge <- percentCounter(ELS, PLS)
rightSliderEdge <- percentCounter(ERS, PRS)
leftChangeupEdge <- percentCounter(ELU, PLU)
rightChangeupEdge <- percentCounter(ERU, PRU)
leftCurveEdge <- percentCounter(ELC, PLC)
rightCurveEdge <- percentCounter(ERC, PRC)

leftFourHeart <- percentCounter(HLF, PLF)
leftTwoHeart <- percentCounter(HLT, PLT)
rightFourHeart <- percentCounter(HRF, PRF)
rightTwoHeart <- percentCounter(HRT, PRT)
leftSliderHeart <- percentCounter(HLS, PLS)
rightSliderHeart <- percentCounter(HRS, PRS)
leftChangeupHeart <- percentCounter(HLU, PLU)
rightChangeupHeart <- percentCounter(HRU, PRU)
leftCurveHeart <- percentCounter(HLC, PLC)
rightCurveHeart <- percentCounter(HRC, PRC)

leftFourBall <- percentCounter(BLF, PLF)
leftTwoBall <- percentCounter(BLT, PLT)
rightFourBall <- percentCounter(BRF, PRF)
rightTwoBall <- percentCounter(BRT, PRT)
leftSliderBall <- percentCounter(BLS, PLS)
rightSliderBall <- percentCounter(BRS, PRS)
leftChangeupBall <- percentCounter(BLU, PLU)
rightChangeupBall <- percentCounter(BRU, PRU)
leftCurveBall <- percentCounter(BLC, PLC)
rightCurveBall <- percentCounter(BRC, PRC)
```

# Find Pitcher Team
```{r}
# Find the teams for each pitcher
team <- vector()
teamFinder <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    base <- unique(x$Pitcher)[i]
    pitcher <- x[x$Pitcher == base,]
    team[i] <- unique(pitcher$PitcherTeam)[1]
  }
  return(team)
}

leftFourTeam <- teamFinder(ncaaLeftFour)
rightFourTeam <- teamFinder(ncaaRightFour)
leftTwoTeam <- teamFinder(ncaaLeftTwo)
rightTwoTeam <- teamFinder(ncaaRightTwo)
leftSliderTeam <- teamFinder(ncaaLeftSlider)
rightSliderTeam <- teamFinder(ncaaRightSlider)
leftChangeupTeam <- teamFinder(ncaaLeftChangeup)
rightChangeupTeam <- teamFinder(ncaaRightChangeup)
leftCurveTeam <- teamFinder(ncaaLeftCurve)
rightCurveTeam <- teamFinder(ncaaRightCurve)
```

# Export Files

```{r}
library(dplyr)
leftFourTable <- mapply(c, unique(ncaaLeftFour$Pitcher), leftFourTeam, ELF, HLF, BLF, PLF, leftFourEdge, leftFourHeart, leftFourBall)

leftTwoTable <- mapply(c, unique(ncaaLeftTwo$Pitcher), leftTwoTeam, ELT, HLT, BLT, PLT, leftTwoEdge, leftTwoHeart, leftTwoBall)

rightFourTable <- mapply(c, unique(ncaaRightFour$Pitcher), rightFourTeam, ERF, HRF, BRF, PRF, rightFourEdge, rightFourHeart, rightFourBall)

rightTwoTable <- mapply(c, unique(ncaaRightTwo$Pitcher), rightTwoTeam, ERT, HRT, BRT, PRT, rightTwoEdge, rightTwoHeart, rightTwoBall)

leftSliderTable <- mapply(c, unique(ncaaLeftSlider$Pitcher), leftSliderTeam, ELS, HLS, BLS, PLS, leftSliderEdge, leftSliderHeart, leftSliderBall)

rightSliderTable <- mapply(c, unique(ncaaRightSlider$Pitcher), rightSliderTeam, ERS, HRS, BRS, PRS, rightSliderEdge, rightSliderHeart, rightSliderBall)

leftChangeupTable <- mapply(c, unique(ncaaLeftChangeup$Pitcher), leftChangeupTeam, ELU, HLU, BLU, PLU, leftChangeupEdge, leftChangeupHeart, leftChangeupBall)

rightChangeupTable <- mapply(c, unique(ncaaRightChangeup$Pitcher), rightChangeupTeam, ERU, HRU, BRU, PRU, rightChangeupEdge, rightChangeupHeart, rightChangeupBall)

leftCurveTable <- mapply(c, unique(ncaaLeftCurve$Pitcher), leftCurveTeam, ELC, HLC, BLC, PLC, leftCurveEdge, leftCurveHeart, leftCurveBall)

rightCurveTable <- mapply(c, unique(ncaaRightCurve$Pitcher), rightCurveTeam, ERC, HRC, BRC, PRC, rightCurveEdge, rightCurveHeart, rightCurveBall)
```

# Export Files
```{r}
write.csv(leftFourTable, "2022 NCAA Left Four Edge Heart Ball%.csv", row.names = FALSE)

write.csv(leftTwoTable, "2022 NCAA Left Two Edge Heart Ball%.csv", row.names = FALSE)

write.csv(rightFourTable, "2022 NCAA Right Four Edge Heart Ball%.csv", row.names = FALSE)

write.csv(rightTwoTable, "2022 NCAA Right Two Edge Heart Ball%.csv", row.names = FALSE)

write.csv(leftSliderTable, "2022 NCAA Left Slider Edge Heart Ball%.csv", row.names = FALSE)

write.csv(rightSliderTable, "2022 NCAA Right Slider Edge Heart Ball%.csv", row.names = FALSE)

write.csv(leftChangeupTable, "2022 NCAA Left Changeup Edge Heart Ball%.csv", row.names = FALSE)

write.csv(rightChangeupTable, "2022 NCAA Right Changeup Edge Heart Ball%.csv", row.names = FALSE)

write.csv(leftCurveTable, "2022 NCAA Left Curve Edge Heart Ball%.csv", row.names = FALSE)

write.csv(rightCurveTable, "2022 NCAA Right Curve Edge Heart Ball%.csv", row.names = FALSE)
```
