---
title: "NCAA Pitcher Statistic Calculation 2022"
author: "Alexander Harriman"
date: "5/9/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file takes all of the NCAA pitchers from games in the 2022 season **May 3rd Last Update** and calculates their percentiles needed for Stuff+, their Barrel% and their Whiff%. All of these will come from a combined Trackman file, compiled in Excel first.

This will be used for validation purposes and for calculating the Stuff+ of all NCAA pitchers.



# Insert ACC Pitcher Files

```{r, warning = FALSE}
library(readxl)
library(plyr)
pitchers2022 <- read_excel("Reduced 2022 Trackman (5-4).xlsx")
newPitchers2022 <- read_excel("Combined 2022 Trackman (5-5 to 5-20).xlsx")
total2022 <- rbind.fill(pitchers2022,newPitchers2022)

acc2022 <- total2022[total2022$TaggedPitchType == "Fastball",]

# Filter by Minimum Speed
acc2022 <- acc2022[acc2022$RelSpeed > 85.999,]


names <- unique(acc2022$Pitcher)

names

head(acc2022)

#There are 4293 pitchers in this dataset.

##We will filter them down by fastball type
```

# Fastball Filtering

```{r}
# Separating lefties and righties

accRight <- acc2022[acc2022$PitcherThrows == "Right",]
accLeft <- acc2022[acc2022$PitcherThrows == "Left",]

dim(accRight)


# Using tilt to separate four-seamers and two-seamers
rightTilt <- function(){
  for(i in 1:dim(accRight)[1]){
    if(is.na(accRight$`Tilt Coordinate`[i]) == TRUE){
      accRight$`Pitch Assignment`[i] <<- "None"
    }
    else if(accRight$`Tilt Coordinate`[i] > 0.07 && accRight$`Tilt Coordinate`[i] < .135){
      accRight$`Pitch Assignment`[i] <<- "Right Two-Seam"
    }
    else if(accRight$`Tilt Coordinate`[i] > 0.04  && accRight$`Tilt Coordinate`[i] < .065){
      accRight$`Pitch Assignment`[i] <<- "Right Four-Seam"
    }
    else if(accRight$`Tilt Coordinate`[i] > 0.499 && accRight$`Tilt Coordinate`[i] < .532){
      accRight$`Pitch Assignment`[i] <<- "Right Four-Seam"
    }
  }
  return(accRight$`Pitch Assignment`)
}

rightTilt()

leftTilt <- function(){
  for(i in 1:dim(accLeft)[1]){
    if(is.na(accLeft$`Tilt Coordinate`[i]) == TRUE){
      accLeft$`Pitch Assignment`[i] <<- "None"
    }
    else if(accLeft$`Tilt Coordinate`[i] > .3749 && accLeft$`Tilt Coordinate`[i] < .428){
      accLeft$`Pitch Assignment`[i] <<- "Left Two-Seam"
    }
    else if(accLeft$`Tilt Coordinate`[i] >  .437 && accLeft$`Tilt Coordinate`[i] < 0.501){
      accLeft$`Pitch Assignment`[i] <<- "Left Four-Seam"
    }
  }
  return(accLeft$`Pitch Assignment`)
}

leftTilt()

```


```{r}
#Adjusting for Known Pitchers

##Not all pitches will be correct simply by tilt
###For Clemson pitchers and Ben Joyce, all of their pitches will be corrected to fit their pitching style

#Exceptions: Jay Dill and Mack Anglin (throw both types)
pitcherChangeR <- function(){
  for(i in 1:length(accRight$Pitcher)){
    if(is.na(accRight$Pitcher[i]) == TRUE){
      accRight$`Pitch Assignment`[i] <<- "None"
    }
    else if(accRight$Pitcher[i] == "Joyce, Ben"){
      accRight$`Pitch Assignment`[i] <<- "Right Four-Seam"
    }
    else if(accRight$Pitcher[i] == "Hoffmann, Nick"){
      accRight$`Pitch Assignment`[i] <<- "Right Four-Seam"
    }
    else if(accRight$Pitcher[i] == "Gordon, Austin"){
      accRight$`Pitch Assignment`[i] <<- "Right Four-Seam"
    }
    else if(accRight$Pitcher[i] == "Williams, Ricky"){
      accRight$`Pitch Assignment`[i] <<- "Right Four-Seam"
    }
    else if(accRight$Pitcher[i] == "Tallent, Casey"){
      accRight$`Pitch Assignment`[i] <<- "Right Four-Seam"
    }
    else if(accRight$Pitcher[i] == "Olenchuck, Ty"){
      accRight$`Pitch Assignment`[i] <<- "Right Four-Seam"
    }
    else if(accRight$Pitcher[i] == "Edmondson, Alex"){
      accRight$`Pitch Assignment`[i] <<- "Right Four-Seam"
    }
    else if(accRight$Pitcher[i] == "Gilbert, Geoffrey"){
      accRight$`Pitch Assignment`[i] <<- "Right Two-Seam"
    }
    else if(accRight$Pitcher[i] == "Ammons, Ryan"){
      accRight$`Pitch Assignment`[i] <<- "Right Two-Seam"
    }
    else if(accRight$Pitcher[i] == "Labriola, Paul"){
      accRight$`Pitch Assignment`[i] <<- "Right Two-Seam"
    }
  }
  return(accRight$`Pitch Assignment`)
}

pitcherChangeL <- function(){
  for(i in 1:length(accLeft$Pitcher)){
    if(is.na(accLeft$Pitcher[i]) == TRUE){
      accLeft$`Pitch Assignment`[i] <<- "None"
    }
    else if(accLeft$Pitcher[i] == "Grice, Caden"){
      accLeft$`Pitch Assignment`[i] <<- "Left Two-Seam"
    }
    else if(accLeft$Pitcher[i] == "Reid, Rocco"){
      accLeft$`Pitch Assignment`[i] <<- "Left Two-Seam"
    }
    else if(accLeft$Pitcher[i] == "Lindley, Jackson"){
      accLeft$`Pitch Assignment`[i] <<- "Left Four-Seam"
    }
    else if(accLeft$Pitcher[i] == "Clayton, Nick"){
      accLeft$`Pitch Assignment`[i] <<- "Left Four-Seam"
    }
    else if(accLeft$Pitcher[i] == "Barlow, Billy"){
      accLeft$`Pitch Assignment`[i] <<- "Left Four-Seam"
    }
  }
  return(accLeft$`Pitch Assignment`)
}

pitcherChangeR()
pitcherChangeL()

```

# Fastball Filtering


```{r}

accRightFour <- accRight[accRight$`Pitch Assignment` == "Right Four-Seam",]
accRightTwo <- accRight[accRight$`Pitch Assignment` == "Right Two-Seam",]

#Filter out bad pitchers
accRightTwo <- accRightTwo[accRightTwo$Pitcher != "Maxwell, Zachary",]

accRightFour <- accRightFour[is.na(accRightFour$Pitcher) == FALSE,]

#Zachary Maxwell's two-seamers break the code due to their unorthodox measurables. As this is the only instance, we will assume these are incorrectly marked as two-seamers based on their Tilt, and omit him from this subset.
##We have an NA in the four-seam dataset when ran normally, which breaks the code.

dim(accRightFour)
dim(accRightTwo)

accLeftFour <- accLeft[accLeft$`Pitch Assignment` == "Left Four-Seam",]
accLeftTwo <- accLeft[accLeft$`Pitch Assignment` == "Left Two-Seam",]
dim(accLeftFour)
dim(accLeftTwo)


```


```{r}
# Set Pitch Minimum at 20 Pitches:

##This code takes 6 runs
###Make sure not to continue until the dimensions line up after running

#### Then continue

pitcherClearLF <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    x <- accLeftFour
    base <- unique(x$Pitcher)[i]
    if(length(x[x$Pitcher == base,]$PitchNo) < 20){
      accLeftFour <<- subset(x, x$Pitcher != base)
    }
  }
  return(accLeftFour)
}

pitcherClearLT <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    x <- accLeftTwo
    base <- unique(x$Pitcher)[i]
    if(length(x[x$Pitcher == base,]$PitchNo) < 20){
      accLeftTwo <<- subset(x, x$Pitcher != base)
    }
  }
  return(accLeftTwo)
}

pitcherClearRF <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    x <- accRightFour
    base <- unique(x$Pitcher)[i]
    if(length(x[x$Pitcher == base,]$PitchNo) < 20){
      accRightFour <<- subset(x, x$Pitcher != base)
    }
  }
  return(accRightFour)
}

pitcherClearRT <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    x <- accRightTwo
    base <- unique(x$Pitcher)[i]
    if(length(x[x$Pitcher == base,]$PitchNo) < 20){
      accRightTwo <<- subset(x, x$Pitcher != base)
    }
  }
  return(accRightTwo)
}

dim(accLeftFour)
pitcherClearLF(accLeftFour)
dim(accLeftFour)

dim(accRightTwo)
pitcherClearRT(accRightTwo)
dim(accRightTwo)

dim(accRightFour)
pitcherClearRF(accRightFour)
dim(accRightFour)

dim(accLeftTwo)
pitcherClearLT(accLeftTwo)
dim(accLeftTwo)

```


# Find Averages and Standard Deviations for Excel


## Find Mean and Standard Deviation for Metrics

```{r}
rightFourVeloMean <- mean(accRightFour$RelSpeed)
rightTwoVeloMean <- mean(accRightTwo$RelSpeed)
leftFourVeloMean <- mean(accLeftFour$RelSpeed)
leftTwoVeloMean <- mean(accLeftTwo$RelSpeed)

rightFourVeloSD <- sd(accRightFour$RelSpeed)
rightTwoVeloSD <- sd(accRightTwo$RelSpeed)
leftFourVeloSD <- sd(accLeftFour$RelSpeed)
leftTwoVeloSD <- sd(accLeftTwo$RelSpeed)

rightFourHorzMean <- abs(mean(accRightFour$HorzBreak))
rightTwoHorzMean <- abs(mean(accRightTwo$HorzBreak))
leftFourHorzMean <- abs(mean(accLeftFour$HorzBreak))
leftTwoHorzMean <- abs(mean(accLeftTwo$HorzBreak))

rightFourHorzSD <- abs(sd(accRightFour$HorzBreak))
rightTwoHorzSD <- abs(sd(accRightTwo$HorzBreak))
leftFourHorzSD <- abs(sd(accLeftFour$HorzBreak))
leftTwoHorzSD <- abs(sd(accLeftTwo$HorzBreak))

rightFourIVBMean <- mean(accRightFour$InducedVertBreak)
rightTwoIVBMean <- mean(accRightTwo$InducedVertBreak)
leftFourIVBMean <- mean(accLeftFour$InducedVertBreak)
leftTwoIVBMean <- mean(accLeftTwo$InducedVertBreak)

rightFourIVBSD <- sd(accRightFour$InducedVertBreak)
rightTwoIVBSD <- sd(accRightTwo$InducedVertBreak)
leftFourIVBSD <- sd(accLeftFour$InducedVertBreak)
leftTwoIVBSD <- sd(accLeftTwo$InducedVertBreak)

leftTwoIVBMean
leftTwoHorzMean
leftTwoVeloMean

leftTwoIVBSD
leftTwoHorzSD
leftTwoVeloSD

leftFourIVBMean
leftFourHorzMean
leftFourVeloMean

leftFourIVBSD
leftFourHorzSD
leftFourVeloSD

rightFourIVBMean
rightFourHorzMean
rightFourVeloMean

rightFourIVBSD
rightFourHorzSD
rightFourVeloSD

rightTwoIVBMean
rightTwoHorzMean
rightTwoVeloMean

rightTwoIVBSD
rightTwoHorzSD
rightTwoVeloSD
```

## Find Average Extension
```{r}
extension <- vector()
extensionFunction <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    base <- unique(x$Pitcher)[i]
    pitcher <- x[x$Pitcher == base,]
    exten <- pitcher[is.na(pitcher$Extension) == FALSE,]
    extension[i] <- mean(exten$Extension)
  }
  return(extension)
}

rightFourExtension <- extensionFunction(accRightFour)
rightTwoExtension <- extensionFunction(accRightTwo)
leftFourExtension <- extensionFunction(accLeftFour)
leftTwoExtension <- extensionFunction(accLeftTwo)



rightFourExtensionMean <- mean(rightFourExtension)
rightTwoExtensionMean <- mean(rightTwoExtension)
leftFourExtensionMean <- mean(leftFourExtension)
leftTwoExtensionMean <- mean(leftTwoExtension)

rightFourExtensionSD <- sd(rightFourExtension)
rightTwoExtensionSD <- sd(rightTwoExtension)
leftFourExtensionSD <- sd(leftFourExtension)
leftTwoExtensionSD <- sd(leftTwoExtension)

rightFourExtensionMean
rightTwoExtensionMean
leftFourExtensionMean
leftTwoExtensionMean

rightFourExtensionSD
rightTwoExtensionSD
leftFourExtensionSD
leftTwoExtensionSD
```

## Find Average Induced Vert by Bucket

```{r}
bucket1 <- vector()
bucket2 <- vector()
bucket3 <- vector()
bucket4 <- vector()
bucket5 <- vector()
bucket6 <- vector()
bucket7 <- vector()
bucket8 <- vector()
bucket9 <- vector()
bucket10 <- vector()
bucket11 <- vector()
bucket12 <- vector()

bucketMeanFunction <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    base <- unique(x$Pitcher)[i]
    pitcher <- x[x$Pitcher == base,]
    for(a in 1:length(pitcher$RelHeight)){
      if(pitcher$RelHeight[a] > 4.5 && pitcher$RelHeight[a] < 4.7){
      bucket1 <<- append(bucket1, pitcher$InducedVertBreak[a])
    }
    else if(pitcher$RelHeight[a] > 4.7 && pitcher$RelHeight[a] < 4.9){
      bucket2 <<- append(bucket2, pitcher$InducedVertBreak[a])
    }
    else if(pitcher$RelHeight[a] > 4.9 && pitcher$RelHeight[a] < 5.1){
      bucket3 <<- append(bucket3, pitcher$InducedVertBreak[a])
    }
    else if(pitcher$RelHeight[a] > 5.1 && pitcher$RelHeight[a] < 5.3){
      bucket4 <<- append(bucket4, pitcher$InducedVertBreak[a])
    }
    else if(pitcher$RelHeight[a] > 5.3 && pitcher$RelHeight[a] < 5.5){
      bucket5 <<- append(bucket5, pitcher$InducedVertBreak[a])
    }
    else if(pitcher$RelHeight[a] > 5.5 && pitcher$RelHeight[a] < 5.7){
      bucket6 <<- append(bucket6, pitcher$InducedVertBreak[a])
    }
    else if(pitcher$RelHeight[a] > 5.7 && pitcher$RelHeight[a] < 5.9){
      bucket7 <<- append(bucket7, pitcher$InducedVertBreak[a])
    }
    else if(pitcher$RelHeight[a] > 5.9 && pitcher$RelHeight[a] < 6.1){
      bucket8 <<- append(bucket8, pitcher$InducedVertBreak[a])
    }
    else if(pitcher$RelHeight[a] > 6.1 && pitcher$RelHeight[a] < 6.3){
      bucket9 <<- append(bucket9, pitcher$InducedVertBreak[a])
    }
    else if(pitcher$RelHeight[a] > 6.3 && pitcher$RelHeight[a] < 6.5){
      bucket10 <<- append(bucket10, pitcher$InducedVertBreak[a])
    }
    else if(pitcher$RelHeight > 6.5 && pitcher$RelHeight < 6.7){
      bucket11 <<- append(bucket11, pitcher$InducedVertBreak) 
    }
    else if(pitcher$RelHeight[a] > 6.7 && pitcher$RelHeight[a] < 6.9){
      bucket12 <<- append(bucket12, pitcher$InducedVertBreak[a])
    }
    else if(pitcher$RelHeight[a] < 4.5){
      bucket1 <<- append(bucket1, pitcher$InducedVertBreak[a])
    }
    else if(pitcher$RelHeight[a] > 6.9){
      bucket12 <<- append(bucket12, pitcher$InducedVertBreak[a])
    }
    }
  }
  return(c(mean(bucket1), mean(bucket2), mean(bucket3), mean(bucket4), mean(bucket5), mean(bucket6), mean(bucket7), mean(bucket8), mean(bucket9), mean(bucket10), mean(bucket11), mean(bucket12)))
}


rightFourBucketMean <- bucketMeanFunction(accRightFour)
leftFourBucketMean <- bucketMeanFunction(accLeftFour)

rightFourBucketMean
leftFourBucketMean

beep(3)
```

```{r}
bucketSDFunction <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    base <- unique(x$Pitcher)[i]
    pitcher <- x[x$Pitcher == base,]
    for(a in 1:length(pitcher$RelHeight)){
      if(pitcher$RelHeight[a] > 4.5 && pitcher$RelHeight[a] < 4.7){
      bucket1 <<- append(bucket1, pitcher$InducedVertBreak[a])
    }
    else if(pitcher$RelHeight[a] > 4.7 && pitcher$RelHeight[a] < 4.9){
      bucket2 <<- append(bucket2, pitcher$InducedVertBreak[a])
    }
    else if(pitcher$RelHeight[a] > 4.9 && pitcher$RelHeight[a] < 5.1){
      bucket3 <<- append(bucket3, pitcher$InducedVertBreak[a])
    }
    else if(pitcher$RelHeight[a] > 5.1 && pitcher$RelHeight[a] < 5.3){
      bucket4 <<- append(bucket4, pitcher$InducedVertBreak[a])
    }
    else if(pitcher$RelHeight[a] > 5.3 && pitcher$RelHeight[a] < 5.5){
      bucket5 <<- append(bucket5, pitcher$InducedVertBreak[a])
    }
    else if(pitcher$RelHeight[a] > 5.5 && pitcher$RelHeight[a] < 5.7){
      bucket6 <<- append(bucket6, pitcher$InducedVertBreak[a])
    }
    else if(pitcher$RelHeight[a] > 5.7 && pitcher$RelHeight[a] < 5.9){
      bucket7 <<- append(bucket7, pitcher$InducedVertBreak[a])
    }
    else if(pitcher$RelHeight[a] > 5.9 && pitcher$RelHeight[a] < 6.1){
      bucket8 <<- append(bucket8, pitcher$InducedVertBreak[a])
    }
    else if(pitcher$RelHeight[a] > 6.1 && pitcher$RelHeight[a] < 6.3){
      bucket9 <<- append(bucket9, pitcher$InducedVertBreak[a])
    }
    else if(pitcher$RelHeight[a] > 6.3 && pitcher$RelHeight[a] < 6.5){
      bucket10 <<- append(bucket10, pitcher$InducedVertBreak[a])
    }
    else if(pitcher$RelHeight > 6.5 && pitcher$RelHeight < 6.7){
      bucket11 <<- append(bucket11, pitcher$InducedVertBreak) 
    }
    else if(pitcher$RelHeight[a] > 6.7 && pitcher$RelHeight[a] < 6.9){
      bucket12 <<- append(bucket12, pitcher$InducedVertBreak[a])
    }
    else if(pitcher$RelHeight[a] < 4.5){
      bucket1 <<- append(bucket1, pitcher$InducedVertBreak[a])
    }
    else if(pitcher$RelHeight[a] > 6.9){
      bucket12 <<- append(bucket12, pitcher$InducedVertBreak[a])
    }
    }
  }
  return(c(sd(bucket1), sd(bucket2), sd(bucket3), sd(bucket4), sd(bucket5), sd(bucket6), sd(bucket7), sd(bucket8), sd(bucket9), sd(bucket10), sd(bucket11), sd(bucket12)))
}

rightFourBucketSD <- bucketSDFunction(accRightFour)
leftFourBucketSD <- bucketSDFunction(accLeftFour)

rightFourBucketSD
leftFourBucketSD
```
# Find Percentiles

```{r}
filterRightFour <- accRightFour
filterRightTwo <- accRightTwo
filterLeftFour <- accLeftFour
filterLeftTwo <- accLeftTwo
```

# Split buckets into RHP and LHP

```{r}

#Four-Seam Righties

bucket1RHPFF <- subset(filterRightFour, filterRightFour$RelHeight > 4.5 & filterRightFour$RelHeight < 4.7)
bucket2RHPFF <- subset(filterRightFour, filterRightFour$RelHeight > 4.7 & filterRightFour$RelHeight < 4.9)
bucket3RHPFF <- subset(filterRightFour, filterRightFour$RelHeight > 4.9 & filterRightFour$RelHeight < 5.1)
bucket4RHPFF <- subset(filterRightFour, filterRightFour$RelHeight > 5.1 & filterRightFour$RelHeight < 5.3)
bucket5RHPFF <- subset(filterRightFour, filterRightFour$RelHeight > 5.3 & filterRightFour$RelHeight < 5.5)
bucket6RHPFF <- subset(filterRightFour, filterRightFour$RelHeight > 5.5 & filterRightFour$RelHeight < 5.7)
bucket7RHPFF <- subset(filterRightFour, filterRightFour$RelHeight > 5.7 & filterRightFour$RelHeight < 5.9)
bucket8RHPFF <- subset(filterRightFour, filterRightFour$RelHeight > 5.9 & filterRightFour$RelHeight < 6.1)
bucket9RHPFF <- subset(filterRightFour, filterRightFour$RelHeight > 6.1 & filterRightFour$RelHeight < 6.3)
bucket10RHPFF <- subset(filterRightFour, filterRightFour$RelHeight > 6.3 & filterRightFour$RelHeight < 6.5)
bucket11RHPFF <- subset(filterRightFour, filterRightFour$RelHeight > 6.5 & filterRightFour$RelHeight < 6.7)
bucket12RHPFF <- subset(filterRightFour, filterRightFour$RelHeight > 6.7 & filterRightFour$RelHeight < 6.9)


#Four-Seam Lefties

bucket1LHPFF <- subset(filterLeftFour, filterLeftFour$RelHeight > 4.5 & filterLeftFour$RelHeight < 4.7)
bucket2LHPFF <- subset(filterLeftFour, filterLeftFour$RelHeight > 4.7 & filterLeftFour$RelHeight < 4.9)
bucket3LHPFF <- subset(filterLeftFour, filterLeftFour$RelHeight > 4.9 & filterLeftFour$RelHeight < 5.1)
bucket4LHPFF <- subset(filterLeftFour, filterLeftFour$RelHeight > 5.1 & filterLeftFour$RelHeight < 5.3)
bucket5LHPFF <- subset(filterLeftFour, filterLeftFour$RelHeight > 5.3 & filterLeftFour$RelHeight < 5.5)
bucket6LHPFF <- subset(filterLeftFour, filterLeftFour$RelHeight > 5.5 & filterLeftFour$RelHeight < 5.7)
bucket7LHPFF <- subset(filterLeftFour, filterLeftFour$RelHeight > 5.7 & filterLeftFour$RelHeight < 5.9)
bucket8LHPFF <- subset(filterLeftFour, filterLeftFour$RelHeight > 5.9 & filterLeftFour$RelHeight < 6.1)
bucket9LHPFF <- subset(filterLeftFour, filterLeftFour$RelHeight > 6.1 & filterLeftFour$RelHeight < 6.3)
bucket10LHPFF <- subset(filterLeftFour, filterLeftFour$RelHeight > 6.3 & filterLeftFour$RelHeight < 6.5)
bucket11LHPFF <- subset(filterLeftFour, filterLeftFour$RelHeight > 6.5 & filterLeftFour$RelHeight < 6.7)
bucket12LHPFF <- subset(filterLeftFour, filterLeftFour$RelHeight > 6.7 & filterLeftFour$RelHeight < 6.9)

#Two-Seam Lefties

bucket1LHPTF <- subset(filterLeftTwo, filterLeftTwo$RelHeight > 4.5 & filterLeftTwo$RelHeight < 4.7)
bucket2LHPTF <- subset(filterLeftTwo, filterLeftTwo$RelHeight > 4.7 & filterLeftTwo$RelHeight < 4.9)
bucket3LHPTF <- subset(filterLeftTwo, filterLeftTwo$RelHeight > 4.9 & filterLeftTwo$RelHeight < 5.1)
bucket4LHPTF <- subset(filterLeftTwo, filterLeftTwo$RelHeight > 5.1 & filterLeftTwo$RelHeight < 5.3)
bucket5LHPTF <- subset(filterLeftTwo, filterLeftTwo$RelHeight > 5.3 & filterLeftTwo$RelHeight < 5.5)
bucket6LHPTF <- subset(filterLeftTwo, filterLeftTwo$RelHeight > 5.5 & filterLeftTwo$RelHeight < 5.7)
bucket7LHPTF <- subset(filterLeftTwo, filterLeftTwo$RelHeight > 5.7 & filterLeftTwo$RelHeight < 5.9)
bucket8LHPTF <- subset(filterLeftTwo, filterLeftTwo$RelHeight > 5.9 & filterLeftTwo$RelHeight < 6.1)
bucket9LHPTF <- subset(filterLeftTwo, filterLeftTwo$RelHeight > 6.1 & filterLeftTwo$RelHeight < 6.3)
bucket10LHPTF <- subset(filterLeftTwo, filterLeftTwo$RelHeight > 6.3 & filterLeftTwo$RelHeight < 6.5)
bucket11LHPTF <- subset(filterLeftTwo, filterLeftTwo$RelHeight > 6.5 & filterLeftTwo$RelHeight < 6.7)
bucket12LHPTF <- subset(filterLeftTwo, filterLeftTwo$RelHeight > 6.7 & filterLeftTwo$RelHeight < 6.9)

#Two-Seam Righties

bucket1RHPTF <- subset(filterRightTwo, filterRightTwo$RelHeight > 4.5 & filterRightTwo$RelHeight < 4.7)
bucket2RHPTF <- subset(filterRightTwo, filterRightTwo$RelHeight > 4.7 & filterRightTwo$RelHeight < 4.9)
bucket3RHPTF <- subset(filterRightTwo, filterRightTwo$RelHeight > 4.9 & filterRightTwo$RelHeight < 5.1)
bucket4RHPTF <- subset(filterRightTwo, filterRightTwo$RelHeight > 5.1 & filterRightTwo$RelHeight < 5.3)
bucket5RHPTF <- subset(filterRightTwo, filterRightTwo$RelHeight > 5.3 & filterRightTwo$RelHeight < 5.5)
bucket6RHPTF <- subset(filterRightTwo, filterRightTwo$RelHeight > 5.5 & filterRightTwo$RelHeight < 5.7)
bucket7RHPTF <- subset(filterRightTwo, filterRightTwo$RelHeight > 5.7 & filterRightTwo$RelHeight < 5.9)
bucket8RHPTF <- subset(filterRightTwo, filterRightTwo$RelHeight > 5.9 & filterRightTwo$RelHeight < 6.1)
bucket9RHPTF <- subset(filterRightTwo, filterRightTwo$RelHeight > 6.1 & filterRightTwo$RelHeight < 6.3)
bucket10RHPTF <- subset(filterRightTwo, filterRightTwo$RelHeight > 6.3 & filterRightTwo$RelHeight < 6.5)
bucket11RHPTF <- subset(filterRightTwo, filterRightTwo$RelHeight > 6.5 & filterRightTwo$RelHeight < 6.7)
bucket12RHPTF <- subset(filterRightTwo, filterRightTwo$RelHeight > 6.7 & filterRightTwo$RelHeight < 6.9)
```

# Creating Vectors for Percentiles
```{r}
# Velocity
percentileFunction <- function(x){
  percentileVector <- quantile(na.omit(x$RelSpeed), probs = seq(0, 1, by = 0.01))
  return(percentileVector)
}

# Induced Vertical Break
percentileFunction2 <- function(x){
  percentileVector <- quantile(na.omit(abs(x$InducedVertBreak)), probs = seq(0, 1, by = 0.01))
  return(percentileVector)
}


# Horizontal Break
percentileFunction3 <- function(x){
  percentileVector <- quantile(na.omit(abs(x$HorzBreak)), probs = seq(0, 1, by = 0.01))
  return(percentileVector)
}

# Extension
percentileFunction4 <- function(x){
  percentileVector <- quantile(na.omit(x$Extension), probs = seq(0, 1, by = 0.01))
  return(percentileVector)
}


```

# Combining the Vectors

```{r}
library(dplyr)

# Velocity
totalVeloTable <- mapply(c, percentileFunction(bucket1RHPFF), percentileFunction(bucket2RHPFF), percentileFunction(bucket3RHPFF), percentileFunction(bucket4RHPFF), percentileFunction(bucket5RHPFF), percentileFunction(bucket6RHPFF), percentileFunction(bucket7RHPFF), percentileFunction(bucket8RHPFF), percentileFunction(bucket9RHPFF), percentileFunction(bucket10RHPFF), percentileFunction(bucket11RHPFF), percentileFunction(bucket12RHPFF), percentileFunction(bucket1RHPTF), percentileFunction(bucket2RHPTF), percentileFunction(bucket3RHPTF), percentileFunction(bucket4RHPTF), percentileFunction(bucket5RHPTF), percentileFunction(bucket6RHPTF), percentileFunction(bucket7RHPTF), percentileFunction(bucket8RHPTF), percentileFunction(bucket9RHPTF), percentileFunction(bucket10RHPTF), percentileFunction(bucket11RHPTF), percentileFunction(bucket12RHPTF), percentileFunction(bucket1LHPFF), percentileFunction(bucket2LHPFF), percentileFunction(bucket3LHPFF), percentileFunction(bucket4LHPFF), percentileFunction(bucket5LHPFF), percentileFunction(bucket6LHPFF), percentileFunction(bucket7LHPFF), percentileFunction(bucket8LHPFF), percentileFunction(bucket9LHPFF), percentileFunction(bucket10LHPFF), percentileFunction(bucket11LHPFF), percentileFunction(bucket12LHPFF), percentileFunction(bucket1LHPTF), percentileFunction(bucket2LHPTF), percentileFunction(bucket3LHPTF), percentileFunction(bucket4LHPTF), percentileFunction(bucket5LHPTF), percentileFunction(bucket6LHPTF), percentileFunction(bucket7LHPTF), percentileFunction(bucket8LHPTF), percentileFunction(bucket9LHPTF), percentileFunction(bucket10LHPTF), percentileFunction(bucket11LHPTF), percentileFunction(bucket12LHPTF), simplify = FALSE)

write.csv(totalVeloTable, "NCAA Velocity Percentiles by Bucket all Fastballs.csv", sep = "\t", row.names = FALSE)


# All Pitches Velocity
allPitchVelo <- mapply(c, percentileFunction(filterRightFour), percentileFunction(filterRightTwo), percentileFunction(filterLeftFour), percentileFunction(filterLeftTwo), simplify = FALSE)

write.csv(allPitchVelo, "NCAA Velocity Percentiles for all Fastballs.csv", sep = "\t", row.names = FALSE)

# Induced Vertical Break
totalIVBTable <- mapply(c, percentileFunction2(bucket1RHPFF), percentileFunction2(bucket2RHPFF), percentileFunction2(bucket3RHPFF), percentileFunction2(bucket4RHPFF), percentileFunction2(bucket5RHPFF), percentileFunction2(bucket6RHPFF), percentileFunction2(bucket7RHPFF), percentileFunction2(bucket8RHPFF), percentileFunction2(bucket9RHPFF), percentileFunction2(bucket10RHPFF), percentileFunction2(bucket11RHPFF), percentileFunction2(bucket12RHPFF), percentileFunction2(bucket1RHPTF), percentileFunction2(bucket2RHPTF), percentileFunction2(bucket3RHPTF), percentileFunction2(bucket4RHPTF), percentileFunction2(bucket5RHPTF), percentileFunction2(bucket6RHPTF), percentileFunction2(bucket7RHPTF), percentileFunction2(bucket8RHPTF), percentileFunction2(bucket9RHPTF), percentileFunction2(bucket10RHPTF), percentileFunction2(bucket11RHPTF), percentileFunction2(bucket12RHPTF), percentileFunction2(bucket1LHPFF), percentileFunction2(bucket2LHPFF), percentileFunction2(bucket3LHPFF), percentileFunction2(bucket4LHPFF), percentileFunction2(bucket5LHPFF), percentileFunction2(bucket6LHPFF), percentileFunction2(bucket7LHPFF), percentileFunction2(bucket8LHPFF), percentileFunction2(bucket9LHPFF), percentileFunction2(bucket10LHPFF), percentileFunction2(bucket11LHPFF), percentileFunction2(bucket12LHPFF), percentileFunction2(bucket1LHPTF), percentileFunction2(bucket2LHPTF), percentileFunction2(bucket3LHPTF), percentileFunction2(bucket4LHPTF), percentileFunction2(bucket5LHPTF), percentileFunction2(bucket6LHPTF), percentileFunction2(bucket7LHPTF), percentileFunction2(bucket8LHPTF), percentileFunction2(bucket9LHPTF), percentileFunction2(bucket10LHPTF), percentileFunction2(bucket11LHPTF), percentileFunction2(bucket12LHPTF), simplify = FALSE)

write.csv(totalIVBTable, "NCAA Induced Vertical Percentiles by Bucket all Fastballs.csv", sep = "\t", row.names = FALSE)


# All Pitch Vertical Break

allPitchIVB <- mapply(c, percentileFunction2(filterRightFour), percentileFunction2(filterRightTwo), percentileFunction2(filterLeftFour), percentileFunction2(filterLeftTwo), simplify = FALSE)

write.csv(allPitchIVB, "NCAA Induced Vertical Percentiles for all Fastballs.csv", sep = "\t", row.names = FALSE)


# Horizontal Break
totalHorzBreakTable <- mapply(c, percentileFunction3(bucket1RHPFF), percentileFunction3(bucket2RHPFF), percentileFunction3(bucket3RHPFF), percentileFunction3(bucket4RHPFF), percentileFunction3(bucket5RHPFF), percentileFunction3(bucket6RHPFF), percentileFunction3(bucket7RHPFF), percentileFunction3(bucket8RHPFF), percentileFunction3(bucket9RHPFF), percentileFunction3(bucket10RHPFF), percentileFunction3(bucket11RHPFF), percentileFunction3(bucket12RHPFF), percentileFunction3(bucket1RHPTF), percentileFunction3(bucket2RHPTF), percentileFunction3(bucket3RHPTF), percentileFunction3(bucket4RHPTF), percentileFunction3(bucket5RHPTF), percentileFunction3(bucket6RHPTF), percentileFunction3(bucket7RHPTF), percentileFunction3(bucket8RHPTF), percentileFunction3(bucket9RHPTF), percentileFunction3(bucket10RHPTF), percentileFunction3(bucket11RHPTF), percentileFunction3(bucket12RHPTF), percentileFunction3(bucket1LHPFF), percentileFunction3(bucket2LHPFF), percentileFunction3(bucket3LHPFF), percentileFunction3(bucket4LHPFF), percentileFunction3(bucket5LHPFF), percentileFunction3(bucket6LHPFF), percentileFunction3(bucket7LHPFF), percentileFunction3(bucket8LHPFF), percentileFunction3(bucket9LHPFF), percentileFunction3(bucket10LHPFF), percentileFunction3(bucket11LHPFF), percentileFunction3(bucket12LHPFF), percentileFunction3(bucket1LHPTF), percentileFunction3(bucket2LHPTF), percentileFunction3(bucket3LHPTF), percentileFunction3(bucket4LHPTF), percentileFunction3(bucket5LHPTF), percentileFunction3(bucket6LHPTF), percentileFunction3(bucket7LHPTF), percentileFunction3(bucket8LHPTF), percentileFunction3(bucket9LHPTF), percentileFunction3(bucket10LHPTF), percentileFunction3(bucket11LHPTF), percentileFunction3(bucket12LHPTF), simplify = FALSE)

write.csv(totalHorzBreakTable, "NCAA Horizontal Break Percentiles by Bucket all Fastballs.csv", sep = "\t", row.names = FALSE)

# All Pitch Horizontal Break

allPitchHorz <- mapply(c, percentileFunction3(filterRightFour), percentileFunction3(filterRightTwo), percentileFunction3(filterLeftFour), percentileFunction3(filterLeftTwo), simplify = FALSE)

write.csv(allPitchHorz, "NCAA Horizontal Break Percentiles for all Fastballs.csv", sep = "\t", row.names = FALSE)


# Extension
totalExtensionTable <- mapply(c, percentileFunction4(bucket1RHPFF), percentileFunction4(bucket2RHPFF), percentileFunction4(bucket3RHPFF), percentileFunction4(bucket4RHPFF), percentileFunction4(bucket5RHPFF), percentileFunction4(bucket6RHPFF), percentileFunction4(bucket7RHPFF), percentileFunction4(bucket8RHPFF), percentileFunction4(bucket9RHPFF), percentileFunction4(bucket10RHPFF), percentileFunction4(bucket11RHPFF), percentileFunction4(bucket12RHPFF), percentileFunction4(bucket1RHPTF), percentileFunction4(bucket2RHPTF), percentileFunction4(bucket3RHPTF), percentileFunction4(bucket4RHPTF), percentileFunction4(bucket5RHPTF), percentileFunction4(bucket6RHPTF), percentileFunction4(bucket7RHPTF), percentileFunction4(bucket8RHPTF), percentileFunction4(bucket9RHPTF), percentileFunction4(bucket10RHPTF), percentileFunction4(bucket11RHPTF), percentileFunction4(bucket12RHPTF), percentileFunction4(bucket1LHPFF), percentileFunction4(bucket2LHPFF), percentileFunction4(bucket3LHPFF), percentileFunction4(bucket4LHPFF), percentileFunction4(bucket5LHPFF), percentileFunction4(bucket6LHPFF), percentileFunction4(bucket7LHPFF), percentileFunction4(bucket8LHPFF), percentileFunction4(bucket9LHPFF), percentileFunction4(bucket10LHPFF), percentileFunction4(bucket11LHPFF), percentileFunction4(bucket12LHPFF), percentileFunction4(bucket1LHPTF), percentileFunction4(bucket2LHPTF), percentileFunction4(bucket3LHPTF), percentileFunction4(bucket4LHPTF), percentileFunction4(bucket5LHPTF), percentileFunction4(bucket6LHPTF), percentileFunction4(bucket7LHPTF), percentileFunction4(bucket8LHPTF), percentileFunction4(bucket9LHPTF), percentileFunction4(bucket10LHPTF), percentileFunction4(bucket11LHPTF), percentileFunction4(bucket12LHPTF), simplify = FALSE)

write.csv(totalExtensionTable, "NCAA Extension Break Percentiles by Bucket all Fastballs.csv", sep = "\t", row.names = FALSE)

# All Pitch Extension

allPitchExt <- mapply(c, percentileFunction4(filterRightFour), percentileFunction4(filterRightTwo), percentileFunction4(filterLeftFour), percentileFunction4(filterLeftTwo), simplify = FALSE)

write.csv(allPitchExt, "NCAA Extension Percentiles for all Fastballs.csv", sep = "\t", row.names = FALSE)
```


# Validation

```{r}
#Load Data


accRightFourStuff <- read.csv("2022 NCAA Right Handed Four Seam Stuff+ Validation.csv")
accRightTwoStuff <- read.csv("2022 NCAA Right Handed Two Seam Stuff+ Validation.csv")
accLeftFourStuff <- read.csv("2022 NCAA Left Handed Four Seam Stuff+ Validation.csv")
accLeftTwoStuff <- read.csv("2022 NCAA Left Handed Two Seam Stuff+ Validation.csv")

accLeftTwoStuff

accRightFourStuff

length(rightFourWhiff)


#Merge Data into one table

mergeFunction <- function(x,y,z){
  whiffTable <- data.frame(Pitcher = unique(x$Pitcher), Whiff = y, stringsAsFactors = FALSE)
  mergedTable <- merge(z, whiffTable, by.y = "Pitcher")
  return(mergedTable)
}

rightFourMerged <- mergeFunction(accRightFour, rightFourWhiff, accRightFourStuff)
rightTwoMerged <- mergeFunction(accRightTwo, rightTwoWhiff, accRightTwoStuff)
leftFourMerged <- mergeFunction(accLeftFour, leftFourWhiff, accLeftFourStuff)
leftTwoMerged <- mergeFunction(accLeftTwo, leftTwoWhiff, accLeftTwoStuff)

rightFourMerged

```

# Multiple Linear Regression:

```{r}
summary(lm(Whiff ~ Stuff., data = rightFourMerged))

summary(lm(Whiff ~ Velo.. + Horz.. + IVB.. + IVB.by.Bucket.., data = rightFourMerged))

summary(lm(Whiff ~ Adjusted.Velo + Horz + IVB, data = rightFourMerged))

```
# Left Four Seam Fastball

```{r}
summary(lm(Whiff ~ Stuff., data = leftFourMerged))

summary(lm(Whiff ~ Velo.. + Horz.. + IVB.. + IVB.by.Bucket.., data = leftFourMerged))

summary(lm(Whiff ~ Adjusted.Velo + Horz + IVB, data = leftFourMerged))
```
## Right Handed Two Seamers

```{r}
summary(lm(Whiff ~ Stuff., data = rightTwoMerged))

summary(lm(Whiff ~ Velo.. + Horz.. + IVB.., data = rightTwoMerged))

summary(lm(Whiff ~ Adjusted.Velo + Horz + IVB, data = rightTwoMerged))
```


```{r}
summary(lm(Whiff ~ Stuff., data = leftTwoMerged))

summary(lm(Whiff ~ Velo.. + Horz.. + IVB.., data = leftTwoMerged))

summary(lm(Whiff ~ Adjusted.Velo + Horz + IVB, data = leftTwoMerged))
```





# Calculate Swinging Strike%



## Find Number of Strikes

```{r}
swing <- vector()

strike <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    base <- unique(x$Pitcher)[i]
    pitcher <- x[x$Pitcher == base,]
    swingStrike <- pitcher[pitcher$PitchCall == "StrikeSwinging",]
    swing[i] <- dim(swingStrike)[1]
  }
  return(swing)
}

KRF <- strike(accRightFour)

KRT <- strike(accRightTwo)
KLF <- strike(accLeftFour)
KLT <- strike(accLeftTwo)

KRF

KRF[93]
```

# Find Number of Pitches

```{r}

#SwStr% involves all pitches, not just swings

count <- vector()
pitchCount2 <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    number <- 0
    base <- unique(x$Pitcher)[i]
    pitcher <- x[x$Pitcher == base,]
    count[i] <- length(pitcher$PitchNo)
  }
  return(count)
}



rightFourCount <- pitchCount2(accRightFour)
leftFourCount <- pitchCount2(accLeftFour)
rightTwoCount <- pitchCount2(accRightTwo)
leftTwoCount <- pitchCount2(accLeftTwo)

rightFourCount
leftFourCount
rightTwoCount
leftTwoCount

rightFourCount[1]
```


# Find SwStr%

```{r}
swStr <- vector()

swStrFunctionRF <- function(){
  for(i in 1:length(KRF)){
    swStr[i] <- 100*KRF[i] / rightFourCount[i]
  }
  return(swStr)
}

swStrFunctionRT <- function(){
  for(i in 1:length(KRT)){
    swStr[i] <- 100*KRT[i] / rightTwoCount[i]
  }
  return(swStr)
}

swStrFunctionLF <- function(){
  for(i in 1:length(KLF)){
    swStr[i] <- 100*KLF[i] / leftFourCount[i]
  }
  return(swStr)
}

whiffFunctionLT <- function(){
  for(i in 1:length(KLT)){
    whiff[i] <- 100*KLT[i] / leftTwoCount[i]
  }
  return(whiff)
}

rightFourWhiff <- whiffFunctionRF()
rightTwoWhiff <- whiffFunctionRT()
leftFourWhiff <- whiffFunctionLF()
leftTwoWhiff <- whiffFunctionLT()

rightFourWhiff

rightFourWhiff[93]
```


# Calculate Barrel %

```{r}
barrels <- vector()
hitting <- vector()

# Counting the Number of Barrels
barrel <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    data <- x[x$Pitcher == unique(x$Pitcher)[i],]
    for(a in 1:length(data$PitchNo)){
      count <- 0
      if(data$TaggedHitType[a] == "FlyBall"){ #Barrel Component 1
        if(data$ExitSpeed[a] > 95 & is.na(data$ExitSpeed[a]) == FALSE){ #Barrel Component 2
          count <- count + 1
        }
      }
      else if(data$TaggedHitType[a] == "LineDrive"){
        if(data$ExitSpeed[a] > 95 & is.na(data$ExitSpeed[a]) == FALSE){
          count <- count + 1
        }
      }
    }
    barrels[i] <- count
  }
  return(barrels)
}

# Counting the Number of Hits
hit <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    count2 <- 0
    if(x$TaggedHitType[i] != "Undefined"){ #Undefined is the default for non-hits
      count2 <- count2 + 1
    }
    hitting[i] <- count2
  }
  return(hitting)
}


rightFourBarrel <- barrel(accRightFour)
leftFourBarrel <- barrel(accLeftFour)
rightTwoBarrel <- barrel(accRightTwo)
leftTwoBarrel <- barrel(accLeftTwo)

rightFourBarrel


rightFourHit <- hit(accRightFour)
leftFourHit <- hit(accLeftFour)
rightTwoHit <- hit(accRightTwo)
leftTwoHit <- hit(accLeftTwo)
```

## Barrel%

```{r}
barreling <- vector()

barrelFunctionRF <- function(){
  for(i in 1:length(rightFourBarrel)){
    barreling[i] <- 100*rightFourBarrel[i] / rightFourHit[i]
  }
  return(barreling)
}

barrelFunctionRT <- function(){
  for(i in 1:length(rightTwoBarrel)){
    barreling[i] <- 100*rightTwoBarrel[i] / rightTwoHit[i]
  }
  return(barreling)
}

barrelFunctionLF <- function(){
  for(i in 1:length(leftFourBarrel)){
    barreling[i] <- 100*leftFourBarrel[i] / leftFourHit[i]
  }
  return(barreling)
}

barrelFunctionLT <- function(){
  for(i in 1:length(leftTwoBarrel)){
    barreling[i] <- 100*leftTwoBarrel[i] / leftTwoHit[i]
  }
  return(barreling)
}

rightFourBarrel <- barrelFunctionRF()
rightTwoBarrel <- barrelFunctionRT()
leftFourBarrel <- barrelFunctionLF()
leftTwoBarrel <- barrelFunctionLT()
```


## Merge Data Tables

```{r}
rightFourMerged2 <- mergeFunction(accRightFour, rightFourBarrel, accRightFourStuff)
rightTwoMerged2 <- mergeFunction(accRightTwo, rightTwoBarrel, accRightTwoStuff)
leftFourMerged2 <- mergeFunction(accLeftFour, leftFourBarrel, accLeftFourStuff)
leftTwoMerged2 <- mergeFunction(accLeftTwo, leftTwoBarrel, accLeftTwoStuff)

rightFourMerged2
```