---
title: "NCAA Fastball Stuff+ Version 8 Part 2"
author: "Alexander Harriman"
date: '2022-05-23'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Percentiles

```{r}
library(beepr)

allVelo <- read.csv("NCAA Velocity Percentiles for all Fastballs.csv")
allInducedVertBucket <- read.csv("NCAA Induced Vertical Percentiles by Bucket all Fastballs.csv")
allInducedVert <- read.csv("NCAA Induced Vertical Percentiles for all Fastballs.csv")
allHorizontal <- read.csv("NCAA Horizontal Break Percentiles for all Fastballs.csv")
allExtension <- read.csv("NCAA Extension Percentiles for all Fastballs.csv")

allVelo
allInducedVertBucket
allInducedVert
allHorizontal
allExtension
```

# Insert ACC Pitcher Files

```{r, warning = FALSE}
library(readxl)
acc2022 <- read_excel("Reduced 2022 Fastballs (5-20, M-Z).xlsx")

# Filter by Minimum Speed
acc2022 <- acc2022[acc2022$RelSpeed > 85.999,]


names <- unique(acc2022$Pitcher)

#There are 4293 pitchers in this dataset.

##We will filter them down by fastball type

beep(3)
```

# Fastball Filtering

```{r}
# Separating lefties and righties

accRight <- acc2022[acc2022$PitcherThrows == "Right",]
accLeft <- acc2022[acc2022$PitcherThrows == "Left",]

dim(accRight)
accRight$`Pitch Assignment`[8420]

# Using tilt to separate four-seamers and two-seamers

rightTilt <- function(){
  for(i in 1:dim(accRight)[1]){
    if(is.na(accRight$`Tilt Coordinate`[i]) == TRUE){
      accRight$`Pitch Assignment`[i] <<- "None"
    }
    else if(accRight$`Tilt Coordinate`[i] > 0.07 && accRight$`Tilt Coordinate`[i] < .135){
      accRight$`Pitch Assignment`[i] <<- "Right Two-Seam"
    }
    else if(accRight$`Tilt Coordinate`[i] > 0.04  && accRight$`Tilt Coordinate`[i] < .065){
      accRight$`Pitch Assignment`[i] <<- "Right Four-Seam"
    }
    else if(accRight$`Tilt Coordinate`[i] > 0.499 && accRight$`Tilt Coordinate`[i] < .532){
      accRight$`Pitch Assignment`[i] <<- "Right Four-Seam"
    }
  }
  return(accRight$`Pitch Assignment`)
}

rightTilt()

accRightFour <- accRight[accRight$`Pitch Assignment` == "Right Four-Seam",]
accRightTwo <- accRight[accRight$`Pitch Assignment` == "Right Two-Seam",]


#Filter Out Bad Data
accRightTwo <- accRightTwo[accRightTwo$Pitcher != "Maxwell, Zachary",]

accRightTwo <- accRightTwo[accRightTwo$Pitcher != "Stohr, Ryan",]
accRightTwo <- accRightTwo[is.na(accRightTwo$Pitcher) == FALSE,]

accRightFour <- accRightFour[is.na(accRightFour$Pitcher) == FALSE,]

dim(accRightFour)
dim(accRightTwo)

leftTilt <- function(){
  for(i in 1:dim(accLeft)[1]){
    if(is.na(accLeft$`Tilt Coordinate`[i]) == TRUE){
      accLeft$`Pitch Assignment`[i] <<- "None"
    }
    else if(accLeft$`Tilt Coordinate`[i] > .3749 && accLeft$`Tilt Coordinate`[i] < .428){
      accLeft$`Pitch Assignment`[i] <<- "Left Two-Seam"
    }
    else if(accLeft$`Tilt Coordinate`[i] >  .437 && accLeft$`Tilt Coordinate`[i] < 0.501){
      accLeft$`Pitch Assignment`[i] <<- "Left Four-Seam"
    }
  }
  return(accRight$`Pitch Assignment`)
}

leftTilt()


accLeftFour <- accLeft[accLeft$`Pitch Assignment` == "Left Four-Seam",]
accLeftTwo <- accLeft[accLeft$`Pitch Assignment` == "Left Two-Seam",]
dim(accLeftFour)
dim(accLeftTwo)
```

```{r}
# Set Pitch Minimum at 20 Pitches:

pitcherClearLF <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    x <- accLeftFour
    base <- unique(x$Pitcher)[i]
    if(length(x[x$Pitcher == base,]$PitchNo) < 20){
      accLeftFour <<- subset(x, x$Pitcher != base)
    }
  }
  return(accLeftFour)
}

pitcherClearLT <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    x <- accLeftTwo
    base <- unique(x$Pitcher)[i]
    if(length(x[x$Pitcher == base,]$PitchNo) < 20){
      accLeftTwo <<- subset(x, x$Pitcher != base)
    }
  }
  return(accLeftTwo)
}

pitcherClearRF <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    x <- accRightFour
    base <- unique(x$Pitcher)[i]
    if(length(x[x$Pitcher == base,]$PitchNo) < 20){
      accRightFour <<- subset(x, x$Pitcher != base)
    }
  }
  return(accRightFour)
}

pitcherClearRT <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    x <- accRightTwo
    base <- unique(x$Pitcher)[i]
    if(length(x[x$Pitcher == base,]$PitchNo) < 20){
      accRightTwo <<- subset(x, x$Pitcher != base)
    }
  }
  return(accRightTwo)
}

dim(accLeftFour)
pitcherClearLF(accLeftFour)
dim(accLeftFour)

dim(accRightTwo)
pitcherClearRT(accRightTwo)
dim(accRightTwo)

dim(accRightFour)
pitcherClearRF(accRightFour)
dim(accRightFour)

dim(accLeftTwo)
pitcherClearLT(accLeftTwo)
dim(accLeftTwo)

beep(3)
```




# Find Average Release Height for All Pitchers
```{r}
# Find the number of unique pitchers in each of the four types

rffName <- unique(accRightFour$Pitcher) #3175 pitchers
lffName <- unique(accLeftFour$Pitcher) #561 pitchers
rtfName <- unique(accRightTwo$Pitcher) #683 pitchers
ltfName <- unique(accLeftTwo$Pitcher) #205 pitchers

length(rffName)
length(lffName)
length(rtfName)
length(ltfName)


# Find the teams for each pitcher
team <- vector()
teamFinder <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    base <- unique(x$Pitcher)[i]
    pitcher <- x[x$Pitcher == base,]
    team[i] <- unique(pitcher$PitcherTeam)[1]
  }
  return(team)
}

rightFourTeam <- teamFinder(accRightFour)
leftFourTeam <- teamFinder(accLeftFour)
rightTwoTeam <- teamFinder(accRightTwo)
leftTwoTeam <- teamFinder(accLeftTwo)

# Find Pitch Count

count <- vector()
pitchCount <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    base <- unique(x$Pitcher)[i]
    pitcher <- x[x$Pitcher == base,]
    count[i] <- dim(pitcher)[1]
  }
  return(count)
}

rightFourCount <- pitchCount(accRightFour)
leftFourCount <- pitchCount(accLeftFour)
rightTwoCount <- pitchCount(accRightTwo)
leftTwoCount <- pitchCount(accLeftTwo)

rightFourCount
leftFourCount
rightTwoCount
leftTwoCount


# Find Level within NCAA

level <- vector()
pitchLevel <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    base <- unique(x$Pitcher)[i]
    pitcher <- x[x$Pitcher == base,]
    level[i] <- unique(pitcher$Level)[1]
  }
  return(level)
}


rightFourLevel <- pitchLevel(accRightFour)
leftFourLevel <- pitchLevel(accLeftFour)
rightTwoLevel <- pitchLevel(accRightTwo)
leftTwoLevel <- pitchLevel(accLeftTwo)

beep(3)
```




```{r}
# Unique Release Height for the Four-Seamers, needed for buckets

RH <- vector()
relHeight <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    base <- unique(x$Pitcher)[i]
    pitcher <- x[x$Pitcher == base,]
    height <- pitcher[is.na(pitcher$RelHeight) == FALSE,]
    RH[i] <- mean(height$RelHeight)
  }
  return(RH)
}

rightFourHeight <- relHeight(accRightFour)
leftFourHeight <- relHeight(accLeftFour)
rightTwoHeight <- relHeight(accRightTwo)
leftTwoHeight <- relHeight(accLeftTwo)

rightFourHeight
```


# Average Horizontal Break

```{r}
horz <- vector()
horzFunction <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    base <- unique(x$Pitcher)[i]
    pitcher <- x[x$Pitcher == base,]
    horiz <- pitcher[is.na(pitcher$HorzBreak) == FALSE,]
    horz[i] <- abs(mean(horiz$HorzBreak))
  }
  return(horz)
}
rightFourHorz <- horzFunction(accRightFour)
rightTwoHorz <- horzFunction(accRightTwo)
leftFourHorz <- horzFunction(accLeftFour)
leftTwoHorz <- horzFunction(accLeftTwo)

rightFourHorz
```

# Average Velo

```{r}
velo <- vector()
veloFunction <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    base <- unique(x$Pitcher)[i]
    pitcher <- x[x$Pitcher == base,]
    veloc <- pitcher[is.na(pitcher$RelSpeed) == FALSE,]
    velo[i] <- abs(mean(veloc$RelSpeed))
  }
  return(velo)
}

rightFourVelo <- veloFunction(accRightFour)
rightTwoVelo <- veloFunction(accRightTwo)
leftFourVelo <- veloFunction(accLeftFour)
leftTwoVelo <- veloFunction(accLeftTwo)

rightFourVelo
```

# Average IVB (No Bucket)

```{r}
IVB <- vector()
IVBFunction <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    base <- unique(x$Pitcher)[i]
    pitcher <- x[x$Pitcher == base,]
    induce <- pitcher[is.na(pitcher$InducedVertBreak) == FALSE,]
    IVB[i] <- abs(mean(induce$InducedVertBreak))
  }
  return(IVB)
}

rightFourIVB <- IVBFunction(accRightFour)
rightTwoIVB <- IVBFunction(accRightTwo)
leftFourIVB <- IVBFunction(accLeftFour)
leftTwoIVB <- IVBFunction(accLeftTwo)

rightTwoIVB
```

# Average Extension
```{r}
extension <- vector()
extensionFunction <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    base <- unique(x$Pitcher)[i]
    pitcher <- x[x$Pitcher == base,]
    exten <- pitcher[is.na(pitcher$Extension) == FALSE,]
    extension[i] <- mean(exten$Extension)
  }
  return(extension)
}

rightFourExtension <- extensionFunction(accRightFour)
rightTwoExtension <- extensionFunction(accRightTwo)
leftFourExtension <- extensionFunction(accLeftFour)
leftTwoExtension <- extensionFunction(accLeftTwo)

rightFourExtension
```

# Finding Percentile Value for Each Pitcher

For each of these functions, x will represent the row value, which is determined by the bucket.

```{r}
# Induced Vertical Break by Bucket 
## Only for Four-Seamers

bucket <- vector()
rightBucketFunction <- function(){
  for(i in 1:length(unique(accRightFour$Pitcher))){
    if(rightFourHeight[i] > 4.5 && rightFourHeight[i] < 4.7){
      bucket[i] <- 1
    }
    else if(rightFourHeight[i] > 4.7 && rightFourHeight[i] < 4.9){
      bucket[i] <- 2
    }
    else if(rightFourHeight[i] > 4.9 && rightFourHeight[i] < 5.1){
      bucket[i] <- 3
    }
    else if(rightFourHeight[i] > 5.1 && rightFourHeight[i] < 5.3){
      bucket[i] <- 4
    }
    else if(rightFourHeight[i] > 5.3 && rightFourHeight[i] < 5.5){
      bucket[i] <- 5
    }
    else if(rightFourHeight[i] > 5.5 && rightFourHeight[i] < 5.7){
      bucket[i] <- 6
    }
    else if(rightFourHeight[i] > 5.7 && rightFourHeight[i] < 5.9){
      bucket[i] <- 7
    }
    else if(rightFourHeight[i] > 5.9 && rightFourHeight[i] < 6.1){
      bucket[i] <- 8
    }
    else if(rightFourHeight[i] > 6.1 && rightFourHeight[i] < 6.3){
      bucket[i] <- 9
    }
    else if(rightFourHeight[i] > 6.3 && rightFourHeight[i] < 6.5){
      bucket[i] <- 10
    }
    else if(rightFourHeight[i] > 6.5 && rightFourHeight[i] < 6.7){
      bucket[i] <- 11
    }
    else if(rightFourHeight[i] > 6.7 && rightFourHeight[i] < 6.9){
      bucket[i] <- 12
    }
    else if(rightFourHeight[i] < 4.5){
      bucket[i] <- 1
    }
    else if(rightFourHeight[i] > 6.9){
      bucket[i] <- 12
    }
  }
  return(bucket)
}


leftBucketFunction <- function(){
  for(i in 1:length(unique(accLeftFour$Pitcher))){
    if(leftFourHeight[i] > 4.5 && leftFourHeight[i] < 4.7){
      bucket[i] <- 25
    }
    else if(leftFourHeight[i] > 4.7 && leftFourHeight[i] < 4.9){
      bucket[i] <- 26
    }
    else if(leftFourHeight[i] > 4.9 && leftFourHeight[i] < 5.1){
      bucket[i] <- 27
    }
    else if(leftFourHeight[i] > 5.1 && leftFourHeight[i] < 5.3){
      bucket[i] <- 28
    }
    else if(leftFourHeight[i] > 5.3 && leftFourHeight[i] < 5.5){
      bucket[i] <- 29
    }
    else if(leftFourHeight[i] > 5.5 && leftFourHeight[i] < 5.7){
      bucket[i] <- 30
    }
    else if(leftFourHeight[i] > 5.7 && leftFourHeight[i] < 5.9){
      bucket[i] <- 31
    }
    else if(leftFourHeight[i] > 5.9 && leftFourHeight[i] < 6.1){
      bucket[i] <- 32
    }
    else if(leftFourHeight[i] > 6.1 && leftFourHeight[i] < 6.3){
      bucket[i] <- 33
    }
    else if(leftFourHeight[i] > 6.3 && leftFourHeight[i] < 6.5){
      bucket[i] <- 34
    }
    else if(leftFourHeight[i] > 6.5 && leftFourHeight[i] < 6.7){
      bucket[i] <- 35
    }
    else if(leftFourHeight[i] > 6.7 && leftFourHeight[i] < 6.9){
      bucket[i] <- 36
    }
    else if(leftFourHeight[i] < 4.5){
      bucket[i] <- 25
    }
    else if(leftFourHeight[i] > 6.9){
      bucket[i] <- 36
    }
  }
  return(bucket)
}

rightFourBucket <- rightBucketFunction()
leftFourBucket <- leftBucketFunction()

beep(3)
```

```{r}
IVBBucketPercentileVect <- vector()
rightComparisonIVBBucket <- function(x){
  for(a in 1:length(unique(x$Pitcher))){
    for(j in 1:101){
    if(rightFourIVB[a] > allInducedVertBucket[rightFourBucket[a],j] && rightFourIVB[a] < allInducedVertBucket[rightFourBucket[a],j+1]){
        IVBBucketPercentileVect[a] <- j
    }
  }
  }
  return(IVBBucketPercentileVect)
}


leftComparisonIVBBucket <- function(x){
  for(a in 1:length(unique(x$Pitcher))){
    for(j in 1:101){
    if(leftFourIVB[a] > allInducedVertBucket[leftFourBucket[a],j] && leftFourIVB[a] < allInducedVertBucket[leftFourBucket[a],j+1]){
        IVBBucketPercentileVect[a] <- j
    }
  }
  }
  return(IVBBucketPercentileVect)
}

# Create Vector of Percentiles 

rightIVBBucketPercentile <- rightComparisonIVBBucket(accRightFour)
leftIVBBucketPercentile <- leftComparisonIVBBucket(accLeftFour)


rightIVBBucketPercentile[93]
```


```{r}
# Induced Vertical Break not by Bucket
rightTwoIVB
IVBPercentile <- vector()

rightFourComparisonIVB <- function(x){
  for(i in 1:length(unique(x$Pitcher))){  #Right Handed Four Seamer
    for(a in 1:101){
      if(rightFourIVB[i] > allInducedVert[1,a] && rightFourIVB[i] < allInducedVert[1,a+1]){
        IVBPercentile[i] <- a
      }
    }
  }
  return(IVBPercentile)
}
rightTwoComparisonIVB <- function(x){
  for(i in 1:length(unique(x$Pitcher))){  #Right Handed Two Seamer
    for(a in 1:101){
      if(rightTwoIVB[i] > allInducedVert[2,101]){
        IVBPercentile[i] <- 100
      }
      else if(rightTwoIVB[i] > allInducedVert[2,a] && rightTwoIVB[i] < allInducedVert[2,a+1]){
        IVBPercentile[i] <- a
      }
    }
  }
  return(IVBPercentile)
}
leftFourComparisonIVB <- function(x){
  for(i in 1:length(unique(x$Pitcher))){  #Left Handed Four Seamer
    for(a in 1:101){
      if(leftFourIVB[i] > allInducedVert[3,a] && leftFourIVB[i] < allInducedVert[3,a+1]){
        IVBPercentile[i] <- a
      }
    }
  }
  return(IVBPercentile)
}
leftTwoComparisonIVB <- function(x){
  for(i in 1:length(unique(x$Pitcher))){  #Left Handed Two Seamer
    for(a in 1:101){
      if(leftTwoIVB[i] > allInducedVert[4,a] && leftTwoIVB[i] < allInducedVert[4,a+1]){
        IVBPercentile[i] <- a
      }
    }
  }
  return(IVBPercentile)
}
 

#Create Vector of Percentiles

## For the Two-Seamers, Flip the Percentiles

rightFourIVBPercentile <- rightFourComparisonIVB(accRightFour)
rightTwoIVBPercentile <- rightTwoComparisonIVB(accRightTwo)
leftFourIVBPercentile <- leftFourComparisonIVB(accLeftFour)
leftTwoIVBPercentile <- leftTwoComparisonIVB(accLeftTwo)


#Flip the two-seamer percentiles:

percentileFlip <- function(x){
  for(i in 1:length(x)){
    IVBPercentile[i] <- 100 - x[i]
  }
  return(IVBPercentile)
}

rightTwoIVBPercentileYes <- percentileFlip(rightTwoIVBPercentile)
leftTwoIVBPercentileYes <- percentileFlip(leftTwoIVBPercentile)

rightFourIVBPercentile

rightTwoIVBPercentileYes

leftTwoIVBPercentile
leftTwoIVBPercentileYes

beep(3)
```

```{r}
# Horizontal Break

horzPercentile <- vector()

rightFourComparisonHorz <- function(x){
  for(i in 1:length(unique(x$Pitcher))){  #Left Handed Four Seamer
    for(a in 1:101){
      if(rightFourHorz[i] > allHorizontal[1,a] && rightFourHorz[i] < allHorizontal[1,a+1]){
        horzPercentile[i] <- a
      }
    }
  }
  return(horzPercentile)
}
rightTwoComparisonHorz <- function(x){
  for(i in 1:length(unique(x$Pitcher))){  #Left Handed Four Seamer
    for(a in 1:101){
      if(rightTwoHorz[i] > allHorizontal[2,a] && rightTwoHorz[i] < allHorizontal[2,a+1]){
        horzPercentile[i] <- a
      }
    }
  }
  return(horzPercentile)
}
leftFourComparisonHorz <- function(x){
  for(i in 1:length(unique(x$Pitcher))){  #Left Handed Four Seamer
    for(a in 1:101){
      if(leftFourHorz[i] > allHorizontal[3,a] && leftFourHorz[i] < allHorizontal[3,a+1]){
        horzPercentile[i] <- a
      }
    }
  }
  return(horzPercentile)
}
leftTwoComparisonHorz <- function(x){
  for(i in 1:length(unique(x$Pitcher))){  #Left Handed Four Seamer
    for(a in 1:101){
      if(leftTwoHorz[i] > allHorizontal[4,a] && leftTwoHorz[i] < allHorizontal[4,a+1]){
        horzPercentile[i] <- a
      }
    }
  }
  return(horzPercentile)
}


rightFourHorzPercentile <- rightFourComparisonHorz(accRightFour)
rightTwoHorzPercentile <- rightTwoComparisonHorz(accRightTwo)
leftFourHorzPercentile <- leftFourComparisonHorz(accLeftFour)
leftTwoHorzPercentile <- leftTwoComparisonHorz(accLeftTwo)

rightFourHorzPercentile
```

# Adjusting the Velocity by Extension

```{r}
veloPercentile <- vector()

rightFourComparisonVelo <- function(x){
  for(i in 1:length(unique(x$Pitcher))){  #Left Handed Four Seamer
    for(a in 1:101){
      if(rightFourAdjustVelo[i] > allVelo[1,a] && rightFourAdjustVelo[i] < allVelo[1,a+1]){
        veloPercentile[i] <- a
      }
    }
  }
  return(veloPercentile)
}
rightTwoComparisonVelo <- function(x){
  for(i in 1:length(unique(x$Pitcher))){  #Left Handed Four Seamer
    for(a in 1:101){
      if(rightTwoAdjustVelo[i] > allVelo[2,101]){
        veloPercentile[i] <- 100
      }
      else if(rightTwoAdjustVelo[i] > allVelo[2,a] && rightTwoAdjustVelo[i] < allVelo[2,a+1]){
        veloPercentile[i] <- a
      }
    }
  }
  return(veloPercentile)
}
leftFourComparisonVelo <- function(x){
  for(i in 1:length(unique(x$Pitcher))){  #Left Handed Four Seamer
    for(a in 1:101){
      if(leftFourAdjustVelo[i] > allVelo[3,a] && leftFourAdjustVelo[i] < allVelo[3,a+1]){
        veloPercentile[i] <- a
      }
    }
  }
  return(veloPercentile)
}
leftTwoComparisonVelo <- function(x){
  for(i in 1:length(unique(x$Pitcher))){  #Left Handed Four Seamer
    for(a in 1:101){
      if(leftTwoAdjustVelo[i] > allVelo[4,a] && leftTwoAdjustVelo[i] < allVelo[4,a+1]){
        veloPercentile[i] <- a
      }
    }
  }
  return(veloPercentile)
}

adjustedVelo <- vector()

rightFourAdjustedVelo <- function(x){
  for(i in 1:length(unique(x$Pitcher))){  #Left Handed Four Seamer
    adjustedVelo[i] <- rightFourVelo[i] + (1.6/12)*(rightFourExtension[i] - allExtension[1,50])
  }
  return(adjustedVelo)
}
rightTwoAdjustedVelo <- function(x){
  for(i in 1:length(unique(x$Pitcher))){  #Left Handed Four Seamer
    adjustedVelo[i] <- rightTwoVelo[i] + (1.6/12)*(rightTwoExtension[i] - allExtension[2,50])
  }
  return(adjustedVelo)
}
leftFourAdjustedVelo <- function(x){
  for(i in 1:length(unique(x$Pitcher))){  #Left Handed Four Seamer
    adjustedVelo[i] <- leftFourVelo[i] + (1.6/12)*(leftFourExtension[i] - allExtension[3,50])
  }
  return(adjustedVelo)
}
leftTwoAdjustedVelo <- function(x){
  for(i in 1:length(unique(x$Pitcher))){  #Left Handed Four Seamer
    adjustedVelo[i] <- leftTwoVelo[i] + (1.6/12)*(leftTwoExtension[i] - allExtension[4,50])
  }
  return(adjustedVelo)
}

rightFourAdjustVelo <- rightFourAdjustedVelo(accRightFour)
rightTwoAdjustVelo <- rightTwoAdjustedVelo(accRightTwo)
leftFourAdjustVelo <- leftFourAdjustedVelo(accLeftFour)
leftTwoAdjustVelo <- leftTwoAdjustedVelo(accLeftTwo)

rightFourAdjustVelo
rightTwoAdjustVelo

max(rightTwoAdjustVelo)


rightFourVeloPercentile <- rightFourComparisonVelo(accRightFour)
rightTwoVeloPercentile <- rightTwoComparisonVelo(accRightTwo)
leftFourVeloPercentile <- leftFourComparisonVelo(accLeftFour)
leftTwoVeloPercentile <- leftTwoComparisonVelo(accLeftTwo)

rightFourVeloPercentile
leftFourVeloPercentile[15]

leftFourAdjustVelo[15]

leftFourAdjustVelo[22]

beep(3)
```

# Find Mean and Standard Deviation for Metrics

```{r}
rightFourVeloMean <- mean(accRightFour$RelSpeed)
rightTwoVeloMean <- mean(accRightTwo$RelSpeed)
leftFourVeloMean <- mean(accLeftFour$RelSpeed)
leftTwoVeloMean <- mean(accLeftTwo$RelSpeed)

rightFourVeloSD <- sd(accRightFour$RelSpeed)
rightTwoVeloSD <- sd(accRightTwo$RelSpeed)
leftFourVeloSD <- sd(accLeftFour$RelSpeed)
leftTwoVeloSD <- sd(accLeftTwo$RelSpeed)

rightFourHorzMean <- abs(mean(accRightFour$HorzBreak))
rightTwoHorzMean <- abs(mean(accRightTwo$HorzBreak))
leftFourHorzMean <- abs(mean(accLeftFour$HorzBreak))
leftTwoHorzMean <- abs(mean(accLeftTwo$HorzBreak))

rightFourHorzSD <- abs(sd(accRightFour$HorzBreak))
rightTwoHorzSD <- abs(sd(accRightTwo$HorzBreak))
leftFourHorzSD <- abs(sd(accLeftFour$HorzBreak))
leftTwoHorzSD <- abs(sd(accLeftTwo$HorzBreak))

rightFourIVBMean <- mean(accRightFour$InducedVertBreak)
rightTwoIVBMean <- mean(accRightTwo$InducedVertBreak)
leftFourIVBMean <- mean(accLeftFour$InducedVertBreak)
leftTwoIVBMean <- mean(accLeftTwo$InducedVertBreak)

rightFourIVBSD <- sd(accRightFour$InducedVertBreak)
rightTwoIVBSD <- sd(accRightTwo$InducedVertBreak)
leftFourIVBSD <- sd(accLeftFour$InducedVertBreak)
leftTwoIVBSD <- sd(accLeftTwo$InducedVertBreak)

leftTwoIVBMean
leftTwoHorzMean
leftTwoVeloMean

leftTwoIVBSD
leftTwoHorzSD
leftTwoVeloSD

leftFourIVBMean
leftFourHorzMean
leftFourVeloMean

leftFourIVBSD
leftFourHorzSD
leftFourVeloSD

rightFourIVBMean
rightFourHorzMean
rightFourVeloMean

rightFourIVBSD
rightFourHorzSD
rightFourVeloSD

rightTwoIVBMean
rightTwoHorzMean
rightTwoVeloMean

rightTwoIVBSD
rightTwoHorzSD
rightTwoVeloSD
```

## Input Average Induced Vert by Bucket

```{r}
rightFourBucketMean <- c(15.40361,
15.2989,
15.41919,
15.70848,
16.47215,
16.99061,
17.53288,
17.83591,
18.43497,
18.73269,
19.50746,
19.13517
)
leftFourBucketMean <- c(15.31076,
15.21698,
15.55029,
15.81432,
16.4152,
16.97381,
17.48811,
17.82028,
18.35664,
18.65516,
19.40202,
19.43625
)

rightFourBucketSD <- c(3.86845,
3.157976,
2.983433,
2.964166,
3.026137,
3.08061,
3.105723,
3.246047,
3.289604,
3.256543,
3.622131,
3.622383
)
leftFourBucketSD <- c(3.779763,
3.129162,
2.947515,
2.973885,
3.037003,
3.084712,
3.099824,
3.233041,
3.286215,
3.225889,
3.577994,
3.550872
)
```

# Create a Stuff Value: Four Seamers

```{r}
stuffVect <- vector()

RightFourStuff <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    stuffVect[i] <- (.45 * (1+ ((rightFourAdjustVelo[i] - rightFourVeloMean) / rightFourVeloSD))) + (.15 * (1+((rightFourHorz[i] - rightFourHorzMean) / rightFourHorzSD))) + (.3 * (1+((rightFourIVB[i] - rightFourIVBMean) / rightFourIVBSD))) + (.1 * (1+((rightFourIVB[i] - rightFourBucketMean[rightFourBucket[i]]) / rightFourBucketSD[rightFourBucket[i]])))
  }
  return(stuffVect)
}

rightFourSeamStuff <- RightFourStuff(accRightFour)

rightFourSeamStuff

```

```{r}
stuffVect <- vector()

LeftFourStuff <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    stuffVect[i] <- (.45 * (1+ ((leftFourAdjustVelo[i] - leftFourVeloMean) / leftFourVeloSD))) + (.15 * (1+((leftFourHorz[i] - leftFourHorzMean) / leftFourHorzSD))) + (.3 * (1+((leftFourIVB[i] - leftFourIVBMean) / leftFourIVBSD))) + (.1 * (1+((leftFourIVB[i] - leftFourBucketMean[leftFourBucket[i] - 24]) / leftFourBucketSD[leftFourBucket[i] - 24])))
  }
  return(stuffVect)
}

leftFourSeamStuff <- LeftFourStuff(accLeftFour)

leftFourSeamStuff

beep(3)
```

# Create a Stuff Value: Two-Seamers
```{r}
stuffVect <- vector()
rightTwoStuffFunction <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    stuffVect[i] <- (.5 * (1+ ((rightTwoAdjustVelo[i] - rightTwoVeloMean) / rightTwoVeloSD))) + (.25 * (1+((rightTwoHorz[i] - rightTwoHorzMean) / rightTwoHorzSD))) + (.25 * (1+(-(rightTwoIVB[i] - rightTwoIVBMean) / rightTwoIVBSD)))
  }
  return(stuffVect)
}

rightTwoStuff <- rightTwoStuffFunction(accRightTwo)


leftTwoStuffFunction <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    stuffVect[i] <- (.5 * (1+ ((leftTwoAdjustVelo[i] - leftTwoVeloMean) / leftTwoVeloSD))) + (.25 * (1+((leftTwoHorz[i] - leftTwoHorzMean) / leftTwoHorzSD))) + (.25 * (1+(-(leftTwoIVB[i] - leftTwoIVBMean) / leftTwoIVBSD)))
  }
  return(stuffVect)
}

leftTwoStuff <- leftTwoStuffFunction(accLeftTwo)

leftTwoStuff

leftTwoIVB[3] - leftTwoIVBMean
```



# Create a Stuff+ Value

```{r}
stuffPlusVect <- vector()

stuffPlusFunction <- function(x){
  for(i in 1:length(x)){
    stuffPlusVect[i] <- 100*(x[i])
  }
  return(stuffPlusVect)
}

rightFourStuffPlus <- stuffPlusFunction(rightFourSeamStuff)
rightTwoStuffPlus <- stuffPlusFunction(rightTwoStuff)
leftFourStuffPlus <- stuffPlusFunction(leftFourSeamStuff)
leftTwoStuffPlus <- stuffPlusFunction(leftTwoStuff)

```

# Export Stuff+ Values

```{r}
#Right Handed Four Seam
rightFourStuffPlusTable <- mapply(c, unique(accRightFour$Pitcher), rightFourTeam, rightFourLevel, rightFourCount, rightFourVelo, rightFourAdjustVelo, rightFourHorz, rightFourIVB, rightFourVeloPercentile, rightFourHorzPercentile, rightFourIVBPercentile, rightIVBBucketPercentile, rightFourStuffPlus, SIMPLIFY = FALSE)

write.csv(rightFourStuffPlusTable, "2022 NCAA Right Handed Four Seam Stuff+.csv")

#Right Handed Two Seam
rightTwoStuffPlusTable <- mapply(c, unique(accRightTwo$Pitcher), rightTwoTeam, rightTwoLevel, rightTwoCount, rightTwoVelo, rightTwoAdjustVelo, rightTwoHorz, rightTwoIVB, rightTwoVeloPercentile, rightTwoHorzPercentile, rightTwoIVBPercentileYes,  rightTwoStuffPlus, SIMPLIFY = FALSE)

write.csv(rightTwoStuffPlusTable, "2022 NCAA Right Handed Two Seam Stuff+.csv", row.names = FALSE)

#Left Handed Four Seam
leftFourStuffPlusTable <- mapply(c, unique(accLeftFour$Pitcher), leftFourTeam, leftFourLevel, leftFourCount, leftFourVelo, leftFourAdjustVelo, leftFourHorz, leftFourIVB, leftFourVeloPercentile, leftFourHorzPercentile, leftFourIVBPercentile, leftIVBBucketPercentile, leftFourStuffPlus, SIMPLIFY = FALSE)

write.csv(leftFourStuffPlusTable, "2022 NCAA Left Handed Four Seam Stuff+.csv")

#Left Handed Two Seam
leftTwoStuffPlusTable <- mapply(c, unique(accLeftTwo$Pitcher), leftTwoTeam, leftTwoLevel, leftTwoCount, leftTwoVelo, leftTwoAdjustVelo, leftTwoHorz, leftTwoIVB, leftTwoVeloPercentile, leftTwoHorzPercentile, leftTwoIVBPercentileYes,  leftTwoStuffPlus, SIMPLIFY = FALSE)

write.csv(leftTwoStuffPlusTable, "2022 NCAA Left Handed Two Seam Stuff+.csv", row.names = FALSE)

beep(3)
```