---
title: "wOBA Calculator"
author: "Alexander Harriman"
date: "2022-08-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file calculates college wOBA based on Trackman data. Equation last updated 2022

# Load Data
```{r, warning = FALSE}
library(readxl)
pitchers2022 <- read.csv("2022 NCAA Trackman with Command.csv")
pitchers2022 <- pitchers2022[is.na(pitchers2022$Pitcher) == FALSE,]
pitchers2022 <- pitchers2022[is.na(pitchers2022$TaggedPitchType) == FALSE,]
pitchers2022 <- pitchers2022[is.na(pitchers2022$PlayResult) == FALSE,]
pitchers2022 <- pitchers2022[is.na(pitchers2022$PitchCall) == FALSE,]
pitchers2022 <- pitchers2022[is.na(pitchers2022$KorBB) == FALSE,]
```

# Find the teams for each pitcher
```{r}
team <- vector()
teamFinder <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    base <- unique(x$Pitcher)[i]
    pitcher <- x[x$Pitcher == base,]
    team[i] <- unique(pitcher$PitcherTeam)[1]
  }
  return(team)
}

teams <- teamFinder(pitchers2022)
```

# Calculate Numerator

The numerator of wOBA requires unintentional walks and all forms of hits. For each pitcher, the number of each is calculated. Then, each weight is assigned and summed up.

```{r}
numerator <- vector()
sums <- 0
wOBA <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    base <- unique(x$Pitcher)[i]
    pitcher <- x[x$Pitcher == base,]
    for(a in 1:dim(pitcher)[1]){
      if(pitcher$PitchCall[a] == "HitByPitch"){
        sums <- sums + .728
      }
      else if(pitcher$PlayResult[a] == "Single"){
        sums <- sums + .883
      }
      else if(pitcher$PlayResult[a] == "Double"){
        sums <- sums + 1.238
      }
      else if(pitcher$PlayResult[a] == "Triple"){
        sums <- sums + 1.558
      }
      else if(pitcher$PlayResult[a] == "HomeRun"){
        sums <- sums + 1.979
      }
      else if(pitcher$KorBB[a] == "Walk"){
        sums <- sums + .699
      }
    }
    numerator[i] <- sums
    sums <- 0
  }
  return(numerator)
}

totalwOBA <- wOBA(pitchers2022)
```

# Calculate Denominator

The denominator is the sum of at bats, walks, sacrifices, and hit by pitches.

```{r}
denominator <- vector()
total <- 0

denom <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    base <- unique(x$Pitcher)[i]
    pitcher <- x[x$Pitcher == base,]
    for(a in 1:dim(pitcher)[1]){
      if(pitcher$PlayResult[a] != "Undefined" && pitcher$PlayResult[a] != "FieldersChoice" && pitcher$PlayResult[a] != "Error"){
        total <- total + 1
      }
      else if(pitcher$KorBB[a] == "Strikeout" | pitcher$KorBB[a] == "Walk"){
        total <- total + 1
      }
      else if(pitcher$PitchCall[a] == "HitByPitch"){
        total <- total + 1
      }
    }
    denominator[i] <- total
    total <- 0
  }
  return(denominator)
}

denominator <- denom(pitchers2022)
```

# Calculate wOBA

```{r}
finalwOBA <- totalwOBA / denominator

finalwOBA
```

# Export Results

```{r}
results <- mapply(c, unique(pitchers2022$Pitcher), teams, denominator, finalwOBA)
write.csv(results, "2022 NCAA Pitcher wOBA.csv")
```