---
title: "Top 100 Batter wOBA"
author: "Alexander Harriman"
date: "2022-11-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file finds the 100 batters with the highest wOBA and creates a subset of Trackman data that will be used to find metrics on them in a separate code file.

# Load Data

```{r}
pitchers2022 <- read.csv("Full 2022 Trackman.csv")
```

# Filter Out NA Values

```{r}
pitchers2022 <- pitchers2022[is.na(pitchers2022$Batter) == FALSE,]
pitchers2022 <- pitchers2022[is.na(pitchers2022$TaggedPitchType) == FALSE,]
pitchers2022 <- pitchers2022[is.na(pitchers2022$PlayResult) == FALSE,]
pitchers2022 <- pitchers2022[is.na(pitchers2022$PitchCall) == FALSE,]
pitchers2022 <- pitchers2022[is.na(pitchers2022$KorBB) == FALSE,]
pitchers2022 <- pitchers2022[pitchers2022$Level == "D1",]
```

# Filter Out Small AB Counts

Batters will only be included if they have a minimum of 150 at-bats. 

```{r}
#Find Batter List
batterAB <- 1

filteredData <- data.frame()

abCount <- function(){
  for(i in 1:length(batterList)){
    batter <- batterList[i]
    base <- pitchers2022[pitchers2022$Batter == batter,]
    if(length(base$PitchNo) < 2){
      next
    }
    for(j in 2:length(base$PitchNo)){
      if(base$PitchNo[j] - base$PitchNo[j-1] != 1){
        batterAB <- batterAB + 1
      }
    }
    if(batterAB >= 150){
      filteredData <<- rbind(filteredData, base)
    }
    batterAB <- 1
  }
  return(filteredData)
}

filteredPitcher <- abCount()

#The double arrow saves filteredData globally. It will be used in later code; if this is changed, then all instances of filteredData will need to be changed to filteredPitcher.
```

# Create New Batter List

Now that the data has been filtered down, a new batter list will be created. The length should be higher than 100 batters; if not, a lower at-bat limit must be created.

```{r}
filteredBatter <- unique(filteredData$Batter)
length(filteredBatter)
```

With 345 batters in the filtered data, we are ready to find wOBA.

# Calculate Batter wOBA

This is the same code for pitcher wOBA, but adjusted to each batter instead.

## Calculate Numerator

The numerator of wOBA requires unintentional walks and all forms of hits. For each batter, the number of each is calculated. Then, each weight is assigned and summed up.

```{r}
numerator <- vector()
sums <- 0
wOBA <- function(x){
  for(i in 1:length(filteredBatter)){
    base <- filteredBatter[i]
    batter <- x[x$Batter == base,]
    for(a in 1:dim(batter)[1]){
      if(batter$PitchCall[a] == "HitByPitch"){
        sums <- sums + .728
      }
      else if(batter$PlayResult[a] == "Single"){
        sums <- sums + .883
      }
      else if(batter$PlayResult[a] == "Double"){
        sums <- sums + 1.238
      }
      else if(batter$PlayResult[a] == "Triple"){
        sums <- sums + 1.558
      }
      else if(batter$PlayResult[a] == "HomeRun"){
        sums <- sums + 1.979
      }
      else if(batter$KorBB[a] == "Walk"){
        sums <- sums + .699
      }
    }
    numerator[i] <- sums
    sums <- 0
  }
  return(numerator)
}

totalwOBA <- wOBA(filteredData)
```

## Calculate Denominator

The denominator is the sum of at bats, walks, sacrifices, and hit by pitches.

```{r}
denominator <- vector()
total <- 0

denom <- function(x){
  for(i in 1:length(filteredBatter)){
    base <- filteredBatter[i]
    batter <- x[x$Batter == base,]
    for(a in 1:dim(batter)[1]){
      if(batter$PlayResult[a] != "Undefined" && batter$PlayResult[a] != "FieldersChoice" && batter$PlayResult[a] != "Error"){
        total <- total + 1
      }
      else if(batter$KorBB[a] == "Strikeout" | batter$KorBB[a] == "Walk"){
        total <- total + 1
      }
      else if(batter$PitchCall[a] == "HitByPitch"){
        total <- total + 1
      }
    }
    denominator[i] <- total
    total <- 0
  }
  return(denominator)
}

denominator <- denom(filteredData)
```

## Calculate wOBA

```{r}
finalwOBA <- totalwOBA / denominator

finalwOBA
```

# Find Top 100 wOBA

Now that all of our qualified batters have a wOBA value, only the top 100 are needed for analysis.

```{r}
#Creating a data.frame with batter name and wOBA
library(qpcR)

#Create a long list
wOBATable <- qpcR:::cbind.na(filteredBatter, finalwOBA)

#Convert to table
wOBATable <- as.data.frame(wOBATable)

#Ensuring that wOBA is a number, not a character variable
wOBATable$finalwOBA <- as.numeric(wOBATable$finalwOBA)

wOBATable
```

```{r}
#Ordering by top 100 wOBA
topDownwOBA <- wOBATable[order(-wOBATable$finalwOBA),]

topDownwOBA

#Now only taking the top 100 rows
top100wOBA <- topDownwOBA[1:100,]
top100wOBA
```

# Filter Data Down to Top 100

Now that we have identified the top 100 batters by wOBA, the final data filtering will begin.

```{r}
topData <- data.frame()

bestwOBA <- function(x){
  for(i in 1:length(filteredBatter)){
    batter <- filteredBatter[i]
    base <- x[x$Batter == batter,]
    if(batter %in% top100wOBA$filteredBatter){
      topData <<- rbind(topData, base)
    }
  }
  return(topData)
}

top100 <- bestwOBA(filteredData)
```

# Export Data

```{r}
write.csv(topData, "Top 100 Batters by wOBA.csv", row.names = FALSE)
```




