---
title: "Maximum Velocity Finder"
author: "Alexander Harriman"
date: '2022-07-14'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file finds the Whiff% for all pitchers on a given team, based on a given pitch.

```{r, warning = FALSE}
library(readxl)
pitchers2022o <- read_excel("Reduced 2022 Trackman (5-4).xlsx")
pitchers2022n <- read_excel("Reduced 2022 Trackman (5-5 to 5-21).xlsx")

pitchers2022 <- rbind(pitchers2022n, pitchers2022o)

beep(3)
```


# Filter down to team(s) studied

If looking at a single conference instead, the code can be adjusted to $League == "Conference".
```{r}
#Clemson
clemson <- pitchers2022[pitchers2022$PitcherTeam == "CLE_TIG",]
clemson <- clemson[is.na(clemson$RelSpeed) == FALSE,]
names <- unique(clemson$Pitcher)
```

# Filter down to pitcher(s) studied

Only use if looking at specific pitchers on the team(s).
```{r}
ammons <- clemson[clemson$Pitcher == "Ammons, Ryan",]
```

# Filter down to pitch(es) studies

Only used if looking at a specific pitch type(s) and not all pitches.

```{r}
#Ammons is a lefty, change based on handedness and pitch type
ammonsSlider <- ammons[ammons$`Tilt Coordinate` >= .0729 & ammons$`Tilt Coordinate` <= .1667,]
```

# Find Swinging Strikes

```{r}
swing <- vector()

strike <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    base <- unique(x$Pitcher)[i]
    pitcher <- x[x$Pitcher == base,]
    swingStrike <- pitcher[pitcher$PitchCall == "StrikeSwinging",]
    swing[i] <- dim(swingStrike)[1]
  }
  return(swing)
}

KAmmons <- strike(ammonsSlider)
KAmmons
```


# Find Number of Pitches
```{r}
count <- vector()
pitchCount <- function(x){
  for(i in 1:length(unique(x$Pitcher))){
    number <- 0
    base <- unique(x$Pitcher)[i]
    pitcher <- x[x$Pitcher == base,]
    for(a in 1:length(pitcher$PitchNo)){
      if(is.na(pitcher$PitchCall[a]) == TRUE){
        number <- number
      }
      else if(pitcher$PitchCall[a] == "StrikeSwinging"){
        number <- number + 1
      }
      else if(pitcher$PitchCall[a] == "FoulBall"){
        number <- number + 1
      }
      else if(pitcher$PitchCall[a] == "InPlay"){
        number <- number + 1
      }
      else{
        number <- number
      }
    }
    count[i] <- number
  }
  return(count)
}



countAmmons <- pitchCount(ammonsSlider)
```




# Export Names and Values

```{r}
library(dplyr)
maximum <- mapply(c, names, clemsonMax)
write.csv(maximum, "Clemson Maximum Velocity by Pitcher 2022.csv")
```