---
title: "Minor League Data Gathering"
author: "Alexander Harriman"
date: "2024-05-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



This code creates leaderboards for a specific MiLB level and puts the information into a dashboard. 

The leaderboards are as follows (up to V1.0):

* Best pitchers by WHIP for a season, including IP, wOBA against, SO/BB/HR per 9
* Best single outings by Tom Tango game score, including all parts of computation
* Bet hitters by OPS+, including OBP, SLG, wOBA, and other metrics
* Best single outings by wOBA, including other hitter metrics

The goal is to show up-to-date best performances across the season and for individual games.


Data is up-to-date for 2024 season. See Minor League Data Gathering.Rmd for data gathering.



# Data Load State

```{r}
allData <- read.csv('2024 Full MiLB May 13.csv')
allData <- allData[-1]
```



```{r}
library(dplyr)
library(stringr)
library(tidyr)
library(weights)
library(beepr)
```




# Pitcher Results 

This section will gather all of the pitcher information across the season.


## Season Performance: Best Pitchers by WHIP


At One Level of MiLB Only

```{r}
pitcherSeason <- allData |>
  distinct(.keep_all = TRUE) |>
  filter(result.eventType != 'field_error') |>
  group_by(matchup.pitcher.id, matchup.pitcher.fullName, home_level_name, fielding_team) |>
  summarise(IP = sum(case_when(
    details.isOut == TRUE & result.eventType %in% c('double_play', 'grounded_into_double_play', 
                                                    'sac_fly_double_play', 'strikeout_double_play') ~ 2,
    details.isOut == TRUE & !result.eventType %in% c('double_play', 'grounded_into_double_play', 
                                                    'sac_fly_double_play', 'strikeout_double_play') ~ 1,
    details.isOut == FALSE ~ 0
  ), na.rm = TRUE) / 3,
  SO = sum(ifelse(count.strikes.start == 3, 1, 0), na.rm = TRUE),
  BB = sum(ifelse(count.balls.start == 4, 1, 0), na.rm = TRUE),
  HR = sum(ifelse(result.eventType == 'home_run' & details.call.description == 'In play, run(s)', 1, 0),
           na.rm = TRUE),
  WHIPNum = sum(ifelse(result.eventType %in% c('walk', 'single', 'double', 'triple', 'home_run') &
                         (count.balls.start == 4 | 
                            details.call.description %in% c('In play, no out', 'In play, run(s)')), 1, 0), 
                na.rm = TRUE),
  wOBANum = sum(case_when(
           result.eventType == 'walk' ~ 0.697,
           result.eventType == 'hit_by_pitch' ~ 0.729,
           result.eventType == 'single' ~ 0.896,
           result.eventType == 'double' ~ 1.280,
           result.eventType == 'triple' ~ 1.626,
           result.eventType == 'home_run' ~ 2.103,
           .default = 0
         ), na.rm = TRUE),
  wOBADenom = length(unique(paste(startTime,matchup.batter.fullName))) - 
           sum(ifelse(result.eventType == 'intent_walk', 1, 0), na.rm = TRUE)) |>
  mutate(WHIP = WHIPNum / IP,
         wOBA = wOBANum / wOBADenom,
         SO9 = (SO / IP) * 9,
         BB9 = (BB / IP) * 9,
         HR9 = (HR / IP) * 9,
         IPDecimal = IP - floor(IP),
         IPWhole = floor(IP)) 


cleanSeason <- pitcherSeason |>
  select(matchup.pitcher.fullName, fielding_team, IPWhole, IPDecimal, SO9, BB9, HR9, wOBA, WHIP) |>
  mutate(IP = case_when(
    round(IPDecimal, 1) == 0.3 ~ IPWhole + 0.1,
    round(IPDecimal, 1) == 0.7 ~ IPWhole + 0.2,
    .default = IPWhole
  ),
  SO9 = round(SO9, 2),
  BB9 = round(BB9, 2),
  HR9 = round(HR9, 2),
  WHIP = round(WHIP, 2),
  wOBA = rd(wOBA, digits = 3)) |>
  select(matchup.pitcher.id, matchup.pitcher.fullName, fielding_team, IP, SO9, BB9, HR9, wOBA, WHIP)

cleanSeason
```

### Top 10 Pitchers by WHIP

Pitchers must have thrown an above-average number of innings to qualify.

```{r}
averageChecker <- distinctData |>
  group_by(matchup.pitcher.id) |>
  summarise(IP = sum(case_when(
    details.isOut == TRUE & result.eventType %in% c('double_play', 'grounded_into_double_play', 
                                                    'sac_fly_double_play', 'strikeout_double_play') ~ 2,
    details.isOut == TRUE & !result.eventType %in% c('double_play', 'grounded_into_double_play', 
                                                    'sac_fly_double_play', 'strikeout_double_play') ~ 1,
    details.isOut == FALSE ~ 0
  ), na.rm = TRUE) / 3)

averageIP <- mean(averageChecker$IP)
```


```{r}
qualifiedPitchers <- cleanSeason |>
  filter(IP > floor(averageIP)) |>
  ungroup() |>
  mutate(WHIPRank = rank(WHIP, ties.method = 'min'))

topPitchers <- qualifiedPitchers |>
  filter(WHIPRank <= 10) |>
  arrange(WHIP) |>
  select(matchup.pitcher.fullName, fielding_team, home_level_name, IP, SO9, BB9, HR9, wOBA, WHIP)

colnames(topPitchers) <- c('Name', 'Team', 'Level', 'IP', 'K/9', 'BB/9', 'HR/9', 'wOBA', 'WHIP')

topPitchers
```


## Add Organization Info and Logos

```{r}
library(mlbplotR)

topPitchersOrg <- topPitchers |>
  mutate(Org = case_when(
    Team %in% c('Norfolk Tides', 'Bowie Baysox', 'Aberdeen IronBirds', 'Delmarva Shorebirds') ~ 'BAL',
    Team %in% c('Worcester Red Sox', 'Portland Sea Dogs', 'Greenville Drive', 'Salem Red Sox') ~ 'BOS',
    Team %in% c('Scranton/Wilkes-Barre RailRiders', 'Somerset Patriots', 'Hudson Valley Renegades', 'Tampa Tarpons') ~ 'NYY',
    Team %in% c('Durham Bulls', 'Montgomery Biscuits', 'Bowling Green Hot Rods', 'Charleston RiverDogs') ~ 'TB',
    Team %in% c('Buffalo Bisons', 'New Hampshire Fisher Cats', 'Vancouver Canadians', 'Dunedin Blue Jays') ~ 'TOR',
    Team %in% c('Charlotte Knights', 'Birmingham Barons', 'Winston-Salem Dash', 'Kannapolis Cannon Ballers') ~ 'CWS',
    Team %in% c('Columbus Clippers', 'Akron RubberDucks', 'Lake County Captains', 'Lynchburg Hillcats') ~ 'CLE',
    Team %in% c('Toledo Mud Hens', 'Erie SeaWolves', 'West Michigan Whitecaps', 'Lakeland Flying Tigers') ~ 'DET',
    Team %in% c('Omaha Storm Chasers', 'Northwest Arkansas Naturals', 'Quad Cities River Bandits', 'Columbia Fireflies') ~ 'KC',
    Team %in% c('St. Paul Saints', 'Wichita Wind Surge', 'Cedar Rapids Kernels', 'Fort Myers Mighty Mussels') ~ 'MIN',
    Team %in% c('Sugar Land Space Cowboys', 'Corpus Christi Hooks', 'Asheville Tourists', 'Fayetteville Woodpeckers') ~ 'HOU',
    Team %in% c('Salt Lake Bees', 'Rocket City Trash Pandas', 'Tri-City Dust Devils', 'Inland Empire 66ers') ~ 'LAA',
    Team %in% c('Las Vegas Aviators', 'Midland RockHounds', 'Lansing Lugnuts', 'Stockton Ports') ~ 'OAK',
    Team %in% c('Tacoma Rainiers', 'Arkansas Travelers', 'Everett AquaSox', 'Modesto Nuts') ~ 'SEA',
    Team %in% c('Round Rock Express', 'Frisco RoughRiders', 'Hickory Crawdads', 'Down East Wood Ducks') ~ 'TEX',
    Team %in% c('Gwinnett Stripers', 'Mississippi Braves', 'Rome Emperors', 'Augusta GreenJackets') ~ 'ATL',
    Team %in% c('Jacksonville Jumbo Shrimp', 'Pensacola Blue Wahoos', 'Beloit Sky Carp', 'Jupiter Hammerheads') ~ 'MIA',
    Team %in% c('Syracuse Mets', 'Binghamton Rumble Ponies', 'Brooklyn Cyclones', 'St. Lucie Mets') ~ 'NYM',
    Team %in% c('Lehigh Valley IronPigs', 'Reading Fightin Phils', 'Jersey Shore BlueClaws', 'Clearwater Threshers') ~ 'PHI',
    Team %in% c('Rochester Red Wings', 'Harrisburg Senators', 'Wilmington Blue Rocks', 'Fredericksburg Nationals') ~ 'WSH',
    Team %in% c('Iowa Cubs', 'Tennessee Smokies', 'South Bend Cubs', 'Myrtle Beach Pelicans') ~ 'CHC',
    Team %in% c('Louisville Bats', 'Chattanooga Lookouts', 'Dayton Dragons', 'Daytona Tortugas') ~ 'CIN',
    Team %in% c('Nashville Sounds', 'Biloxi Shuckers', 'Wisconsin Timber Rattlers', 'Carolina Mudcats') ~ 'MIL',
    Team %in% c('Indianapolis Indians', 'Altoona Curve', 'Greensboro Grasshoppers', 'Bradenton Marauders') ~ 'PIT',
    Team %in% c('Memphis Redbirds', 'Springfield Cardinals', 'Peoria Chiefs', 'Palm Beach Cardinals') ~ 'STL',
    Team %in% c('Reno Aces', 'Amarillo Sod Poodles', 'Hillsboro Hops', 'Visalia Rawhide') ~ 'AZ',
    Team %in% c('Albuquerque Isotopes', 'Hartford Yard Goats', 'Spokane Indians', 'Fresno Grizzlies') ~ 'COL',
    Team %in% c('Oklahoma City Baseball Club', 'Tulsa Drillers', 'Great Lakes Loons', 'Rancho Cucamonga Quakes') ~ 'LAD',
    Team %in% c('El Paso Chihuahuas', 'San Antonio Missions', 'Fort Wayne TinCaps', 'Lake Elsinore Storm') ~ 'SD',
    Team %in% c('Sacramento River Cats', 'Richmond Flying Squirrels', 'Eugene Emeralds', 'San Jose Giants') ~ 'SF',
    .default = 'MLB'
  ) 
  ) |>
    select(Name, Team, Level, Org, IP, `K/9`, `BB/9`, `HR/9`, wOBA, WHIP)

topPitchersOrg
```



```{r}
library(gt)
seasonBestPitchers <- topPitchersOrg |>
  gt() |>
  tab_header(title = md('**Top MiLB Pitchers by WHIP, 2024**')) |>
  cols_align(align = 'center', columns = c('Team', 'Level', 'Org', 'IP', 'K/9', 'BB/9', 'HR/9', 'wOBA', 'WHIP')) |>
  opt_stylize(style = 5, color = 'red') |>
  cols_width(Name ~ px(175), Team ~ px(175), Level ~ px(90), Org ~ px(90), everything() ~ px(65)) |>
  tab_footnote(footnote = 'Last Updated May 13th') |>
  gt_fmt_mlb_logo(columns = 'Org') |>
  tab_options(
    table.font.size = px(16)
  )

seasonBestPitchers
beep(3)
```



## Game Performance: Best Pitchers by Tom Tango Game Score

Formula for the Tango Game Score:

• Start with 40 points
• Add 2 points for each out recorded (or 6 points per inning)
• Add 1 additional point for every strikeout
• Remove 2 points for each walk allowed
• Remove 2 points for each hit allowed
• Remove 3 points for each run allowed (earned or unearned)
• Remove 6 additional points for each home run allowed


```{r}

```



## Season Performance: Best Hitters by OPS+

### Finding Season Averages


```{r}

```


### Individual Calculations


```{r}

```



## Game Performance: Best Hitters by Number of Bases


```{r}

```

