---
title: "Minor League Data Gathering"
author: "Alexander Harriman"
date: "2024-05-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



This code creates leaderboards for a specific MiLB level and puts the information into a dashboard. 

The leaderboards are as follows (up to V1.0):

* Best pitchers by WHIP for a season, including IP, wOBA against, SO/BB/HR per 9
* Best single outings by Tom Tango game score, including all parts of computation
* Bet hitters by OPS+, including OBP, SLG, wOBA, and other metrics
* Best single outings by wOBA, including other hitter metrics

The goal is to show up-to-date best performances across the season and for individual games.


Data is up-to-date for 2024 season. See code below (or Minor League Data Gathering.Rmd) for data gathering.


# Finding Game PKs

ID for each level:

* MLB - 1
* AAA - 11
* AA - 12
* A+ - 13
* A - 14
* Rookie Ball - 16

```{r, warning=FALSE}
library(baseballr)
library(lubridate)
library(dplyr)
library(data.table)
library(beepr)

dateSequence <- seq(ymd('2024-04-05'), Sys.Date(), by = '1 day')

# AA, High-A, and Low-A start on April 5
## AAA start on March 29
### MLB Korea Series start on March 20, rest of league on March 28
#### Complex leagues start on May 4 and end July 25

gamePKs <- c()

dataAcquisition <- function(){
  for(i in 1:length(dateSequence)){
  allPK <- mlb_game_pks(date = dateSequence[i], level_ids = 13)$game_pk #Adjust level_id as needed, see above
  gamePKs <- append(gamePKs, allPK)
  }
  return(gamePKs)
}

pkList <- dataAcquisition()
```

# Getting PBP Data from Game PKs


```{r, warning = FALSE, error = TRUE}
pitchData <- list()

combiningData <- function(){
  for(i in 1:length(pkList)){
    skipFile <- FALSE
    tryCatch(
    {
      gameData <- mlb_pbp(game_pk = pkList[i]) |>
        filter(game_date != TRUE)
      pitchData[[i]] <- gameData
    },
    error = function(e){
      skipFile == TRUE
      print(gameData$game_pk[1])
    }
    )
  if(skipFile == TRUE){
    next
  }
  }
  return(pitchData)
}

dataList <- combiningData()

allData <- rbindlist(dataList, fill = TRUE)

dim(allData)[1]
```


# Data Save State

```{r}
write.csv(allData, '2024 High A May 7th.csv')
beep(3)
```

Saved Data List:

2024 Low A: 2:38 PM May 3rd
2024 High A: 11:00 AM May 7th


Approximate run time from start to save state: 20 mins.

# Data Load State

If a CSV needs to be loaded to save time. DO NOT INCLUDE IN CODE POSTED TO GITHUB

```{r}
lowAData <- read.csv('2024 High A May 6th.csv')
```



# Pitcher Results 

This section will gather all of the pitcher information across the season.


## Season Performance: Best Pitchers by WHIP


```{r}
library(dplyr)
library(stringr)
library(tidyr)
library(weights)
pitcherSeason <- allData |>
  distinct(.keep_all = TRUE) |>
  group_by(matchup.pitcher.fullName, fielding_team) |>
  summarise(IP = sum(case_when(
    details.isOut == TRUE & result.eventType %in% c('double_play', 'grounded_into_double_play', 
                                                    'sac_fly_double_play', 'strikeout_double_play') ~ 2,
    details.isOut == TRUE & !result.eventType %in% c('double_play', 'grounded_into_double_play', 
                                                    'sac_fly_double_play', 'strikeout_double_play') ~ 1,
    details.isOut == FALSE ~ 0
  ), na.rm = TRUE) / 3,
  SO = sum(ifelse(count.strikes.start == 3, 1, 0), na.rm = TRUE),
  BB = sum(ifelse(count.balls.start == 4, 1, 0), na.rm = TRUE),
  HR = sum(ifelse(result.eventType == 'home_run' & details.call.description == 'In play, run(s)', 1, 0),
           na.rm = TRUE),
  WHIPNum = sum(ifelse(result.eventType %in% c('walk', 'single', 'double', 'triple', 'home_run') &
                         (count.balls.start == 4 | 
                            details.call.description %in% c('In play, no out', 'In play, run(s)')), 1, 0), 
                na.rm = TRUE),
  wOBANum = sum(case_when(
           result.eventType == 'walk' ~ 0.697,
           result.eventType == 'hit_by_pitch' ~ 0.729,
           result.eventType == 'single' ~ 0.896,
           result.eventType == 'double' ~ 1.280,
           result.eventType == 'triple' ~ 1.626,
           result.eventType == 'home_run' ~ 2.103,
           .default = 0
         ), na.rm = TRUE),
  wOBADenom = length(unique(paste(startTime,matchup.batter.fullName))) - 
           sum(ifelse(result.eventType == 'intent_walk', 1, 0), na.rm = TRUE)) |>
  mutate(WHIP = WHIPNum / IP,
         wOBA = wOBANum / wOBADenom,
         SO9 = (SO / IP) * 9,
         BB9 = (BB / IP) * 9,
         HR9 = (HR / IP) * 9,
         IPDecimal = IP - floor(IP),
         IPWhole = floor(IP)) 


cleanSeason <- pitcherSeason |>
  select(matchup.pitcher.fullName, fielding_team, IPWhole, IPDecimal, SO9, BB9, HR9, wOBA, WHIP) |>
  mutate(IP = case_when(
    round(IPDecimal, 1) == 0.3 ~ IPWhole + 0.1,
    round(IPDecimal, 1) == 0.7 ~ IPWhole + 0.2,
    .default = IPWhole
  ),
  SO9 = round(SO9, 2),
  BB9 = round(BB9, 2),
  HR9 = round(HR9, 2),
  WHIP = round(WHIP, 2),
  wOBA = rd(wOBA, digits = 3)) |>
  select(matchup.pitcher.fullName, fielding_team, IP, SO9, BB9, HR9, wOBA, WHIP)

cleanSeason
```

### Top 10 Pitchers by WHIP

Pitchers must have thrown an above-average number of innings to qualify.


```{r}
averageIP <- mean(pitcherSeason$IP)

qualifiedPitchers <- cleanSeason |>
  filter(IP > floor(averageIP)) |>
  ungroup() |>
  mutate(WHIPRank = rank(WHIP, ties.method = 'min'))

topPitchers <- qualifiedPitchers |>
  filter(WHIPRank <= 10) |>
  arrange(WHIP) |>
  select(matchup.pitcher.fullName, fielding_team, IP, SO9, BB9, HR9, wOBA, WHIP)

colnames(topPitchers) <- c('Name', 'Team', 'IP', 'K/9', 'BB/9', 'HR/9', 'wOBA', 'WHIP')

topPitchers
```



```{r}
library(gt)
seasonBestPitchers <- topPitchers |>
  gt() |>
  tab_header(title = md('**Top High-A Pitchers by WHIP, 2024**')) |>
  cols_align(align = 'center', columns = c('IP', 'K/9', 'BB/9', 'HR/9', 'wOBA', 'WHIP')) |>
  opt_stylize(style = 5, color = 'red') |>
  cols_width(Name ~ px(150), Team ~ px(225), everything() ~ px(75)) |>
  tab_footnote(footnote = 'Last Updated May 7th')

seasonBestPitchers
```



## Game Performance: Best Pitchers by Tom Tango Game Score

Formula for the Tango Game Score:

• Start with 40 points
• Add 2 points for each out recorded (or 6 points per inning)
• Add 1 additional point for every strikeout
• Remove 2 points for each walk allowed
• Remove 2 points for each hit allowed
• Remove 3 points for each run allowed (earned or unearned)
• Remove 6 additional points for each home run allowed


```{r}

```



## Season Performance: Best Hitters by OPS+

### Finding Season Averages


```{r}

```


### Individual Calculations


```{r}

```



## Game Performance: Best Hitters by Number of Bases


```{r}

```

