---
title: "Minor League Data Gathering"
author: "Alexander Harriman"
date: "2024-05-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



This code creates leaderboards for a specific MiLB level and puts the information into a dashboard. 

The leaderboards are as follows (up to V1.0):

* Best pitchers by WHIP for a season, including IP, wOBA against, SO/BB/HR per 9
* Best single outings by Tom Tango game score, including all parts of computation
* Bet hitters by OPS+, including OBP, SLG, wOBA, and other metrics
* Best single outings by wOBA, including other hitter metrics

The goal is to show up-to-date best performances across the season and for individual games.


Data is up-to-date for 2024 season. See Minor League Data Gathering.Rmd for data gathering.



# Data Load State

```{r}
allData <- read.csv('2024 Full MiLB May 22.csv')

#Check for how many row index columns exist, and remove that number from this imported file
##One per write.csv used to create the csv file loaded above

allData <- allData[-1]
allData <- allData[-1]
```


```{r}
library(dplyr)
library(stringr)
library(tidyr)
library(weights)
library(beepr)
```




# Pitcher Results 

This section will gather all of the pitcher information across the season.


## Season Performance: Best Pitchers by WHIP


At One Level of MiLB Only

```{r}
pitcherSeason <- allData |>
  distinct(paste(game_pk, startTime, endTime), .keep_all = TRUE) |>
  filter(result.eventType != 'field_error') |>
  group_by(matchup.pitcher.id, matchup.pitcher.fullName, home_level_name, fielding_team) |>
  summarise(IP = sum(case_when(
    details.isOut == TRUE & result.eventType %in% c('double_play', 'grounded_into_double_play', 
                                                    'sac_fly_double_play', 'strikeout_double_play') ~ 2,
    details.isOut == TRUE & !result.eventType %in% c('double_play', 'grounded_into_double_play', 
                                                    'sac_fly_double_play', 'strikeout_double_play') ~ 1,
    details.isOut == FALSE ~ 0
  ), na.rm = TRUE) / 3,
  SO = sum(ifelse(count.strikes.start == 3, 1, 0), na.rm = TRUE),
  BB = sum(ifelse(count.balls.start == 4, 1, 0), na.rm = TRUE),
  HR = sum(ifelse(result.eventType == 'home_run' & details.call.description == 'In play, run(s)', 1, 0),
           na.rm = TRUE),
  WHIPNum = sum(ifelse(result.eventType %in% c('walk', 'single', 'double', 'triple', 'home_run') &
                         (count.balls.start == 4 | 
                            details.call.description %in% c('In play, no out', 'In play, run(s)')), 1, 0), 
                na.rm = TRUE),
  wOBANum = sum(case_when(
           result.eventType == 'walk' ~ 0.697,
           result.eventType == 'hit_by_pitch' ~ 0.729,
           result.eventType == 'single' ~ 0.896,
           result.eventType == 'double' ~ 1.280,
           result.eventType == 'triple' ~ 1.626,
           result.eventType == 'home_run' ~ 2.103,
           .default = 0
         ), na.rm = TRUE),
  wOBADenom = length(unique(paste(startTime,matchup.batter.fullName))) - 
           sum(ifelse(result.eventType == 'intent_walk', 1, 0), na.rm = TRUE)) |>
  mutate(WHIP = WHIPNum / IP,
         wOBA = wOBANum / wOBADenom,
         SO9 = (SO / IP) * 9,
         BB9 = (BB / IP) * 9,
         HR9 = (HR / IP) * 9,
         IPDecimal = IP - floor(IP),
         IPWhole = floor(IP)) 


cleanSeason <- pitcherSeason |>
  select(matchup.pitcher.fullName, fielding_team, IPWhole, IPDecimal, SO9, BB9, HR9, wOBA, WHIP) |>
  mutate(IP = case_when(
    round(IPDecimal, 1) == 0.3 ~ IPWhole + 0.1,
    round(IPDecimal, 1) == 0.7 ~ IPWhole + 0.2,
    .default = IPWhole
  ),
  SO9 = round(SO9, 2),
  BB9 = round(BB9, 2),
  HR9 = round(HR9, 2),
  wOBA = rd(wOBA, digits = 3)) |>
  select(matchup.pitcher.id, matchup.pitcher.fullName, fielding_team, IP, SO9, BB9, HR9, wOBA, WHIP)

cleanSeason
```

### Top 10 Pitchers by WHIP

Pitchers must have thrown an above-average number of innings to qualify.

```{r}
averageChecker <- allData |>
  group_by(matchup.pitcher.id) |>
  summarise(IP = sum(case_when(
    details.isOut == TRUE & result.eventType %in% c('double_play', 'grounded_into_double_play', 
                                                    'sac_fly_double_play', 'strikeout_double_play') ~ 2,
    details.isOut == TRUE & !result.eventType %in% c('double_play', 'grounded_into_double_play', 
                                                    'sac_fly_double_play', 'strikeout_double_play') ~ 1,
    details.isOut == FALSE ~ 0
  ), na.rm = TRUE) / 3)

averageIP <- mean(averageChecker$IP)
```


```{r}
qualifiedPitchers <- cleanSeason |>
  filter(IP > floor(averageIP)) |>
  ungroup() 
```



## Add Organization Info and Logos

```{r}
library(mlbplotR)

topPitchersOrg <- qualifiedPitchers |>
  mutate(Org = case_when(
    Team %in% c('Norfolk Tides', 'Bowie Baysox', 'Aberdeen IronBirds', 'Delmarva Shorebirds') ~ 'BAL',
    Team %in% c('Worcester Red Sox', 'Portland Sea Dogs', 'Greenville Drive', 'Salem Red Sox') ~ 'BOS',
    Team %in% c('Scranton/Wilkes-Barre RailRiders', 'Somerset Patriots', 'Hudson Valley Renegades', 'Tampa Tarpons') ~ 'NYY',
    Team %in% c('Durham Bulls', 'Montgomery Biscuits', 'Bowling Green Hot Rods', 'Charleston RiverDogs') ~ 'TB',
    Team %in% c('Buffalo Bisons', 'New Hampshire Fisher Cats', 'Vancouver Canadians', 'Dunedin Blue Jays') ~ 'TOR',
    Team %in% c('Charlotte Knights', 'Birmingham Barons', 'Winston-Salem Dash', 'Kannapolis Cannon Ballers') ~ 'CWS',
    Team %in% c('Columbus Clippers', 'Akron RubberDucks', 'Lake County Captains', 'Lynchburg Hillcats') ~ 'CLE',
    Team %in% c('Toledo Mud Hens', 'Erie SeaWolves', 'West Michigan Whitecaps', 'Lakeland Flying Tigers') ~ 'DET',
    Team %in% c('Omaha Storm Chasers', 'Northwest Arkansas Naturals', 'Quad Cities River Bandits', 'Columbia Fireflies') ~ 'KC',
    Team %in% c('St. Paul Saints', 'Wichita Wind Surge', 'Cedar Rapids Kernels', 'Fort Myers Mighty Mussels') ~ 'MIN',
    Team %in% c('Sugar Land Space Cowboys', 'Corpus Christi Hooks', 'Asheville Tourists', 'Fayetteville Woodpeckers') ~ 'HOU',
    Team %in% c('Salt Lake Bees', 'Rocket City Trash Pandas', 'Tri-City Dust Devils', 'Inland Empire 66ers') ~ 'LAA',
    Team %in% c('Las Vegas Aviators', 'Midland RockHounds', 'Lansing Lugnuts', 'Stockton Ports') ~ 'OAK',
    Team %in% c('Tacoma Rainiers', 'Arkansas Travelers', 'Everett AquaSox', 'Modesto Nuts') ~ 'SEA',
    Team %in% c('Round Rock Express', 'Frisco RoughRiders', 'Hickory Crawdads', 'Down East Wood Ducks') ~ 'TEX',
    Team %in% c('Gwinnett Stripers', 'Mississippi Braves', 'Rome Emperors', 'Augusta GreenJackets') ~ 'ATL',
    Team %in% c('Jacksonville Jumbo Shrimp', 'Pensacola Blue Wahoos', 'Beloit Sky Carp', 'Jupiter Hammerheads') ~ 'MIA',
    Team %in% c('Syracuse Mets', 'Binghamton Rumble Ponies', 'Brooklyn Cyclones', 'St. Lucie Mets') ~ 'NYM',
    Team %in% c('Lehigh Valley IronPigs', 'Reading Fightin Phils', 'Jersey Shore BlueClaws', 'Clearwater Threshers') ~ 'PHI',
    Team %in% c('Rochester Red Wings', 'Harrisburg Senators', 'Wilmington Blue Rocks', 'Fredericksburg Nationals') ~ 'WSH',
    Team %in% c('Iowa Cubs', 'Tennessee Smokies', 'South Bend Cubs', 'Myrtle Beach Pelicans') ~ 'CHC',
    Team %in% c('Louisville Bats', 'Chattanooga Lookouts', 'Dayton Dragons', 'Daytona Tortugas') ~ 'CIN',
    Team %in% c('Nashville Sounds', 'Biloxi Shuckers', 'Wisconsin Timber Rattlers', 'Carolina Mudcats') ~ 'MIL',
    Team %in% c('Indianapolis Indians', 'Altoona Curve', 'Greensboro Grasshoppers', 'Bradenton Marauders') ~ 'PIT',
    Team %in% c('Memphis Redbirds', 'Springfield Cardinals', 'Peoria Chiefs', 'Palm Beach Cardinals') ~ 'STL',
    Team %in% c('Reno Aces', 'Amarillo Sod Poodles', 'Hillsboro Hops', 'Visalia Rawhide') ~ 'AZ',
    Team %in% c('Albuquerque Isotopes', 'Hartford Yard Goats', 'Spokane Indians', 'Fresno Grizzlies') ~ 'COL',
    Team %in% c('Oklahoma City Baseball Club', 'Tulsa Drillers', 'Great Lakes Loons', 'Rancho Cucamonga Quakes') ~ 'LAD',
    Team %in% c('El Paso Chihuahuas', 'San Antonio Missions', 'Fort Wayne TinCaps', 'Lake Elsinore Storm') ~ 'SD',
    Team %in% c('Sacramento River Cats', 'Richmond Flying Squirrels', 'Eugene Emeralds', 'San Jose Giants') ~ 'SF',
    .default = 'MLB'
  ) 
  ) |>
    select(Name, Team, Level, Org, IP, `K/9`, `BB/9`, `HR/9`, wOBA, WHIP)

topPitchersOrg
```


```{r}
topPitchers <- qualifiedPitchers |>
  mutate(WHIPRank = rank(WHIP, ties.method = 'min'))
  filter(WHIPRank <= 10) |>
  arrange(WHIP) |>
  select(matchup.pitcher.fullName, fielding_team, home_level_name, IP, SO9, BB9, HR9, wOBA, WHIP) |>
  mutate(WHIP = round(WHIP,2))

colnames(topPitchers) <- c('Name', 'Team', 'Level', 'IP', 'K/9', 'BB/9', 'HR/9', 'wOBA', 'WHIP')

topPitchers
```


```{r}
pitcherSeasonExport <- topPitchersOrg |>
  group_by(Team, .keep_all = TRUE) |>
  mutate(ScoreRank = rank(WHIP, ties.method = 'min')) |>
  filter(ScoreRank <= 10) 

pitcherSeasonExport
```


```{r}
write.csv(pitcherGameExport, 'MiLB Best Games by Team.csv')
```



```{r}
library(gt)
seasonBestPitchers <- topPitchersOrg |>
  gt() |>
  tab_header(title = md('**Top MiLB Pitchers by WHIP, 2024**')) |>
  cols_align(align = 'center', columns = c('Team', 'Level', 'Org', 'IP', 'K/9', 'BB/9', 'HR/9', 'wOBA', 'WHIP')) |>
  opt_stylize(style = 5, color = 'red') |>
  cols_width(Name ~ px(200), Team ~ px(200), Level ~ px(80), everything() ~ px(60)) |>
  tab_footnote(footnote = 'Last Updated May 13th') |>
  gt_fmt_mlb_logo(columns = 'Org') |>
  tab_options(
    table.font.size = px(16)
  )

seasonBestPitchers
```



## Game Performance: Best Pitchers by Tom Tango Game Score

Formula for the Tango Game Score:

• Start with 40 points
• Add 2 points for each out recorded (or 6 points per inning)
• Add 1 additional point for every strikeout
• Remove 2 points for each walk allowed
• Remove 2 points for each hit allowed
• Remove 3 points for each run allowed (earned or unearned)
• Remove 6 additional points for each home run allowed


```{r}
pitcherGame <- allData |>
  distinct(paste(game_pk, startTime, endTime, matchup.pitcher.fullName), .keep_all = TRUE) |>
  filter(result.eventType != 'field_error') |>
  mutate(runs_scored = case_when(
    lead(game_pk) != game_pk ~ 0,
    about.halfInning == 'top' ~ result.awayScore - lead(result.awayScore), 
    about.halfInning == 'bottom' ~ result.homeScore - lead(result.homeScore),
    .default = 0
  )) |>
  group_by(game_pk, matchup.pitcher.id, matchup.pitcher.fullName, fielding_team, game_date, home_level_name) |>
  summarise(Outs = sum(case_when(
    details.isOut == TRUE & result.eventType %in% c('double_play', 'grounded_into_double_play', 
                                                    'sac_fly_double_play', 'strikeout_double_play') ~ 2,
    details.isOut == TRUE & !result.eventType %in% c('double_play', 'grounded_into_double_play', 
                                                    'sac_fly_double_play', 'strikeout_double_play') ~ 1,
    details.isOut == FALSE ~ 0
  ), na.rm = TRUE),
  SO = sum(ifelse(count.strikes.start == 3, 1, 0), na.rm = TRUE),
  BB = sum(ifelse(count.balls.start == 4, 1, 0), na.rm = TRUE),
  HR = sum(ifelse(result.eventType == 'home_run' & details.call.description == 'In play, run(s)', 1, 0),
           na.rm = TRUE),
  H = sum(ifelse(
    result.eventType %in% c('single', 'double', 'triple', 'home_run') & 
    details.call.description %in% c('In play, no out', 'In play, run(s)'), 1, 0)),
  RA = sum(runs_scored),
  .groups = 'keep') |>
  mutate(game_score = 40 + 2*Outs + SO - 2*BB - 2*H - 3*RA - 6*HR) |>
  ungroup() |>
#  mutate(ScoreRank = rank(desc(game_score), ties.method = 'min')) |>
#  filter(ScoreRank <= 10) |>
  arrange(desc(game_score)) |>
  select(matchup.pitcher.fullName, game_date, fielding_team, home_level_name, Outs, SO, BB, H, HR, RA, game_score)

```


```{r}
colnames(pitcherGame) <- c('Name', 'Date', 'Team', 'Level', 'Outs', 'K', 'BB', 'H', 'HR', 'RA', 'Score')

topGameOrg <- pitcherGame |>
  mutate(Org = case_when(
    Team %in% c('Norfolk Tides', 'Bowie Baysox', 'Aberdeen IronBirds', 'Delmarva Shorebirds') ~ 'BAL',
    Team %in% c('Worcester Red Sox', 'Portland Sea Dogs', 'Greenville Drive', 'Salem Red Sox') ~ 'BOS',
    Team %in% c('Scranton/Wilkes-Barre RailRiders', 'Somerset Patriots', 'Hudson Valley Renegades', 'Tampa Tarpons') ~ 'NYY',
    Team %in% c('Durham Bulls', 'Montgomery Biscuits', 'Bowling Green Hot Rods', 'Charleston RiverDogs') ~ 'TB',
    Team %in% c('Buffalo Bisons', 'New Hampshire Fisher Cats', 'Vancouver Canadians', 'Dunedin Blue Jays') ~ 'TOR',
    Team %in% c('Charlotte Knights', 'Birmingham Barons', 'Winston-Salem Dash', 'Kannapolis Cannon Ballers') ~ 'CWS',
    Team %in% c('Columbus Clippers', 'Akron RubberDucks', 'Lake County Captains', 'Lynchburg Hillcats') ~ 'CLE',
    Team %in% c('Toledo Mud Hens', 'Erie SeaWolves', 'West Michigan Whitecaps', 'Lakeland Flying Tigers') ~ 'DET',
    Team %in% c('Omaha Storm Chasers', 'Northwest Arkansas Naturals', 'Quad Cities River Bandits', 'Columbia Fireflies') ~ 'KC',
    Team %in% c('St. Paul Saints', 'Wichita Wind Surge', 'Cedar Rapids Kernels', 'Fort Myers Mighty Mussels') ~ 'MIN',
    Team %in% c('Sugar Land Space Cowboys', 'Corpus Christi Hooks', 'Asheville Tourists', 'Fayetteville Woodpeckers') ~ 'HOU',
    Team %in% c('Salt Lake Bees', 'Rocket City Trash Pandas', 'Tri-City Dust Devils', 'Inland Empire 66ers') ~ 'LAA',
    Team %in% c('Las Vegas Aviators', 'Midland RockHounds', 'Lansing Lugnuts', 'Stockton Ports') ~ 'OAK',
    Team %in% c('Tacoma Rainiers', 'Arkansas Travelers', 'Everett AquaSox', 'Modesto Nuts') ~ 'SEA',
    Team %in% c('Round Rock Express', 'Frisco RoughRiders', 'Hickory Crawdads', 'Down East Wood Ducks') ~ 'TEX',
    Team %in% c('Gwinnett Stripers', 'Mississippi Braves', 'Rome Emperors', 'Augusta GreenJackets') ~ 'ATL',
    Team %in% c('Jacksonville Jumbo Shrimp', 'Pensacola Blue Wahoos', 'Beloit Sky Carp', 'Jupiter Hammerheads') ~ 'MIA',
    Team %in% c('Syracuse Mets', 'Binghamton Rumble Ponies', 'Brooklyn Cyclones', 'St. Lucie Mets') ~ 'NYM',
    Team %in% c('Lehigh Valley IronPigs', 'Reading Fightin Phils', 'Jersey Shore BlueClaws', 'Clearwater Threshers') ~ 'PHI',
    Team %in% c('Rochester Red Wings', 'Harrisburg Senators', 'Wilmington Blue Rocks', 'Fredericksburg Nationals') ~ 'WSH',
    Team %in% c('Iowa Cubs', 'Tennessee Smokies', 'South Bend Cubs', 'Myrtle Beach Pelicans') ~ 'CHC',
    Team %in% c('Louisville Bats', 'Chattanooga Lookouts', 'Dayton Dragons', 'Daytona Tortugas') ~ 'CIN',
    Team %in% c('Nashville Sounds', 'Biloxi Shuckers', 'Wisconsin Timber Rattlers', 'Carolina Mudcats') ~ 'MIL',
    Team %in% c('Indianapolis Indians', 'Altoona Curve', 'Greensboro Grasshoppers', 'Bradenton Marauders') ~ 'PIT',
    Team %in% c('Memphis Redbirds', 'Springfield Cardinals', 'Peoria Chiefs', 'Palm Beach Cardinals') ~ 'STL',
    Team %in% c('Reno Aces', 'Amarillo Sod Poodles', 'Hillsboro Hops', 'Visalia Rawhide') ~ 'AZ',
    Team %in% c('Albuquerque Isotopes', 'Hartford Yard Goats', 'Spokane Indians', 'Fresno Grizzlies') ~ 'COL',
    Team %in% c('Oklahoma City Baseball Club', 'Tulsa Drillers', 'Great Lakes Loons', 'Rancho Cucamonga Quakes') ~ 'LAD',
    Team %in% c('El Paso Chihuahuas', 'San Antonio Missions', 'Fort Wayne TinCaps', 'Lake Elsinore Storm') ~ 'SD',
    Team %in% c('Sacramento River Cats', 'Richmond Flying Squirrels', 'Eugene Emeralds', 'San Jose Giants') ~ 'SF',
    .default = 'MLB'
  ) 
  ) |>
    select(Name, Team, Date, Level, Org, Outs, K, BB, H, HR, RA, Score)

topGameOrg
```

### Clean Up Date


```{r}
topGameOrg <- topGameOrg |>
  mutate(Date = as.Date(Date)) |>
  mutate(Date = format(Date, format = '%m-%d'))

topGameOrg
```



```{r}
pitcherGameExport <- topGameOrg |>
  group_by(Team, .keep_all = TRUE) |>
  mutate(ScoreRank = rank(desc(Score), ties.method = 'min')) |>
  filter(ScoreRank <= 10) 
```


```{r}
write.csv(pitcherGameExport, 'MiLB Best Games by Team.csv')
```


```{r}
topGameOrg <- topGameOrg |>
  mutate(ScoreRank = rank(desc(Score), ties.method = 'min')) |>
  filter(ScoreRank <= 10) |>
  select(Name, Team, Date, Level, Org, Outs, K, BB, H, HR, RA, Score)
```



```{r}
gameBestPitchers <- topGameOrg |>
  gt() |>
  tab_header(title = md('**Top MiLB Pitcher Outings by Tango Game Score, 2024**')) |>
  cols_align(align = 'center', columns = c('Team', 'Level', 'Date', 'Org', 'Outs', 'K', 'BB', 'H', 'HR', 'RA', 'Score')) |>
  opt_stylize(style = 5, color = 'blue') |>
  cols_width(Name ~ px(200), Date ~ px(80), Team ~ px(200), Level ~ px(80), everything() ~ px(60)) |>
  tab_footnote(footnote = 'Last Updated May 20th') |>
  gt_fmt_mlb_logo(columns = 'Org') |>
  tab_options(
    table.font.size = px(16)
  ) 

gameBestPitchers
```



## Season Performance: Best Hitters by OPS+

### Finding Season Averages


```{r}
abCheck <- allData |>
  distinct(paste(game_pk, atBatIndex), .keep_all = TRUE) |>
  group_by(paste(game_pk, atBatIndex), home_level_name) |>
  summarise(isAB = ifelse(!result.eventType %in% c('hit_by_pitch', 'sac_fly', 'intent_walk', 'walk', 'catcher_interf', 'sac_bunt'), 1, 0),
            slgCount = case_when(
              result.eventType == 'single' ~ 1,
              result.eventType == 'double' ~ 2,
              result.eventType == 'triple' ~ 3,
              result.eventType == 'home_run' ~ 4,
              .default = 0))

abByLeague <- abCheck |>
  group_by(home_level_name) |>
  summarise(abCount = sum(isAB),
            SLG = sum(slgCount) / sum(isAB))

abByLeague

  
obpAverages <- allData |>
  group_by(home_level_name) |>
  summarise(OBP = sum(ifelse(result.eventType %in% c('walk', 'hit_by_pitch', 'single', 'double', 'triple', 'home_run') &
            (count.balls.start == 4 | details.call.description %in% c('In play, no out', 'In play, run(s)', 'Hit By Pitch')),
            1, 0)) / length(unique(paste(game_pk, atBatIndex))))

colnames(obpAverages) <- c('Drop', 'OBP') #To remove duplicate column name created by grouping

leagueAverages <- cbind(obpAverages, abByLeague) |>
  select(home_level_name, abCount, OBP, SLG)

leagueAverages
```


### Individual Calculations


```{r}
indABCheck <- allData |>
  distinct(paste(game_pk, atBatIndex), .keep_all = TRUE) |>
  group_by(matchup.batter.id, matchup.batter.fullName, home_level_name, batting_team) |>
  summarise(isAB = ifelse(!result.eventType %in% c('hit_by_pitch', 'sac_fly', 'intent_walk', 'walk', 'catcher_interf', 'sac_bunt'), 1, 0),
            slgCount = case_when(
              result.eventType == 'single' ~ 1,
              result.eventType == 'double' ~ 2,
              result.eventType == 'triple' ~ 3,
              result.eventType == 'home_run' ~ 4,
              .default = 0))

abByPlayer <- indABCheck |>
  group_by(matchup.batter.id, matchup.batter.fullName, home_level_name, batting_team) |>
  summarise(abCount = sum(isAB),
            SLG = sum(slgCount) / sum(isAB))

abByPlayer

indyHittersBasic <- allData |>
  distinct(paste(game_pk, atBatIndex), .keep_all = TRUE) |>
  group_by(matchup.batter.id, matchup.batter.fullName, home_level_name, batting_team) |>
  summarise(PA = length(unique(paste(game_pk, atBatIndex))),
            SO = sum(ifelse(count.strikes.start == 3, 1, 0), na.rm = TRUE),
            BB = sum(ifelse(count.balls.start == 4, 1, 0), na.rm = TRUE),
            HR = sum(ifelse(result.eventType == 'home_run' & details.call.description == 'In play, run(s)', 1, 0),
                     na.rm = TRUE),
            H = sum(ifelse(
              result.eventType %in% c('single', 'double', 'triple', 'home_run') & 
              details.call.description %in% c('In play, no out', 'In play, run(s)'), 1, 0)))


indyHittersAdvanced <- allData |>
  distinct(paste(game_pk, atBatIndex), .keep_all = TRUE) |>
  group_by(matchup.batter.id, matchup.batter.fullName, home_level_name, batting_team) |>
  summarise(OBP = sum(ifelse(result.eventType %in% c('walk', 'hit_by_pitch', 'single', 'double', 'triple', 'home_run') &
            (count.balls.start == 4 | details.call.description %in% c('In play, no out', 'In play, run(s)', 'Hit By Pitch')),
            1, 0)) / length(unique(paste(game_pk, atBatIndex))))

indyHitters <- inner_join(indyHittersBasic, indyHittersAdvanced, by = c('matchup.batter.id', 'matchup.batter.fullName', 'home_level_name', 'batting_team'))
allHitters <- inner_join(indyHitters, abByPlayer, by = c('matchup.batter.id', 'matchup.batter.fullName', 'home_level_name', 'batting_team'))
```

```{r}
cleanHitters <- allHitters |>
  select(matchup.batter.id, matchup.batter.fullName, home_level_name, batting_team, abCount, PA, SO, BB, H, HR, OBP, SLG) |>
  ungroup() |>
  mutate(OPS = OBP + SLG)

allHitterInfo <- left_join(cleanHitters, leagueAverages, by = 'home_level_name')

allHitterClean <- allHitterInfo |>
  mutate(OPSPlus = (100*((OBP.x/OBP.y) + (SLG.x/SLG.y) - 1))) |>
  select(matchup.batter.fullName, home_level_name, batting_team, abCount.x, PA, OBP.x, SLG.x, OPS, OPSPlus) |>
  filter(PA >= mean(PA)) |>
  mutate(OPSPlusRank = rank(-OPSPlus),
         OBP.x = rd(OBP.x, 3),
         SLG.x = rd(SLG.x, 3),
         OPS = rd(OPS, 3),
         OPSPlus = round(OPSPlus, 0)) |>
  arrange(desc(OPSPlus))

colnames(allHitterClean) <- c('Name', 'Level', 'Team', 'AB', 'PA', 'OBP', 'SLG', 'OPS', 'OPS+', 'Rank')

allHitterClean
```


```{r}
write.csv(allHitterClean, 'MiLB Hitter 2024 Season.csv')
```


```{r}
allHitterClean <- allHitterClean |>
  filter(OPSPlusRank <= 10)
```


```{r}
topHitterSeason <- allHitterClean |>
  mutate(Org = case_when(
    Team %in% c('Norfolk Tides', 'Bowie Baysox', 'Aberdeen IronBirds', 'Delmarva Shorebirds') ~ 'BAL',
    Team %in% c('Worcester Red Sox', 'Portland Sea Dogs', 'Greenville Drive', 'Salem Red Sox') ~ 'BOS',
    Team %in% c('Scranton/Wilkes-Barre RailRiders', 'Somerset Patriots', 'Hudson Valley Renegades', 'Tampa Tarpons') ~ 'NYY',
    Team %in% c('Durham Bulls', 'Montgomery Biscuits', 'Bowling Green Hot Rods', 'Charleston RiverDogs') ~ 'TB',
    Team %in% c('Buffalo Bisons', 'New Hampshire Fisher Cats', 'Vancouver Canadians', 'Dunedin Blue Jays') ~ 'TOR',
    Team %in% c('Charlotte Knights', 'Birmingham Barons', 'Winston-Salem Dash', 'Kannapolis Cannon Ballers') ~ 'CWS',
    Team %in% c('Columbus Clippers', 'Akron RubberDucks', 'Lake County Captains', 'Lynchburg Hillcats') ~ 'CLE',
    Team %in% c('Toledo Mud Hens', 'Erie SeaWolves', 'West Michigan Whitecaps', 'Lakeland Flying Tigers') ~ 'DET',
    Team %in% c('Omaha Storm Chasers', 'Northwest Arkansas Naturals', 'Quad Cities River Bandits', 'Columbia Fireflies') ~ 'KC',
    Team %in% c('St. Paul Saints', 'Wichita Wind Surge', 'Cedar Rapids Kernels', 'Fort Myers Mighty Mussels') ~ 'MIN',
    Team %in% c('Sugar Land Space Cowboys', 'Corpus Christi Hooks', 'Asheville Tourists', 'Fayetteville Woodpeckers') ~ 'HOU',
    Team %in% c('Salt Lake Bees', 'Rocket City Trash Pandas', 'Tri-City Dust Devils', 'Inland Empire 66ers') ~ 'LAA',
    Team %in% c('Las Vegas Aviators', 'Midland RockHounds', 'Lansing Lugnuts', 'Stockton Ports') ~ 'OAK',
    Team %in% c('Tacoma Rainiers', 'Arkansas Travelers', 'Everett AquaSox', 'Modesto Nuts') ~ 'SEA',
    Team %in% c('Round Rock Express', 'Frisco RoughRiders', 'Hickory Crawdads', 'Down East Wood Ducks') ~ 'TEX',
    Team %in% c('Gwinnett Stripers', 'Mississippi Braves', 'Rome Emperors', 'Augusta GreenJackets') ~ 'ATL',
    Team %in% c('Jacksonville Jumbo Shrimp', 'Pensacola Blue Wahoos', 'Beloit Sky Carp', 'Jupiter Hammerheads') ~ 'MIA',
    Team %in% c('Syracuse Mets', 'Binghamton Rumble Ponies', 'Brooklyn Cyclones', 'St. Lucie Mets') ~ 'NYM',
    Team %in% c('Lehigh Valley IronPigs', 'Reading Fightin Phils', 'Jersey Shore BlueClaws', 'Clearwater Threshers') ~ 'PHI',
    Team %in% c('Rochester Red Wings', 'Harrisburg Senators', 'Wilmington Blue Rocks', 'Fredericksburg Nationals') ~ 'WSH',
    Team %in% c('Iowa Cubs', 'Tennessee Smokies', 'South Bend Cubs', 'Myrtle Beach Pelicans') ~ 'CHC',
    Team %in% c('Louisville Bats', 'Chattanooga Lookouts', 'Dayton Dragons', 'Daytona Tortugas') ~ 'CIN',
    Team %in% c('Nashville Sounds', 'Biloxi Shuckers', 'Wisconsin Timber Rattlers', 'Carolina Mudcats') ~ 'MIL',
    Team %in% c('Indianapolis Indians', 'Altoona Curve', 'Greensboro Grasshoppers', 'Bradenton Marauders') ~ 'PIT',
    Team %in% c('Memphis Redbirds', 'Springfield Cardinals', 'Peoria Chiefs', 'Palm Beach Cardinals') ~ 'STL',
    Team %in% c('Reno Aces', 'Amarillo Sod Poodles', 'Hillsboro Hops', 'Visalia Rawhide') ~ 'AZ',
    Team %in% c('Albuquerque Isotopes', 'Hartford Yard Goats', 'Spokane Indians', 'Fresno Grizzlies') ~ 'COL',
    Team %in% c('Oklahoma City Baseball Club', 'Tulsa Drillers', 'Great Lakes Loons', 'Rancho Cucamonga Quakes') ~ 'LAD',
    Team %in% c('El Paso Chihuahuas', 'San Antonio Missions', 'Fort Wayne TinCaps', 'Lake Elsinore Storm') ~ 'SD',
    Team %in% c('Sacramento River Cats', 'Richmond Flying Squirrels', 'Eugene Emeralds', 'San Jose Giants') ~ 'SF',
    .default = 'MLB'
  ) 
  ) |>
    select(Name, Team, Level, Org, AB, PA, OBP, SLG, OPS, `OPS+`)
```


```{r}
seasonBestHitters <- topHitterSeason|>
  gt() |>
  tab_header(title = md('**Top MiLB Hitters by OPS+, 2024**')) |>
  cols_align(align = 'center', columns = c('Team', 'Level', 'Org', 'AB', 'PA', 'OBP', 'SLG', 'OPS', 'OPS+')) |>
  opt_stylize(style = 5, color = 'blue') |>
  cols_width(Name ~ px(200), Team ~ px(200), Level ~ px(80), everything() ~ px(60)) |>
  tab_footnote(footnote = 'Last Updated May 22nd') |>
  gt_fmt_mlb_logo(columns = 'Org') |>
  tab_options(
    table.font.size = px(16)
  ) 

seasonBestHitters
beep(3)
```



## Game Performance: Best Hitters by Number of Bases


```{r}

```

