---
title: "Minor League Data Gathering"
author: "Alexander Harriman"
date: "2024-05-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



This file should be used as the basis for any file that requires pitch-by-pitch data across the 2024 season. The code below will load all data into the R environment (or a CSV if necessary) for analysis in other files.

# Finding Game PKs

ID for each level:

* MLB - 1
* AAA - 11
* AA - 12
* A+ - 13
* A - 14
* Rookie Ball - 16

```{r, warning=FALSE}
library(baseballr)
library(lubridate)
library(dplyr)
library(data.table)
library(beepr)

dateSequence <- seq(ymd('2024-04-05'), Sys.Date(), by = '1 day')

# AA, High-A, and Low-A start on April 5
## AAA start on March 29
### MLB Korea Series start on March 20, rest of league on March 28
#### Complex leagues start on May 4 and end July 25

gamePKs <- c()

dataAcquisition <- function(){
  for(i in 1:length(dateSequence)){
  allPK <- mlb_game_pks(date = dateSequence[i], level_ids = 12)$game_pk #Adjust level_id as needed, see above
  gamePKs <- append(gamePKs, allPK)
  }
  return(gamePKs)
}

pkList <- dataAcquisition()
```

# Getting PBP Data from Game PKs


```{r, warning = FALSE, error = TRUE}
pitchData <- list()

combiningData <- function(){
  for(i in 1:length(pkList)){
    skipFile <- FALSE
    tryCatch(
    {
      gameData <- mlb_pbp(game_pk = pkList[i]) |>
        filter(game_date != TRUE)
      pitchData[[i]] <- gameData
    },
    error = function(e){
      skipFile == TRUE
      print(gameData$game_pk[1])
    }
    )
  if(skipFile == TRUE){
    next
  }
  }
  return(pitchData)
}

dataList <- combiningData()

allData <- rbindlist(dataList, fill = TRUE)

dim(allData)[1]
```


# Data Save State

```{r}
write.csv(allData, '2024 AA May 3rd.csv')
beep(3)
```

Saved Data List:

2024 Low A: 2:38 PM May 3rd
2024 High A: 3:03 PM May 3rd


Approximate run time from start to save state: 20 mins.

# Data Load State

If a CSV needs to be loaded to save time. DO NOT INCLUDE IN CODE POSTED TO GITHUB

```{r}
lowAData <- read.csv('2024 Low A May 3rd.csv')
```




