---
title: "MLB Exit Velocity EVA Analysis V0.1"
author: "Alexander Harriman"
date: "2024-06-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(dplyr)
library(stringr)
library(tidyr)
library(weights)
library(beepr)
library(lubridate)
library(extRemes)
library(ismev)
```



# Load MLB Data

```{r}
mlb24 <- read.csv('2024 Full MLB August 12.csv')
mlb22 <- read.csv('2022 MLB Season.csv')
mlb23 <- read.csv('2023 MLB Season.csv')


mlbBIP <- bind_rows(mlb24, mlb22, mlb23) |>
  filter(!details.call.description %in% c('Foul', 'Foul Tip', 'Foul Bunt') & is.na(hitData.launchSpeed) == FALSE)
```







## Histogram of Exit Velocity

```{r}
plot(mlbBIP$hitData.launchSpeed~as.Date(mlbBIP$game_date),
     xlab = "Game Date", ylab = "Exit Velocity (MPH)",
     main = "2024 MLB Ball In Play EVs", type = "h",
     pch = 19, cex = 0.5, col = "lightsalmon",
     las = 1,
     yaxt = "n")
axis(2, at = seq(0, 120, 10), labels = seq(0, 120, 10), las = 1)
```


# Extract Daily Maximum Values


```{r, warning = FALSE}
maxEVs <- mlbBIP |>
  group_by(game_date) |>
  summarise(bip = n(),
            maxEV = max(hitData.launchSpeed))

maxEVs <- maxEVs |>
  mutate(maxEV = ifelse(maxEV == -Inf, 0, maxEV)) |>
  filter(maxEV != 0) |>
  mutate(game_date = row_number())
```



# EVA: Peak-Over-Threshold Method


## Find the Best Threshold: Mean Residual Life Plot

Choosing a quality threshold is up to the analyst. A Mean Residual Life Plot (MRL plot) is used as a visual guide: the threshold is placed where the plot begins to flatten into a horizontal line. 

```{r}
mrlplot(maxEVs$maxEV, main = 'Mean Residual Exit Velocity', xlab = 'Exit Velocity (MPH)', las = 1, xlim = c(110, 120), ylim = c(0, 5))

abline(v = 115, col = "blue", lty = 2)

threshold <- 115
```


Using this curve, the mean excess plot begins to flatten out around 115 MPH. Therefore, that will be the threshold set for this analysis.


## Plot the Time Series

FILL IN BEFORE THE END



# Find High-EV Balls in Play


```{r}
highEV <- which(maxEVs$maxEV > threshold)
extremEV <- maxEVs[highEV,]
length(highEV)
```






## Assess Best Fit



```{r}
#Fit the GPD
gpdFit <- fevd(c(maxEVs$maxEV), threshold = threshold, type = 'GP')

gpdFit


MLE <- gpdFit$results$par

##QQ-Plot
p <- 1:length(extremEV$maxEV) / (length(extremEV$maxEV) + 1)
qm <- gpdq(MLE, 115, 1 - p)



plot(qm, sort(extremEV$maxEV),
     pch = 16, cex = 1, col = alpha("blue", 0.5),
     xlab = "Model", ylab = "Empirical", main = "Quantile Plot",
     las = 1)
abline(0, 1, lwd = 1.5)
```




## Perform Inference on Return Levels



```{r}
#Find 1-week, 1-month, and 1-season return levels

##Using 7 days, 30 days, and 180 days
returnLevels <- return.level(gpdFit, return.period = c(7, 30, 180))

returnLevels
```



### Confidence Intervals


The 1-Month Confidence Interval will be used for these graphs.

```{r}
#1-Month Return Level
returnLevel <- return.level(gpdFit, return.period = 30)

#Delta CI

deltaCI <- ci(gpdFit, return.period = 30, verbose = F)
deltaCI


#Profile-Likelihood CI

profCI <- ci(gpdFit, method = 'proflik', xrange = c(115, 130), return.period = 30, verbose = F)
profCI


# Plot CIs

hist(extremEV$maxEV, col = alpha("lightblue", 0.2), border = "gray",
     xlim = c(115, 130), ylim = c(0, 120),
     xlab = "Exit Velocity (MPH)",
     main = "95% CI for 1-Month Maximum EV on Balls in Play")
xg <- seq(threshold, 10, len = 1000)
mle <- gpdFit$results$par
lines(xg, gpd.dens(mle, threshold, xg), lwd = 1.5)
abline(v = returnLevel, lwd = 1.5, lty = 2)
for (i in c(1, 3)) abline(v = deltaCI[i], lty = 2, col = "blue")
for (i in c(1,3)) abline(v = profCI[i], lty = 2, col = "red")
legend("topleft", legend = c("Delta CI", "Prof CI"),
       col = c("blue", "red"), lty = c(2, 3), bty = "n",
       lwd = 1.5)
```


