---
title: "Passer Rating Aging Curve Project V0.1"
author: "Alexander Harriman"
date: "2024-07-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



This project will create a baseline aging curve for NFL Passer Rating. After generating a curve, expected growth and decline will be used to determine which NFL QBs overperformed and underperformed in the 2023 season before providing a baseline expectation for QBs entering the 2024 season.



```{r, warning = FALSE, echo = FALSE}
library(dplyr)
library(tidyverse)
library(rvest)
library(xml2)
library(gam)
library(ggplot2)
```


# Load NFL Stats Data

Information will be loaded starting in the 2006 season, as ESPN QBR (found on Football Reference) dates back to this year.


```{r, warning = FALSE}
qbData <- data.frame()

for(i in 2006:2023){
  fbRefLink <<- paste0('https://www.pro-football-reference.com/years/',
                      i,
                      '/passing.htm')

  webScrape <- fbRefLink |>
    read_html() |>
    html_nodes('table') |>
    html_table()
  
  singleSeason <- webScrape[[1]]
  

  singleSeason <- singleSeason[-c(27)] |>
    mutate(Year = i) 

  #Clean up QB Names
  singleSeason$Player <- stringr::str_replace(singleSeason$Player, '\\*', '')
  singleSeason$Player <- stringr::str_replace(singleSeason$Player, '\\+', '')
  
  
  #Convert Stats to Numeric Values
  numericCols <- c(4, 6, 7, 9:31)
  
  singleSeason[numericCols] <- sapply(singleSeason[numericCols], as.numeric)
  
  qbData <- rbind(qbData, singleSeason)
   
}

#Filter down to QBs with 100+ Attempts
qbData <- qbData |>
  filter(Player != 'Player',
         Att >= 100)


qbData
```


# Adjust Statistics to Season Average

Each NFL season has a different passing environment for several factors (rule changes, injury variation, etc.). To compare each year on a level playing field, each statistic will be adjusted to a plus metric, similar to baseball. 100 will be NFL average, with higher representing better relative to their peers.


## Passer Rating


```{r}
avgPasserRating <- qbData |> 
  mutate(prWeight = Att * Rate) |>
  group_by(Year) |>
  summarise(avgPR = sum(prWeight) / sum(Att))

avgPasserRating
```



```{r}
for(i in 1:dim(qbData)[1]){
  qbData$adjustedRate[i] = (100.0*qbData$Rate[i]) / avgPasserRating[avgPasserRating$Year == qbData$Year[i],]$avgPR
}

qbData
```


## Quarterback Rating (QBR)


```{r}
avgQBR <- qbData |> 
  mutate(qbrWeight = Att * QBR) |>
  group_by(Year) |>
  summarise(avgQBRate = sum(qbrWeight, na.rm = TRUE) / sum(Att))

avgQBR
```



```{r}
for(i in 1:dim(qbData)[1]){
  qbData$adjustedQBR[i] = (100.0*qbData$QBR[i]) / avgQBR[avgQBR$Year == qbData$Year[i],]$avgQBRate
}

qbData
```



## Adjusted Net Yards Per Attempt (ANY/A)




```{r}
avgANYA <- qbData |> 
  mutate(anyaWeight = Att * `ANY/A`) |>
  group_by(Year) |>
  summarise(avgANYA = sum(anyaWeight, na.rm = TRUE) / sum(Att))

avgANYA
```



```{r}
for(i in 1:dim(qbData)[1]){
  qbData$adjustedANYA[i] = (100.0*qbData$`ANY/A`[i]) / avgANYA[avgANYA$Year == qbData$Year[i],]$avgANYA
}

qbData
```



## Yards Per Attempt

Used as a proxy for QB arm power



```{r}
avgYPA <- qbData |> 
  mutate(ypaWeight = Att * `Y/A`) |>
  group_by(Year) |>
  summarise(avgYPA = sum(ypaWeight, na.rm = TRUE) / sum(Att))

avgYPA
```



```{r}
for(i in 1:dim(qbData)[1]){
  qbData$adjustedYPA[i] = (100.0*qbData$`Y/A`[i]) / avgYPA[avgYPA$Year == qbData$Year[i],]$avgYPA
}

qbData
```




## Sack Percentage

Used as a proxy for awareness





```{r}
avgSack <- qbData |> 
  mutate(sackWeight = Att * `Sk%`) |>
  group_by(Year) |>
  summarise(avgSack = sum(sackWeight, na.rm = TRUE) / sum(Att))

avgSack
```



```{r}
for(i in 1:dim(qbData)[1]){
  qbData$adjustedSack[i] = (100.0*avgSack[avgSack$Year == qbData$Year[i],]$avgSack) / qbData$`Sk%`[i]
}

qbData
```




## Completion Percentage

Used as a proxy for accuracy 





```{r}
avgComp <- qbData |> 
  mutate(compWeight = Att * `Cmp%`) |>
  group_by(Year) |>
  summarise(avgComp = sum(compWeight, na.rm = TRUE) / sum(Att))

avgComp
```



```{r}
for(i in 1:dim(qbData)[1]){
  qbData$adjustedComp[i] = (100.0*qbData$`Cmp%`[i]) / avgComp[avgComp$Year == qbData$Year[i],]$avgComp
}

qbData
```





# Find Career Averages for Each Player


To account for selection bias at both ends of the aging curve, the player statistic in relation to their career average will also be included. 




```{r}
qbData <- qbData |>
  mutate(rateWeight = Att * adjustedRate,
         qbrWeight = Att * adjustedQBR,
         anyaWeight = Att * adjustedANYA,
         ypaWeight = Att * adjustedYPA,
         sackWeight = Att * adjustedSack,
         compWeight = Att * adjustedComp)



careerAvg <- qbData |>
  group_by(Player) |>
  summarise(Years = n(),
            Attempts = sum(Att),
            PasserRating = sum(rateWeight) / sum(Att),
            QBRate = sum(qbrWeight) / sum(Att),
            ANYA = sum(anyaWeight) / sum(Att),
            YPA = sum(ypaWeight) / sum(Att),
            SackRate = sum(sackWeight) / sum(Att),
            CompRate = sum(compWeight) / sum(Att))


careerAvg
```



## Join Season and Career Stats



```{r}
allInformation <- left_join(qbData, careerAvg, by = 'Player')


allInformation
```




# Create Aging Curve


The adjusted QB+ metrics will be used as part of separate aging curves to determine how player performance relative to league average age over time. A generative additive model (GAM) will be used to create these curves.

Since selection bias would keep only the best young and old players (and thus cause the aging curve to warp), the player's career average will be added to the model equation as an equalizing measure.



## Filter Out Application Data

The curve will only include players through the 2022 seasons; the 2023 data will be used as an application later in the project.



```{r}
inputData <- allInformation |>
  filter(Year <= 2022)

applicationData <- allInformation |>
  filter(Year == 2023)
```





## Passer Rate Aging Curve


```{r}
passerRateModel <- gam(adjustedRate ~ s(Age) + PasserRating, data = inputData)


summary(passerRateModel)
```



```{r}
plot(passerRateModel, se = TRUE)
```



```{r}
passerRateModel$coefficients

passerRateModel$fitted.values


plot(inputData$adjustedRate, passerRateModel$fitted.values)
```

