---
title: "Passer Rating Aging Curve Project V0.1"
author: "Alexander Harriman"
date: "2024-07-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



This project will create a baseline aging curve for NFL Passer Rating. After generating a curve, expected growth and decline will be used to determine which NFL QBs overperformed and underperformed in the 2023 season before providing a baseline expectation for QBs entering the 2024 season.


# Load NFL Stats Data


```{r}
qbData <- 
    list.files(pattern = "\\.csv$") %>% 
    map_df(~fread(.))

qbData <- data.frame()

fileList <- list.files(pattern = '\\.csv$')

for(i in 1:length(fileList)){
  season <- read.csv(fileList[i]) |>
    mutate(Year = 2005 + i)
  
  qbData <- rbind(qbData, season)
}


qbData <- qbData |>
  filter(GS > 0)

qbData$Player <- stringr::str_replace(qbData$Player, '\\*', '')
qbData$Player <- stringr::str_replace(qbData$Player, '\\+', '')

qbData
```



# Adjust Passer Rating to Season Average

Each NFL season has a different passing environment for several factors (rule changes, injury variation, etc.). To compare each year on a level playing field, passer rating will be adjusted to PasserRating+. 100 will be NFL average, with higher representing better relative to their peers.


```{r}
avgPasserRating <- qbData |> 
  mutate(prWeight = GS * Rate) |>
  group_by(Year) |>
  summarise(avgPR = sum(prWeight) / sum(GS))

avgPasserRating
```



```{r}
for(i in 1:dim(qbData)[1]){
  qbData$adjustedRate[i] = (100.0*qbData$Rate[i]) / avgPasserRating[avgPasserRating$Year == qbData$Year[i],]$avgPR
}

qbData
```



# Create Aging Curve


