---
title: "Passer Rating Aging Curve Project V0.1"
author: "Alexander Harriman"
date: "2024-07-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



This project will create a baseline aging curve for NFL Passer Rating. After generating a curve, expected growth and decline will be used to determine which NFL QBs overperformed and underperformed in the 2023 season before providing a baseline expectation for QBs entering the 2024 season.



```{r, warning = FALSE}
library(dplyr)
library(tidyverse)
library(rvest)
library(xml2)
```


# Load NFL Stats Data

Information will be loaded starting in the 2006 season, as ESPN QBR (found on Football Reference) dates back to this year.


```{r, warning = FALSE}
qbData <- data.frame()

for(i in 2006:2023){
  print(i)
  fbRefLink <<- paste0('https://www.pro-football-reference.com/years/',
                      i,
                      '/passing.htm')

  webScrape <- fbRefLink |>
    read_html() |>
    html_nodes('table') |>
    html_table()
  
  singleSeason <- webScrape[[1]]
  
  #Filter down to QBs with 100+ Attempts
  singleSeason <- singleSeason[-c(27)] |>
    mutate(Year = i) |>
    filter(Att >= 100)

  #Clean up QB Names
  singleSeason$Player <- stringr::str_replace(singleSeason$Player, '\\*', '')
  singleSeason$Player <- stringr::str_replace(singleSeason$Player, '\\+', '')
  
  
  #Convert Stats to Numeric Values
  numericCols <- c(4, 6, 7, 9:31)
  
  singleSeason[numericCols] <- sapply(singleSeason[numericCols], as.numeric)
  
  qbData <- rbind(qbData, singleSeason)
   
}

qbData <- qbData |>
  filter(Player != 'Player')


qbData
```


# Adjust Passer Rating to Season Average

Each NFL season has a different passing environment for several factors (rule changes, injury variation, etc.). To compare each year on a level playing field, passer rating will be adjusted to PasserRating+. 100 will be NFL average, with higher representing better relative to their peers.


```{r}
avgPasserRating <- qbData |> 
  mutate(prWeight = GS * Rate) |>
  group_by(Year) |>
  summarise(avgPR = sum(prWeight) / sum(GS))

avgPasserRating
```



```{r}
for(i in 1:dim(qbData)[1]){
  qbData$adjustedRate[i] = (100.0*qbData$Rate[i]) / avgPasserRating[avgPasserRating$Year == qbData$Year[i],]$avgPR
}

qbData
```



# Find Career Averages for Each Player


To account for selection bias at both ends of the aging curve, the player statistic in relation to their career average will also be included. 


# Create Aging Curve



Adjust the approach seen here to create an aging curve for the QB metrics:

model <- gam(stat ~ s(Age) + career_stat, data = dataset, weights = PA)



