---
title: "Passer Rating Aging Curve Project V0.1"
author: "Alexander Harriman"
date: "2024-07-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



This project will create a baseline aging curve for NFL Passer Rating. After generating a curve, expected growth and decline will be used to determine which NFL QBs overperformed and underperformed in the 2023 season before providing a baseline expectation for QBs entering the 2024 season.

The first step is to use a random forest to create a salary estimator for QBs based on passer rating and aging metrics. The passer rating will be normalized to the yearly average to account for changes in the NFL passing environment. A random forest will be used to project values, with the following inputs:

* Age during season
* Starts during the season (to account for injuries)
* Starts over their NFL career (also to account for injuries and as a placeholder for experience)
* Attempts during the last season (volume passer metric)
* Career attempts
* Passer rating during the season
* Career passer rating

After predicting salaries for all players through the 2023 season, the random forest model will be applied to the 2023 season to project what a player should have earned relative to their 2024 FA contract. All salary information is web scraped from Over The Cap.

Due to NFL inflation outpacing government economy inflation, the percentage of the cap consumed will be used instead.

QBs must have started at least one game to qualify for the model.

```{r}
library(data.table)
library(dplyr)
library(tidyverse)
library(randomForest)
library(rvest)
library(xml2)
library(stringr)
library(readr)
library(stats)
```


# Load NFL Stats Data


```{r}
#Load all CSVs in the folder

qbData <- 
    list.files(pattern = "\\.csv$") %>% 
    map_df(~fread(.))

qbData <- data.frame()

fileList <- list.files(pattern = '\\.csv$')

for(i in 1:length(fileList)){
  season <- read.csv(fileList[i]) |>
    mutate(Year = 2005 + i)
  
  qbData <- rbind(qbData, season)
}


qbData <- qbData |>
  filter(GS > 0)

qbData$Player <- stringr::str_replace(qbData$Player, '\\*', '')
qbData$Player <- stringr::str_replace(qbData$Player, '\\+', '')

qbData
```

# Adjust Passer Rating to Season Average

Each NFL season has a different passing environment for several factors (rule changes, injury variation, etc.). To compare each year on a level playing field, passer rating will be adjusted to PasserRating+. 100 will be NFL average, with higher representing better relative to their peers.


```{r}
avgPasserRating <- qbData |> 
  mutate(prWeight = GS * Rate) |>
  group_by(Year) |>
  summarise(avgPR = sum(prWeight) / sum(GS))

avgPasserRating
```



```{r}
for(i in 1:dim(qbData)[1]){
  qbData$adjustedRate[i] = (100.0*qbData$Rate[i]) / avgPasserRating[avgPasserRating$Year == qbData$Year[i],]$avgPR
}

qbData
```


# Gather Single-Season Information

For this salary analysis, single season age, single season games started, and single season adjusted passer rating will be used.


```{r}
singleSeason <- qbData |>
  select(Player, Year, Age, GS, Att, adjustedRate) |>
  mutate(rowID = paste(Player, Year))

singleSeason
```



# Find Career Since 2006

Information used will be career to date. 

```{r}
ytd <- data.frame()

for(i in 1:dim(qbData)[1]){
  qbName <- qbData$Player[i]
  qbYear <- qbData$Year[i]
  
  sampleSize <- qbData |>
    filter(Year <= qbYear & Player == qbName)
  
  playerYTD <- data.frame(Player = qbName,
                          Year = qbYear,
                          careerRate = sum(sampleSize$adjustedRate * sampleSize$GS) / sum(sampleSize$GS),
                          careerGames = sum(sampleSize$GS),
                          careerAtt = sum(sampleSize$Att),
                          rowID = paste(qbName, qbYear))
  
  ytd <- rbind(ytd, playerYTD)
}

ytd
```


# Find Last 3 Years Information


```{r}
lastThree <- data.frame()

for(i in 1:dim(qbData)[1]){
  qbName <- qbData$Player[i]
  qbYear <- qbData$Year[i]
  
  sampleSize <- qbData |>
    filter(Year <= qbYear & Year >= (qbYear - 2) & Player == qbName)
  
  playerLast3 <- data.frame(Player = qbName,
                          Year = qbYear,
                          last3Rate = sum(sampleSize$adjustedRate * sampleSize$GS) / sum(sampleSize$GS),
                          last3Games = sum(sampleSize$GS),
                          last3Att = sum(sampleSize$Att),
                          rowID = paste(qbName, qbYear))
  
  lastThree <- rbind(lastThree, playerLast3)
}

lastThree
```



# Web Scrape NFL Salary Information


```{r}
uniqueYears <- unique(singleSeason$Year)

salaryTable <- data.frame()


for(i in 1:length(uniqueYears)){
  
salaryLink <- paste0('https://overthecap.com/position/quarterback/',
                      uniqueYears[i])

seasonSalary <- salaryLink |>
  read_html() |>
  html_elements(xpath = '/html/body/div[1]/div/div[1]/div[2]/table') |>
  html_table()

salaryClean <- seasonSalary[[1]] |>
  mutate(Year = uniqueYears[i]) |>
  mutate(SalaryCap = case_when(
    uniqueYears[i] == 2014 ~ 133000000,
    uniqueYears[i] == 2015 ~ 143280000,
    uniqueYears[i] == 2016 ~ 155270000,
    uniqueYears[i] == 2017 ~ 167000000,
    uniqueYears[i] == 2018 ~ 177200000,
    uniqueYears[i] == 2019 ~ 188200000,
    uniqueYears[i] == 2020 ~ 198200000,
    uniqueYears[i] == 2021 ~ 182500000,
    uniqueYears[i] == 2022 ~ 208200000,
    uniqueYears[i] == 2023 ~ 224800000,
    .default = 0
  ))

salaryTable <- rbind(salaryTable, salaryClean)
}

salaryTable$`Cap Number` <- parse_number(salaryTable$`Cap Number`)

salaryTableClean <- salaryTable |>
  mutate(percentUsed = `Cap Number` / SalaryCap,
         rowID = paste(Player, Year))

salaryTableClean
```




# Join Salary and Career Information


```{r}
groupedData1 <- inner_join(singleSeason, lastThree, by = c('rowID', 'Player', 'Year'))
groupedData2 <- inner_join(groupedData1, ytd, by = c('rowID', 'Player', 'Year'))
allData <- inner_join(groupedData2, salaryTableClean, by = c('rowID', 'Player', 'Year'))


modelData <- allData |>
  select(Player, Year, Age, GS, Att, adjustedRate, last3Rate, last3Games, last3Att, careerRate, careerGames, careerAtt, percentUsed)


modelData
```



# Create the Random Forest Model

Use all information from 2016 (first year with a potential 3-year sample size) to 2023 for the model.

```{r}
inputs <- modelData |>
  filter(Year >= 2016 & Year < 2023) |>
  select(Age, 
         GS, 
         Att,
         adjustedRate, 
         last3Rate, 
         #last3Games, 
         last3Att, careerRate, #careerGames, 
         #careerAtt, 
         percentUsed)


# Tune a Model

tunedModel <- tuneRF(x = inputs[-8],
                     y = inputs$percentUsed,
                     ntreeTry = 1500,
                     mtryStart = 4,
                     stepFactor = 0.5,
                     improve = 0.0001)

tunedModel


salaryModel <- randomForest(percentUsed ~ ., data = inputs, ntree = 1500, mtry = 3)

plot(salaryModel)
varImpPlot(salaryModel)

#Model RMSE
sqrt(salaryModel$mse[which.min(salaryModel$mse)]) 
```


# Check Predictions vs. Actual Values on In-Sample Data


```{r}
predictions <- predict(salaryModel, inputs)


plot(inputs$percentUsed, predictions)
```



# Apply to 2023 Data to Predict 2024 Salary Numbers


```{r}
application <- allData |>
  filter(Year == 2023) |>
  select(Player, Age, GS, 
         Att,
         adjustedRate, 
         last3Rate, 
         last3Att, careerRate)



predictions <- predict(salaryModel, application)


allOutputs <- cbind(application, predictions)

allOutputs
```

