---
title: "Player Promotion Model"
author: "Alexander Harriman"
date: "2024-07-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This project will create a robust player promotion model containing in-season performance and prospect ranking information. Other information will be added as available and if valuable to the overall final output.

The result will be a percentage-based model predicting the odds of being called up to the next level, regardless of level. A 2/3 vs 1/3 split will be used for train/test splitting, with the inputs being 2022 and 2023 data across all levels (Rookie Ball, four full season levels, MLB). Then, once the model is validated, it will be applied to the 2024 season.


```{r}
library(tidyverse)
library(gt)
library(randomForest)
library(rvest)
library(xml2)
library(tidyr)
library(beepr)
library(purrr)
library(rvest)
library(xml2)
library(data.table)
```



# Load Data


```{r}
#2022 Season
filesMI2022 <- list.files(path = '.', pattern = '2022 MiLB Season')
filesML2022 <- list.files(path = '.', pattern = '2022 MLB Season')
files2022 <- c(filesMI2022, filesML2022)

dataList2022 <- map(files2022, read.csv)

data2022 <- rbindlist(dataList2022, fill = TRUE)
```



```{r}
#2023 Season
filesMI2023 <- list.files(path = '.', pattern = '2023 MiLB Season AA and Below')
filesML2023 <- list.files(path = '.', pattern = '2023 MLB Season')
filesAAA2023 <- list.files(path = '.', pattern = '2023 AAA Season')
files2023 <- c(filesMI2023, filesAAA2023, filesML2023)

dataList2023 <- map(files2023, read.csv)

data2023 <- rbindlist(dataList2023, fill = TRUE)
```



# Find Promotion Status of All Players


A player will be considered promoted if they satisfy one of the following:

* The player started the season at a lower level and ended the season playing 10+ games at a higher level.
* The player played 10+ games at a level higher than their starting level, even if they were later demoted. 


Players called up as temporary injury replacements (fewer than 10 games played) do not count.

If a player was promoted twice in the same season, their data will count as one Yes, not two. 

Both Rookie Ball levels (DSL, Complex Leagues) are considered as separate leagues, but only moving from the DSL to the Complex Leagues is a promotion for this model.

Demotions are not considered. If a player gets demoted, then called back up to the level they started out at, this is not a promotion for the purposes of the model.


```{r}
promo2022
```



# Player Position and Value Web Scraping

This information will be gathered from FanGraphs' Top X Prospects articles from the start of each season. The data scraped includes Age, Position, and Future Value, all of which will be included in the model.



## List of Links

Without a consistent format for links, a data frame containing the proper information must be used to elegantly scrape all 90 websites necessary for the project.


```{r}
scrapingFrame <- data.frame(Team = c(),
                            Number = c(),
                            Year = c())
```



## Gather Prospect Data

```{r}
propsectRankings <- data.frame()

for(i in 1:90){
  fanGraphsLink <- paste0('https://blogs.fangraphs.com/-',
                           scrapingFrame$Team[i],
                           '-top-',
                           scrapingFrame$Number[i],
                           '-prospects/',
                           scrapingFrame$Year[i])

  rosters <- fanGraphsLink |>
  read_html() |>
  html_nodes(xpath = '//*[@id="blogcontent"]/div[1]/article/div[2]/div[3]/div[3]/table') |>
  html_table()
  
  prospectRankings <- rbind(prospectRankings, rosters)
}


prospectRankings
```

